### 2015-05-03 abstract.tex
stuff

\selectlanguage{english}
% \renewcommand{\abstractname}{\Large{The \lego Motorcyclist robot}}
\addcontentsline{toc}{chapter}{\numberline{}Abstract}
\begin{abstract}
	\begin{verse}
	\small \textit{
	by Quentin Delhaye, Arnaud Dumont, Camille Giaux, Pierre Lasbleis and  Gauthier Roig; Université Libre de Bruxelles, 2011–2012.}
	\normalsize
	\end{verse}
\end{abstract}

\clearpage

\selectlanguage{francais}
% \renewcommand{\abstractname}{\Large{Le robot motard \lego}}
\addcontentsline{toc}{chapter}{\numberline{}Résumé}
\begin{abstract}
	\begin{verse}
	\small \textit{
	par Quentin Delhaye, Arnaud Dumont, Camille Giaux, Pierre Lasbleis et Gauthier Roig ; Université Libre de Bruxelles, 2011–2012.}
	\normalsize
	\end{verse}
\end{abstract}

### 2015-05-03 aknowledgements.tex
stuff

\addcontentsline{toc}{chapter}{\numberline{}Aknowledgements}
\chapter*{Aknowledgements}

### 2015-05-03 main.tex
stuff

\documentclass[a4paper, 12pt, oneside]{memoir}

\input{preamble}

\begin{document}

\frontmatter

\include{PageDeGarde}

\include{abstract}

\include{aknowledgements}

\selectlanguage{english}
\tableofcontents \pagebreak

\mainmatter

\appendix

\backmatter

% Bibliography

\end{document}

### 2015-05-03 introduction.tex
stuff

\chapter{Introduction}

\section{Challenge}

\section{Network security}

### 2015-05-03 presentation.tex
stuff

\chapter{Presentation}
Here we talk about the protocols, the platform.
Show The OS stack (kernel/user)

\section{Experimental setup}
The experimental environment is build around a standard x86 host and an ARM Cortex-A9 alongside an Altera Cyclone V FPGA as the target.

\subsection{x86 host}

\begin{framed}
\begin{description}
	\item[OS] Ubuntu 12.04 LTS, kernel 3.16
	\item[CPU] Intel Core-i3 ... (two logical core out of four)
	\item[RAM] 1.5GB DDR3
\end{description}
\end{framed}

\subsection{Altera Socrates SoCFPGA}

\begin{framed}
\begin{description}
	\item[OS] Yocto project, kernel 3.14
	\item[CPU] Dual core ARM Cortex-A9, 800MHz
	\item[RAM] ...GB DDR3
	\item[FPGA] Altera Cyclone V
\end{description}
\end{framed}

### 2015-05-04 conclusion.tex
conclusion

\chapter{Conclusion}

\section{Future work}

Although the present work presents some promissing results, the implementation can certainly be improved in several ways and some further experiments should be conducted.

\paragraph{Driver improvement}
The driver of the BA411E can be made less ressources hungry by improving the initialisation of the descriptors and their linking, but the gain would not be significant enough to justify the time investment at this point.
% TODO refer to a stack analysis of an ipsec file transfer, showing the most expensive instructions. If I remember correctly, it was something like the DMA transfer and DMA something freeing.
A better alternative would be to avoid descriptors altogether by modifiying the interface with the IP so that it can use the scatterlist directly.
We would then spare a lot of DMA mapping instruction and thus some precious cycles on the software side.

As we already remarked in \ref{sec:section-where-we-see-the-CPU-usage-of-ba411e-usage-with-active-polling}, the use cases involving IPSec were conducted using a previous revision of the driver still activelly polling the IP for its results.
A better and cleaner way to proceed is to use interruption routines, as shown in \ref{sec:section-ba411e-with-irq}.
However, the kernel does not support their current implementation and panics upon usage.
If one were to be willing to spend the time replacing the active polling by clean asynchronous interruptions, he should be aware of the overhead imposed by an interruption.
In some cases, when the operation is just a few clock cycle long for the IP, an active polling could still be the better way to go.
A more thorough comparison of the mutual trade-off deserves some investigation, and as a starting point, the packet size could be treated as a branching point between the two solutions.

\paragraph{Registering public key verification with the crypto API}
As we saw in \ref{sec:section-with-RSA-BA414}, the driver is already capable of offloading a large portion of public key operations to the IP, but only with very specific libraries at the time being -- openSSL in our case.
The next step is to register the very same operations with the crypto API so it can be used without having to rely on a custom openSSL engine.

\paragraph{Conditional offloading in cryptodev}
%Note this is still under heavy criticism. If the hardware lags so far behind, why the relatively good score in openvpn?
The figure \ref{fig:openssl-benchmark-cbc} clearly shows a threshold on the packet size from which the hardware has a clear advantage, and below which the user mode software implementation is to go for.
%TODO include a listing to where it could be a good place to branch on the packet size. I already found it.
%TODO at least add a pseudo-code to illustrate this shit.
Using this tipping point, one could set a conditional branch as shown in listing~\ref{list:cryptodev-conditional-offload-pseudocode}

\lstset{language=c}
\lstset{commentstyle=\color{colComments}\textit,
float=hbp,%
basicstyle=\ttfamily\small, %
identifierstyle=\color{colIdentifier}, %
keywordstyle=\color{colKeys}, %
stringstyle=\color{colString}, %
columns=flexible, %
tabsize=2, %
extendedchars=true, %
showspaces=false, %
showstringspaces=false, %
numbers=left, %
numberstyle=\tiny, %
breaklines=true, %
breakautoindent=true, %
captionpos=b}
\begin{lstlisting}[caption=cryptodev conditional offloading, label=list:cryptodev-conditional-offload-pseudocode]
int some_function() {
	if(packet_size < 1024*1024) {
		callback_function();
	}
}
\end{lstlisting}

He should however be aware that as the tipping point is around 1024kB, the performance for a network application should very close to those of a full software implementation, knowing that the ethernet frame size, the MTU, is set by default at 1500kB.

% This work is just the beginning of something far larger. The aim is to build a fully secured platform, from software to hardware, using maximum security at each stage.

### 2015-05-05 conclusion.tex
daily update

@@
 
+\citet{Xenakis20063225}
+
+\paragraph{Disk encryption}
+As the hardware is better used with larger blocks of data, disk encryption could be an interesting application to look into.
+
+\paragraph{Cryptographic libraries}
+OpenSSL is not the only cryptographic library available; PolarSSL and GnuTLS could also be viable solutions.
+%TODO see if they can use cryptodev, that would be perfect.
+
+\paragraph{Cryptodev}
+If patches adding the GCM support to cryptodev have already been released, those are not compatible with Barco Silex' driver.
+Adapting the interface would open the GCM hardware offload to the whole user space applications park.
+
+\paragraph{MAC offloading}
+While the symetric and asymetric encryption ciphers IPs are usable from the operating system, the IP computing MACs does not have a usable driver yet.
+Wherever there is encryption, authentication is also needed.
+As such, any real day-to-day use case can not be fully offloaded to hardware yet, even if some tricks and patches allowed us to bypass this requirement.
+The implementation of the GCM mode, combining encryption and authentication, showed us that stopping relying on the software implementation of MACs would be a huge step forward.
+
 % This work is just the beginning of something far larger. The aim is to build a fully secured platform, from software to hardware, using maximum security at each stage.

### 2015-05-06 conclusion.tex
Daily update: conclusion

@@
 
-\citet{Xenakis20063225}
-
 \paragraph{Disk encryption}
@@
 \paragraph{Cryptographic libraries}
-OpenSSL is not the only cryptographic library available; PolarSSL and GnuTLS could also be viable solutions.
-%TODO see if they can use cryptodev, that would be perfect.
+OpenSSL is not the only cryptographic library available; GnuTLS is also a very popular alternative and supports cryptodev engines too.
+
+However, one library definitely worth to keep an eye on is mbed TLS, formally known as PolarSSL, recently bought by ARM~\cite{2015-arm-buy-polarssl}.
+We can expect the future releases of this library to be more optimized for ARM platforms, and maybe the software footprint and overhead to be reduced.
 

### 2015-05-08 abstract.tex
stuff
daily update

@@
+\thispagestyle{plain}% This way, we can have the page numbering in the foot.
 \selectlanguage{english}
@@
 	\small \textit{
-	by Quentin Delhaye, Arnaud Dumont, Camille Giaux, Pierre Lasbleis and  Gauthier Roig; Université Libre de Bruxelles, 2011–2012.}
+	by Quentin Delhaye,  Master in Computer Science and Engineering, Professional Focus, Université Libre de Bruxelles, 2014--2015.}
 	\normalsize
 	\end{verse}
+	\begin{center}
+		\textbf{Implementation of high level cryptographic protocol using a SoC platform}
+	\end{center}
+This is an abstract.
 \end{abstract}
-
 \clearpage
 
+\thispagestyle{plain}
 \selectlanguage{francais}
@@
 	\small \textit{
-	par Quentin Delhaye, Arnaud Dumont, Camille Giaux, Pierre Lasbleis et Gauthier Roig ; Université Libre de Bruxelles, 2011–2012.}
+	par Quentin Delhaye, Master en ingénieur civil en informatique, à finalité spécialisée, Université Libre de Bruxelles, 2014--2015.}
 	\normalsize

### 2015-05-08 technical-background.tex
daily update

\chapter{Technical background}

This chapter will address the technical ground inherent to this work: first the operating system, followed by the cryptography and networking.

\section{Operating system}

\section{Cryptography}

\subsection{Asymetric cryptography}

\subsection{Symetric cryptography}

\section{Network and VPN implementation}

% Present the TCP/IP layering

% First explain what a VPN is.

There exist several major implementations of VPN: SSL, IPSec and PPTP.
The later was developped by a vendor consortium and proposed in the RFC 2637 and will not be discussed further.

\subsection{SSL/TLS}
% Introduce SSL/TLS, talk about the protocol, the key exchange and stuff, but leaver OpenVPN for the 'implementation' chapter.
% Question the security? Apparently, it does mac-then-encrypt, which is insecure regarding certain types of attacks \cite{cryptoeprint:2001:045}.
% Plus, SSLv3 is to be deprecated, according to a queued RFC: http://www.rfc-editor.org/internet-drafts/draft-ietf-tls-sslv3-diediedie-03.txt

\subsection{IPSec}
% IPSec as a protocol, strongswan come in the 'implementation' chapter.

### 2015-05-15 conclusion.tex
openvpn stuff etc

@@
-\chapter{Conclusion}
+\chapter{Discussion and conclusion}
 
@@
 The next step is to register the very same operations with the crypto API so it can be used without having to rely on a custom openSSL engine.
+This feature would require to work very closely to the linux kernel developpement.
+Indeed, if the signature verification using public key cryptography has been available in the kernel since 2013, a public key encryption API has only been introduced in late April 2015~\cite{cyrpto-api-pk-encryption} and is still under request for comments.
 

### 2015-05-15 implementation.tex
openvpn stuff etc

\chapter{Implementation}

\section{Software}

\subsection{OpenVPN}
% By default, point-to-point, but since the version 2.0 can be configured in server mode, hence being able to manage hundreds of clients.

% Pages 20-21, I have the workflow of openvpn, showing the encryption happening before the fragmentation.
\lstset{language=c}
\lstset{commentstyle=\color{colComments}\textit,
float=hbp,%
basicstyle=\ttfamily\footnotesize, %
identifierstyle=\color{colIdentifier}, %
keywordstyle=\color{colKeys}, %
stringstyle=\color{colString}, %
columns=flexible, %
tabsize=2, %
extendedchars=true, %
showspaces=false, %
showstringspaces=false, %
numbers=left, %
numberstyle=\tiny, %
breaklines=true, %
breakautoindent=true, %
captionpos=b}
\begin{lstlisting}[caption=openvpn compress then encrypt -- sample from \texttt{forward.c}, label=list:]
/* Compress, fragment, encrypt and HMAC-sign an outgoing packet. */
void encrypt_sign (struct context *c, bool comp_frag)
{
	struct context_buffers *b = c->c2.buffers;
	const uint8_t *orig_buf = c->c2.buf.data;

	if (comp_frag){
	/* Compress the packet. */
		if (lzo_defined (&c->c2.lzo_compwork))
			lzo_compress (&c->c2.buf, b->lzo_compress_buf, &c->c2.lzo_compwork, &c->c2.frame);
	/* Fragment the packet. */
		if (c->c2.fragment)
			fragment_outgoing (c->c2.fragment, &c->c2.buf, &c->c2.frame_fragment);
	}

  /* Encrypt the packet and write an optional HMAC signature. */
	openvpn_encrypt (&c->c2.buf, b->encrypt_buf, &c->c2.crypto_options, &c->c2.frame);
}
\end{lstlisting}

\begin{figure}[ht]
\includegraphics[width=\textwidth]{OpenVPN-packetflow}
\caption{OpenVPN packet flow as advertized on the openVPN wiki.}
\label{fig:openvpn-packet-flow}
\end{figure}

\subsection{OpenSSL}
% Talk about the debug flags in the compilation and show it has close to none impact on the performance (using openssl speed. openssl speed is a relevant test because openssl uses 100% of the CPU, hence maximizing the potential performance reduction.)

% Talk about the assembly implementation, maybe show the difference in an appendice with the C implementation. Not /that/ interresting.

\subsection{OpenSSH}
% Talk about the MAC=none patch
% connection using SSL/TLS, then AES, but without fragmentation (unlike openvpn)

\subsection{Strongswan}

\subsection{Linux drivers}
% Mention the fact that we use the assembly implementation of AES, not the generic C one.

\section{Offload}

\subsection{Silex engine}

\subsection{BA411E Driver}

\subsection{BA414E Driver}

### 2015-05-15 presentation.tex
openvpn stuff etc

@@
 \section{Experimental setup}
-The experimental environment is build around a standard x86 host and an ARM Cortex-A9 alongside an Altera Cyclone V FPGA as the target.
+The experimental environment is built around a standard x86 host and an ARM Cortex-A9 alongside an Altera Cyclone V FPGA as the target.
+Both are linked toghether through a network capped by 100Mbps switches.
+Both stations have gigabits ethernet interface and could hence be diretly connected to each other, but in that case the communication would be limited by the I/O transfers of the storage units -- a hard drive disk in one case, an micro-SD card in the second -- on which we can not depend to set a constant throughput limitation, as it is highly influenced by the data block size and general health of the support.
 
 \subsection{x86 host}
+The desktop host runs on Windows 7 Professional 64-bit, but a virtual machine using a Linux distribution is used for the developpement and testings.
 
@@
 	\item[CPU] Intel Core-i3 ... (two logical core out of four)
-	\item[RAM] 1.5GB DDR3
+	\item[RAM] 1GB DDR3
 \end{description}
@@
 \end{framed}
+
+\subsection{ARM DS-5 Streamline}
+% Use a daemon and a driver to get system info and send them to another machine. Has almost no impact on the performance and the load of the system.
+
+\subsection{Barco Silex' IPs}
+% table with the BA411E and BA414E showing the respective operations and crypto alg

### 2015-05-15 results.tex
openvpn stuff etc

\chapter{Results}

\section{Response time -- latency}

\section{TLS connections}

\section{File transfer}
%FTP over openvpn and IPSec

%OpenSSH: 	- normal transfer: shows perf difference
%			- capped to software max: show CPU offload

### 2015-05-17 conclusion.tex
diffie-hellman

@@
 %TODO at least add a pseudo-code to illustrate this shit.
-Using this tipping point, one could set a conditional branch as shown in listing~\ref{list:cryptodev-conditional-offload-pseudocode}
+Using this breakeven point, one could patch the cryptodev cryptodev engine of OpenSSL to branch on the packet size, using the hardware if the packet is large enough, or fallback on the default software implementation otherwise.
+%openssl-1.0.2a/crypto/engine/eng_cryptodev.c
+It would probably not be as trivial and beneficial as it may sound:
+\begin{itemize}
+	\item the encryption contexts would need to be synchronized between the hardware and the software;
+	% \item we would already suffer from some context switches when sending the information to the cryptodev driver; %No, we are still in the openssl engine.
+	% \item if we always branch to the fallback functions, the time overhead due to crytodev would lower the performance; %No, we don't leave the user space, the overhead is minimum.
+	\item as the breakeven point is around 1024kB, the performance for a network application would be very close to those of a full software implementation, knowing that the ethernet frame size, the MTU, is set by default at 1500 bytes.
+\end{itemize}
 
-\lstset{language=c}
-\lstset{commentstyle=\color{colComments}\textit,
-float=hbp,%
-basicstyle=\ttfamily\small, %
-identifierstyle=\color{colIdentifier}, %
-keywordstyle=\color{colKeys}, %
-stringstyle=\color{colString}, %
-columns=flexible, %
-tabsize=2, %
-extendedchars=true, %
-showspaces=false, %
-showstringspaces=false, %
-numbers=left, %
-numberstyle=\tiny, %
-breaklines=true, %
-breakautoindent=true, %
-captionpos=b}
-\begin{lstlisting}[caption=cryptodev conditional offloading, label=list:cryptodev-conditional-offload-pseudocode]
-int some_function() {
-	if(packet_size < 1024*1024) {
-		callback_function();
-	}
-}
-\end{lstlisting}
-
-He should however be aware that as the tipping point is around 1024kB, the performance for a network application should very close to those of a full software implementation, knowing that the ethernet frame size, the MTU, is set by default at 1500kB.
+Such a conditional offloading would be interresting for applications involving mainly very large packets and a few periodic smaller ones, like large data transfer between two hosts on a infrastructure supporting ethernet jumbo frames\footnote{Jumbo frames have an ethernet MTU of 9000 bytes, whilst standard frames are set to 1500.} with periodic ICMP heartbeat.
 

### 2015-05-17 technical-background.tex
openvpn stuff etc
diffie-hellman

@@
 
+\subsection{Symetric cryptography}
+
 \subsection{Asymetric cryptography}
+% Diffie-Hellman: show the math about the shared key, maybe from the course INFOH405 or RFC2631, and point which operation will be offloaded.
+Asymetric cryptography relies on a pair of keys: one private known only to the owner of the certificate, and one public available to anyone.
+Such cryptography uses two kinds of operations: encryption using the public key of the recipient and digital signature using the private key of the sender.
 
-\subsection{Symetric cryptography}
+\subsubsection{RSA}
+
+\subsubsection{Diffie-Hellman}
+Diffie-Hellman is a secret key exchange protocol: two parties compute a shared secret $ZZ$ that can be used as a symetric key during the following exchanges.
+It uses the same kind of operation as RSA, that is modular exponentiation.
+The RFC 2631~\cite{rfc2631} decribes the protocol:
+
+\begin{enumerate}
+	\item Alice generates once $p$ and $g$ (using precomputed parameters):
+	\begin{description}[nosep]
+		\item[p] large prime number
+		\item[g] a generator of $\mathds{Z}_p^*$
+	\end{description}
+	\item Alice picks a random integer $x_a$ and computes $g^{x_a} mod\ p = y_a$.
+	\item Alice sends $p$, $g$ and $y_a$ to Bob, signing everything using her private certificate.
+	\item Bob checks the signature and picks $x_b$.
+	\item Bob computes $y_a^{x_b} mod\ p = g^{x_a x_b} mod\ p = ZZ$, the shared secret to use as a premaster key from which will be derived the symetric key for further communications.
+	\item Bob sends $y_b = g^{x_b} mod\ p$, signing everything with his private certificate.
+	\item Alice checks the signature and computes the same shared secret: $ZZ = y_b^{x_a} mod\ p = g^{x_a x_b}$
+\end{enumerate}
+
+If the server is Alice, it has to do at least one signature, one signature verification and two modular exponentiations.
+If the certificate use RSA, we end up with four modular exponentiations, which can become quite heavy computing wise for certain sizes of prime numbers.
+We will see in chapter~\ref{chap:results} that while a 1024-bit prime is easily manageable by full software implementation, hardware offloading become a necessity for 4096-bit primes.
+Moreover, 1024-bit parameter size, both RSA and Diffie-Hellman, are disallowed by the NIST recommandations since 2013~\cite{nist-sp800-131A}.
 

### 2015-05-17 technical-background.tex
RSA

@@
 \subsection{Symetric cryptography}
+Talk about encryption, integrity and authentication.
+
+\subsubsection{AES}
+Many modes, CBC is mainly used, GCM is great.
+
+\subsubsection{SHA}
+Keyed signature algorithm, several versions in place. SHA-1 is depreciated, SHA-2 is widely used and SHA-3 is already defined and begins to be implemented.
 
@@
 Asymetric cryptography relies on a pair of keys: one private known only to the owner of the certificate, and one public available to anyone.
-Such cryptography uses two kinds of operations: encryption using the public key of the recipient and digital signature using the private key of the sender.
+Such cryptography uses two kinds of operations: encryption using the public key of the recipient and digital signature, which is an ecnryption using the private key of the sender.
 
 \subsubsection{RSA}
+RSA is a public-key scheme proposed in 1978 by three MIT researchers who gave it their name~\cite{Rivest:1978:MOD:359340.359342}.
+A few years later, they founded RSA Laboratories, which is now in charge of maintaining its standards, alongside many others, as the first Public-Key Cryptography Standards, \textit{aka} PKCS \#1.
+The last version of the standard is the version 2.2~\cite{pkcs1} and is defined as a precise key generation protocol allowing encryption and decryption.
+The keys can be generated by respecting a few steps:
+\begin{enumerate}
+	\item randomy choose two large primes $p$ and $q$;
+	\item compute the modulus $n = p q$, and consequently we have $\phi(n) = (p-1)(q-1)$, with $\phi(n)$ as the Euler function;
+	\item randomly choose the public exponent $e \in ]1,\phi(n)[\ s.t.\ GCD(e,\phi(n)) = 1$;
+	\item compute $d \in ]1,\phi(n)[\ s.t.\ e \cdot d \equiv 1 (mod\ \phi(n))$
+\end{enumerate}
+
+With those parameters, we can form a public key with the pair $(n, e)$ and a private key with the pair $(n, d)$.
+
+The encryption and decryption of a given message $m \in \mathds{Z}_n$ are defined as follows:
+\begin{description}
+	\item[Encryption] $c = m^e\ mod\ n$
+	\item[Decryption] $m = c^d\ mod\ n$
+\end{description}
 
@@
 	\item Bob sends $y_b = g^{x_b} mod\ p$, signing everything with his private certificate.
-	\item Alice checks the signature and computes the same shared secret: $ZZ = y_b^{x_a} mod\ p = g^{x_a x_b}$
+	\item Alice checks the signature and computes the same shared secret: $ZZ = y_b^{x_a} mod\ p = g^{x_a x_b} mod\ p$
 \end{enumerate}

### 2015-05-17 implementation.tex
diffie-hellman
ipsec implementation

@@
 \subsection{Strongswan}
+Strongswan is a full implementation of IPSec relying on the kernel drivers for the networking part, on the crypto API for the cryptographic part, and on user space crypto libraries for the connection negociation.
+An other popular implementation is ipsec-tools, but its development lags behind modern Linux and is not up-to-date with the 3.14 Linux kernel headers, making its cross-compilation painfull.
+%Use strongswan 5.3.0
+Strongswan as two advantages: it has a tremendous and exhaustive documentation, and its uer interface is straightforward.
+Once configured, a simple \texttt{ipsec start \&\& ipsec up <connection>} on both sides is enough to create a ready to use VPN.
+
+The figure~\ref{fig:ipsec-workflow} illustrates the workflow of Alice communicating with Bob via an IPSec ESP tunnel.
+The XFRM, read "transform", framework is implementing IPSec and handles the incomming and outgoing packets for established VPNs~\cite{rosen2014}.
+Its name comes from the fact that the kernel transforms packet frames to incorporate IPSec security.
+Depending on the configuration, XFRM uses the AH or ESP kernel module, which in turn calls the crypto API to encrypt and/or sign the IP packet.
+
+We can also clearly see one of the main advantages of IPSec: it works in the kernel space.
+Since it does not require a virtual network interface like OpenVPN, the only transfer between the user/kernel space happens when the former wishes to send a packet on the network, passing it to the later -- or \textit{vice versa} for incomming packets.
+
+\begin{figure}[ht]
+\Large
+\resizebox{\linewidth}{!}{%
+\input{ipsec-transfer}
+}
+\caption{IPSec user/kernel space workflow, using \texttt{ping} as a test case.}
+\label{fig:ipsec-workflow}
+\end{figure}
 

### 2015-05-17 implementation.tex
Move openssl bench dgb on/off to implem.

@@
 
-% Pages 20-21, I have the workflow of openvpn, showing the encryption happening before the fragmentation.
 \lstset{language=c}
@@
 
+It is to be noted that the present work uses an implementation with all the debug flags activated.
+If it does have an impact on the performance, it is minimal: in the worst case of the benchmark, the throughput drops only by 2.4\%.
+Moreover, the benchmark maximizes this difference by doing only OpenSSL operations.
+When OpenSSL will have to share the CPU with other applications, the loss will be even less noticeable.
+\begin{figure}[ht]
+\begin{tikzpicture}
+\begin{axis}[
+		title = {Openssl speed with debug on/off},
+        width  = 0.85*\textwidth,
+        height = 8cm,
+        major x tick style = transparent,
+        ybar,
+        bar width=8pt,
+        ymajorgrids = true,
+        ylabel = {Throughput [$10^3$kB/s]},
+        xlabel = {Packet size},
+        ymin=0,
+        symbolic x coords={16,64,256,1024,8192},
+        xtick = data,
+        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
+        legend style={at={(0.5,-0.25)},	anchor=north,legend columns=2},
+        enlarge x limits=0.1,
+    ]
+\addplot[style={MidnightBlue,fill=MidnightBlue,mark=none}]
+	coordinates {(16,18400.60) (64,20750.61) (256,21507.33) (1024,21681.15) (8192,21725.18)};
+	\label{aes-256-cbc-dbg-off}
+
+\addplot[style={Aquamarine,fill=Aquamarine,mark=none}]
+	coordinates {(16,19091.65) (64,20579.24) (256,21106.60) (1024,21216.26) (8192,21203.63)};
+	\label{aes-256-cbc-dbg-on}
+
+\addplot[style={ForestGreen,fill=ForestGreen,mark=none}]
+	coordinates {(16,23882.41) (64,27781.80) (256,29285.63) (1024,29692.25) (8192,29807.96)};
+	\label{aes-128-cbc-dbg-off}
+
+\addplot[style={YellowGreen,fill=YellowGreen,mark=none}]
+	coordinates {(16,23806.36) (64,27728.96) (256,29181.70) (1024,29573.46) (8192,29687.81)};
+	\label{aes-128-cbc-dbg-on}
+
+\legend{AES-256-CBC debug [off], AES-256-CBC debug [on], AES-128-CBC debug [off], AES-128-CBC debug [on]}
+\end{axis}
+\end{tikzpicture}
+\label{fig:openssl-speed-dbg-on-off}
+\caption{Software benchmark of Openssl speed for AES mode CBC, with 128- and 256-bit keys, debugging flags (de)activated at compilation (\texttt{-fno-inline -g -marm}).}
+\end{figure}
+
 % Talk about the assembly implementation, maybe show the difference in an appendice with the C implementation. Not /that/ interresting.
@@
 \subsection{Strongswan}
+% TODO talk about IKE and SA management
 Strongswan is a full implementation of IPSec relying on the kernel drivers for the networking part, on the crypto API for the cryptographic part, and on user space crypto libraries for the connection negociation.

### 2015-05-18 implementation.tex
diagrams, implem drivers & OS theory

@@
 \chapter{Implementation}
+
+\begin{figure}[ht]
+\center
+\subfloat[\label{fig:os-path-generic-a}Generic]{%
+	\Large
+	\resizebox{.4\linewidth}{!}{%
+	\input{os-path-generic}
+	}
+}
+\subfloat[\label{fig:os-path-generic-b}Specific]{%
+	\Large
+	\resizebox{.4\linewidth}{!}{%
+	\input{os-path-specific}
+	}
+}
+\caption{OS path generic}
+\label{fig:os-path-generic}
+\end{figure}
 
@@
 captionpos=b}
-\begin{lstlisting}[caption=openvpn compress then encrypt -- sample from \texttt{forward.c}, label=list:]
+\begin{lstlisting}[caption=openvpn compress then encrypt -- sample from \texttt{forward.c}, label=list:openvpn-workflow]
 /* Compress, fragment, encrypt and HMAC-sign an outgoing packet. */
@@
 \subsection{OpenSSL}
-% Talk about the debug flags in the compilation and show it has close to none impact on the performance (using openssl speed.
-openssl speed is a relevant test because openssl uses 100% of the CPU, hence maximizing the potential performance reduction.)
 
@@
 
-\subsection{Silex engine}
+\begin{table}[ht]
+\begin{tabular}{|l|c|}\hline
+IP & Ciphers \\ \hline
+BA414E & RSA, DH, DSA, EC \\ 
+BA411E & AES modes CBC, CTR, GCM, CCM, CFB, OFB, CTS, ECB \\ \hline
+\end{tabular}
+\caption{Summary of the ciphers supported by two of Barco Silex' IPs.}
+\label{tab:ip-ciphers}
+\end{table}
 
 \subsection{BA411E Driver}
+Takes the rightern path, which is the cleanest because the most versatile.
+By pluging the driver into the crypto API, we offer a standard kernel interface that can be used by any other kernel driver, such as the ESP driver, or user-space application via the cryptodev driver and OpenSSL engine.
 
-\subsection{BA414E Driver}
+However, at the current state of development, the user-space and the kernel applications do not share the same driver.
+The former is using an IRQ-based driver, whilst the latter is actively polling the hardware.
+The reason why the IRQ-based driver can not be used by kernel-space application is because the current implementation uses sleep methods that, when called on other kernel drivers, put the kernel in panic mode and require the system to reboot.
+An alternative is under development and is further discussed in the last part of this work, in section~\ref{sec:future-work}.
+
+Be it IRQ or polling, the inner workings are the same and follow the exact same canvas as the default software implementation the different AES modes.
+
+\subsection{BA414E Driver and Silex engine}
+At the moment of development, there is no asymetric cryptography interface in the crypto API.
+The BA414E can thus not be accessed using the same path as the BA411E and a user-space driver will be needed, that is the leftern path of figure~\ref{fig:os-path-generic-b}.
+As presented in section~\ref{sec:theory-driver}, we will need to rely on an UIO driver to access the device from the user space.

### 2015-05-18 technical-background.tex
diagrams, implem drivers & OS theory

@@
 \section{Operating system}
+
+%User/kernel space. Networking application: need to send the data through the kernel: huge overhead.
+
+%Crypto API
+
+\subsection{Device Driver}\label{sec:theory-driver}
+% General idea of it works
+% device driver in userspace: linux userspace i/o (/dev/mem)
+Userspace I/O~\cite{koch2011}
+
+There are two ways to get the results from a hardware device: either by using interruption, or by actively polling the device, relentlessly asking it if it finished its operations.
+The first case is the cleanest and the most common: when the device has something to send to the driver, or if anything unexpected happened, it sends an interruption request (IRQ) to the processor, which will in turn execute the interruption routine registered by the driver~\citep[chap.
+10]{Corbet:2005:LDD:1209083}.
+The second case is the easiest and is always guaranteed to work, but won't let go off the processor willingly, loading it at 100\%, and avoiding any other task to be executed.
+Hopefully, modern monolithic kernels such as the Linux kernel from 2.6 provide preemptive scheduling~\cite{Santhanam2003}, that is the scheduler interrupts the running task and assigns the processor ressources it used to an other one.
+Hence, systems with a lot of processes in need for CPU ressources would not be stalled, but it would not change anything if the only process heavily requesting processor time is the one using the driver.
+
+\section{FPGA}
 

### 2015-05-18 test-proto.tex
diagrams, implem drivers & OS theory

\chapter{Test protocol and Prediction}

% Streamline

### 2015-05-19 conclusion.tex
diagrams, implem drivers & OS theory
rfc

@@
 
-\section{Future work}
+\section{Future work}\label{sec:future-work}
 
@@
 This feature would require to work very closely to the linux kernel developpement.
-Indeed, if the signature verification using public key cryptography has been available in the kernel since 2013, a public key encryption API has only been introduced in late April 2015~\cite{cyrpto-api-pk-encryption} and is still under request for comments.
+Indeed, if the signature verification using public key cryptography has been available in the kernel since 2013, a public key encryption API has only been proposed in late April 2015~\cite{crypto-api-pk-encryption} and is still under request for comments.
 

### 2015-05-20 graphs/ftp-bench-comparison.tex
results: graphs

\begin{tikzpicture}

%%%%%%%%%%%%%%%%%%%%%%%%
% CPU in the background
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        width  = 0.95*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar=2pt,%space between the bars
        bar width=16pt,
        enlarge x limits={abs=1},
        ylabel = {CPU\#0 usage},
        hide x axis,
        axis y line*=right,
        ymin=0, ymax=100,
        symbolic x coords={Out-of-tunnel, openSSH, OpenVPN, IPSec},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.15)}, anchor=north,legend columns=4},
        enlarge x limits=0.1,
    ]

\addplot[style={black,fill=LimeGreen,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (Out-of-tunnel, 0)
        (openSSH, 89.29)
        (OpenVPN, 76.03)
        (IPSec, 74.64)
    };
    \label{software}

\addplot[style={black,fill=RedOrange,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (Out-of-tunnel, 0)
        (openSSH, 68.29)
        (OpenVPN, 80.89)
        (IPSec, 17.25)
    };
    \label{ba411e}

\addplot[style={black,fill=gray,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (Out-of-tunnel, 7.16)
        (openSSH, 0)
        (OpenVPN, 0)
        (IPSec, 0)
    };
    \label{out-of-tunnel}%"oot" for "out of tunnel"
\legend{software, ba411e, out-of-tunnel}
\end{axis}

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        title = {Comparison over file transfer methods},
        width  = 0.95*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar=10pt,
        bar width=8pt,
        enlarge x limits={abs=1},
        ymajorgrids = true,
        ylabel = {Throughtput [MB/s]},
        xlabel = {},
        ymin=0, ymax=12,
        symbolic x coords={Out-of-tunnel, openSSH, OpenVPN, IPSec},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.25)}, anchor=north,legend columns=2},
        enlarge x limits=0.1,
    ]

\addplot[style={black,fill=ForestGreen,mark=none}]
    coordinates {
        (Out-of-tunnel, 0)
        (openSSH, 8.19)
        (OpenVPN, 3.87)
        (IPSec, 6.47)
    };
    \label{soft-tp}

\addplot[style={black,fill=BrickRed,mark=none}]
    coordinates {
        (Out-of-tunnel, 0)
        (openSSH, 10.39)
        (OpenVPN, 2.84)
        (IPSec, 5.80)
    };
    \label{ba411e-tp}

\addplot[style={black,fill=black,mark=none}]
    coordinates {
        (Out-of-tunnel, 11.39)
        (openSSH, 0)
        (OpenVPN, 0)
        (IPSec, 0)
    };
    \label{oot-tp}%"tp" for "throughput"
\legend{}
\end{axis}

\end{tikzpicture}

### 2015-05-20 graphs/ipsec-ftp-bench.tex
results: graphs

\begin{tikzpicture}

%%%%%%%%%%%%%%%%%%%%%%%%
% CPU in the background
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        width  = 0.95*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar=2pt,%space between the bars
        bar width=16pt,
        enlarge x limits={abs=1},
        ylabel = {CPU\#0 usage},
        hide x axis,
        axis y line*=right,
        ymin=0, ymax=100,
        symbolic x coords={none:none, aes256cbc:none, aes256cbc:sha256, aes256gcm},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.15)}, anchor=north,legend columns=4},
        enlarge x limits=0.1,
    ]

\addplot[style={black,fill=LimeGreen,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 63.74)
        (aes256cbc:sha256, 74.64)
        (aes256gcm, 89.66)
    };
    \label{software}

\addplot[style={black,fill=RedOrange,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 14.87)
        (aes256cbc:sha256, 17.25)
        (aes256gcm, 0)
    };
    \label{ba411e}

\addplot[style={black,fill=gray,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 7.16)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
        (aes256gcm, 0)
    };
    \label{out-of-tunnel}%"oot" for "out of tunnel"

\addplot[style={black,fill=brown,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 14.68)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
        (aes256gcm, 0)
    };
    \label{inside tunnel}%"it" for "in tunnel"
\legend{software, ba411e, out-of-tunnel, inside tunnel}
\end{axis}

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        title = {FTP transfer insed IPSec tunnel},
        width  = 0.95*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar=10pt,
        bar width=8pt,
        enlarge x limits={abs=1},
        ymajorgrids = true,
        ylabel = {Throughtput [MB/s]},
        xlabel = {},
        ymin=0, ymax=12,
        symbolic x coords={oot, none:none, aes256cbc:none, aes256cbc:sha256, aes256gcm},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.25)}, anchor=north,legend columns=2},
        enlarge x limits=0.1,
    ]

\addplot[style={black,fill=ForestGreen,mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 8.83)
        (aes256cbc:sha256, 6.47)
        (aes256gcm, 5.09)
    };
    \label{soft-tp}

\addplot[style={black,fill=BrickRed,mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 8.52)
        (aes256cbc:sha256, 5.80)
        (aes256gcm, 0)
    };
    \label{ba411e-tp}

\addplot[style={black,fill=black,mark=none}]
    coordinates {
        (none:none, 11.39)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
        (aes256gcm, 0)
    };
    \label{oot-tp}%"tp" for "throughput"

\addplot[style={black,fill=RawSienna,mark=none}]
    coordinates {
        (none:none, 10.21)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
        (aes256gcm, 0)
    };
    \label{it-tp}
\legend{}
\end{axis}

\end{tikzpicture}

### 2015-05-20 graphs/openssh-bench.tex
results: graphs

\begin{tikzpicture}

%%%%%%%%%%%%%%%%%%%%%%%%
% CPU in the background
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        width  = 0.95*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar=2pt,%space between the bars
        bar width=16pt,
        enlarge x limits={abs=1},
        ylabel = {CPU\#0 usage},
        hide x axis,
        axis y line*=right,
        ymin=0, ymax=100,
        symbolic x coords={none:none, aes256cbc:none, aes256cbc:sha256},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.15)}, anchor=north,legend columns=4},
        enlarge x limits=0.1,
    ]

\addplot[style={black,fill=LimeGreen,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 96.26)
        (aes256cbc:sha256, 89.29)
    };
    \label{software}

\addplot[style={black,fill=RedOrange,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 46.01)
        (aes256cbc:sha256, 68.19)
    };
    \label{ba411e}

\addplot[style={black,fill=gray,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (none:none, 7.16)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
    };
    \label{out-of-tunnel}%"oot" for "out of tunnel"
\legend{software, ba411e, out-of-tunnel}
\end{axis}

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        title = {file transfer over SSH tunnel},
        width  = 0.95*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar=10pt,
        bar width=8pt,
        enlarge x limits={abs=1},
        ymajorgrids = true,
        ylabel = {Throughtput [MB/s]},
        xlabel = {},
        ymin=0, ymax=12,
        symbolic x coords={oot, none:none, aes256cbc:none, aes256cbc:sha256},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.25)}, anchor=north,legend columns=2},
        enlarge x limits=0.1,
    ]

\addplot[style={black,fill=ForestGreen,mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 10.89)
        (aes256cbc:sha256, 8.19)
    };
    \label{soft-tp}

\addplot[style={black,fill=BrickRed,mark=none}]
    coordinates {
        (none:none, 0)
        (aes256cbc:none, 10.67)
        (aes256cbc:sha256, 10.39)
    };
    \label{ba411e-tp}

\addplot[style={black,fill=black,mark=none}]
    coordinates {
        (none:none, 11.39)
        (aes256cbc:none, 0)
        (aes256cbc:sha256, 0)
    };
    \label{oot-tp}%"tp" for "throughput"
\legend{}
\end{axis}

\end{tikzpicture}

### 2015-05-20 graphs/openvpn-ftp-bench.tex
results: graphs

--- 2015-05-20 graphs/openssh-bench.tex
@@
         (none:none, 0)
-        (aes256cbc:none, 96.26)
-        (aes256cbc:sha256, 89.29)
+        (aes256cbc:none, 76.60)
+        (aes256cbc:sha256, 76.03)
     };
@@
         (none:none, 0)
-        (aes256cbc:none, 46.01)
-        (aes256cbc:sha256, 68.19)
+        (aes256cbc:none, 83.74)
+        (aes256cbc:sha256, 80.89)
     };
@@
     \label{out-of-tunnel}%"oot" for "out of tunnel"
-\legend{software, ba411e, out-of-tunnel}
+
+\addplot[style={black,fill=brown,postaction={pattern=north east lines},mark=none}]
+    coordinates {
+        (none:none, 42.60)
+        (aes256cbc:none, 0)
+        (aes256cbc:sha256, 0)
+    };
+    \label{inside tunnel}%"it" for "in tunnel"
+\legend{software, ba411e, out-of-tunnel, inside tunnel}
 \end{axis}
@@
 \begin{axis}[
-        title = {file transfer over SSH tunnel},
+        title = {FTP transfer inside OpenVPN tunnel},
         width  = 0.95*\textwidth,
@@
         (none:none, 0)
-        (aes256cbc:none, 10.89)
-        (aes256cbc:sha256, 8.19)
+        (aes256cbc:none, 4.78)
+        (aes256cbc:sha256, 3.87)
     };
@@
         (none:none, 0)
-        (aes256cbc:none, 10.67)
-        (aes256cbc:sha256, 10.39)
+        (aes256cbc:none, 3.35)
+        (aes256cbc:sha256, 2.84)
     };
@@
     \label{oot-tp}%"tp" for "throughput"
+
+\addplot[style={black,fill=RawSienna,mark=none}]
+    coordinates {
+        (none:none, 5.18)
+        (aes256cbc:none, 0)
+        (aes256cbc:sha256, 0)
+    };
+    \label{it-tp}
 \legend{}

### 2015-05-20 results.tex
diffie-hellman
diagrams, implem drivers & OS theory
results: graphs

@@
-\chapter{Results}
+\chapter{Results and Analysis}\label{chap:results}
 
@@
 \section{TLS connections}
+% Begin with the benchmark done by Bastien on the raw number of verif/s with openssl.
+
+This benchmark is done only with OpenVPN.
+Since there is no standard support for those operation in the kernel yet, it would not have made sense to use IPSec, since it would have had to fallback to OpenSSL, then following the same path as OpenVPN do.
+As soon as the public key operations can be plugged into the crypto API, this use case should however be tested immediatly.
 
@@
 %			- capped to software max: show CPU offload
+
+\subsection{openSSH}
+
+\begin{figure}[ht]
+\input{graphs/openssh-bench}
+\caption{file transfer over an SSH tunnel. The background stripped bars are the CPU usage.}
+\label{fig:openssh-bench}
+\end{figure}
+
+\subsection{OpenVPN}
+
+\begin{figure}[ht]
+\input{graphs/openvpn-ftp-bench}
+\caption{FTP file transfer over an OpenVPN tunnel. The background stripped bars are the CPU usage.}
+\label{fig:openvpn-ftp-bench}
+\end{figure}
+
+We can see that adding a MAC computation aside the encryption merely lowers the performance when using the hardware.
+Even though OpenSSL uses here an hihgly ARM-optimzed assembly implementation of SHA-256, it shows that the bottleneck is on the hardware side.
+
+Indeed,
+
+\subsection{IPSec}
+
+\begin{figure}[ht]
+\input{graphs/ipsec-ftp-bench}
+\caption{FTP file transfer over an IPSec tunnel. The background stripped bars are the CPU usage.}
+\label{fig:ipsec-ftp-bench}
+\end{figure}
+
+Some results have to be put into perspective with the fact that the implementation of SHA-256 is entirely C-based.
+A more recent one using assembly instructions optimized for the NEON SIMD instruction set of the ARMv7 core could be used and would most probably yield better results.
+The CPU usage of the software implementation would drop -- even if not significantly -- as for the hardware, it would be less limited by the software MAC counter part, and if the CPU usage could stay at the same level, we could expect a better throughput.
+
+\subsection{Comparison}
+
+\begin{figure}[ht]
+\input{graphs/ftp-bench-comparison}
+\caption{Comparison of file transfer methods. The background stripped bars are the CPU usage.}
+\label{fig:ftp-bench-comparison}
+\end{figure}

### 2015-05-20 technical-background.tex
rfc
results: graphs

@@
 It uses the same kind of operation as RSA, that is modular exponentiation.
-The RFC 2631~\cite{rfc2631} decribes the protocol:
+The protocol can be one of two type (\cite{rfc2631}, ...):
+\begin{itemize}
+	\item Static: the actors use their authenticated certificate to compute the shared secret.
+	\item Ephemeral: the actors create a new pair of public/private keys from which the secret key is derived.
+\end{itemize}
+A static scheme is easier to implement and requires much less operations, but using ephemeral keys is essential to ensure perfect forward secrecy.
+Imagine that somehow, an opponent lays his hand on the shared secret.
+If that secret has already been used, he can decipher all data transfered during past connections.
+However, if the secret is new for every new connection, the compromission of the shared secret des not jeopardize past communications.
+This is perfect forward secrecy: using a new key to protect the past.
 
+Hereunder is the generation of an ephemeral shared secret.
+For a static secret, Alice and Bob will simply use their static certificate, sparing one modular exponentiation.
 \begin{enumerate}
@@
 If the server is Alice, it has to do at least one signature, one signature verification and two modular exponentiations.
+Note that the client, B in our case, could have one signature less because RFC 5246~\cite{rfc5246} leave it as an optional feature, and the server would then have one verification less.
+However, any sane configuration will have both actors signing their ephemeral public key.
 If the certificate use RSA, we end up with four modular exponentiations, which can become quite heavy computing wise for certain sizes of prime numbers.

### 2015-05-20 implementation.tex
FPGA stuff and new caption command.

@@
 }
-\caption{OS path generic}
+\caption{OS path generic}{}
 \label{fig:os-path-generic}
@@
 \includegraphics[width=\textwidth]{OpenVPN-packetflow}
-\caption{OpenVPN packet flow as advertized on the openVPN wiki.}
+\caption{OpenVPN packet flow as advertized on the openVPN wiki.}{}
 \label{fig:openvpn-packet-flow}
@@
 \label{fig:openssl-speed-dbg-on-off}
-\caption{Software benchmark of Openssl speed for AES mode CBC, with 128- and 256-bit keys, debugging flags (de)activated at compilation (\texttt{-fno-inline -g -marm}).}
+\caption{Software benchmark of Openssl speed for AES mode CBC, with 128- and 256-bit keys, debugging flags (de)activated at compilation (\texttt{-fno-inline -g -marm}).}{}
 \end{figure}
@@
 }
-\caption{IPSec user/kernel space workflow, using \texttt{ping} as a test case.}
+\caption{IPSec user/kernel space workflow, using \texttt{ping} as a test case.}{}
 \label{fig:ipsec-workflow}
@@
 \end{tabular}
-\caption{Summary of the ciphers supported by two of Barco Silex' IPs.}
+\caption{Summary of the ciphers supported by two of Barco Silex' IPs.}{}
 \label{tab:ip-ciphers}

### 2015-05-20 technical-background.tex
FPGA stuff and new caption command.

@@
 
-This chapter will address the technical ground inherent to this work: first the operating system, followed by the cryptography and networking.
+This chapter will address the technical ground inherent to this work.
+First comes an overview of Linux operating systems, the distinction between user and kernel mode, and the design of device drivers.
+
+Follow a quick presentation of FPGAs and how they can be driven from the operating system.
 
@@
 \section{FPGA}
+
+Driving from the OS: basically, they will need to share some memory.
+That memory can be directly mapped and accessed from the user-space using \texttt{/dev/mem}, or can use a direct memory access module (DMA).
+From the operation system, we build a bunch of scatterlist in ther kernel-space memory, then map those pages to memory descriptor that have a physical address on the DMA.
+They can be mapped three different ways: \texttt{DMA\_BIDIRECTIONAL}, \texttt{DMA\_TO\_DEVICE} or \texttt{DMA\_FROM\_DEVICE}.
+When the CPU write something in those descriptor and synchronize them with the DMA, it does not have to care about them anymore, the DMA in now in charge to send them to the device where registers are ready to read the incomming data.
+The same goes from the device to the CPU: when the device wants to communicate data to the OS, it writes it on the DMA that will transfer them to the CPU, triggering a flag on the way to notify it.
 
@@
 It uses the same kind of operation as RSA, that is modular exponentiation.
-The protocol can be one of two type (\cite{rfc2631}, ...):
+The protocol can be one of two type (\cite{rfc2631}, \cite{Frankel:2005:SGI:2206289}):
 \begin{itemize}
@@
 	\item Ephemeral: the actors create a new pair of public/private keys from which the secret key is derived.
+	\item Anonymous: same as ephemeral, but without signing anything, hence not identifying neither of the actors. This mode is not advisable since it's vulnerable to main-in-the-middle attack.
 \end{itemize}
@@
 Hereunder is the generation of an ephemeral shared secret.
-For a static secret, Alice and Bob will simply use their static certificate, sparing one modular exponentiation.
+For a static secret, Alice and Bob will simply use their static certificate, sparing the modular exponentiation of the ephemeral public key generation.
 \begin{enumerate}

### 2015-05-22 results.tex
FPGA stuff and new caption command.
ping ipsec

@@
 \section{Response time -- latency}
+
+\subsection{IPSec}
+
+\begin{figure}[ht]
+\input{graphs/ipsec-ping-benchmark}
+\caption{Ping min/avg/max response time}{for different packet sizes using IPsec.
+For each packet size, 1000 requests were flooded to the board, that is \textit{"outputs packets as fast as they come back or one hundred times per second, whichever is more"}, according to the \texttt{ping} command manual.}
+\label{fig:ping-benchmark-ipsec}
+\end{figure}
 
@@
 \input{graphs/openssh-bench}
-\caption{file transfer over an SSH tunnel. The background stripped bars are the CPU usage.}
+\caption{file transfer over an SSH tunnel. The background stripped bars are the CPU usage.}{}
 \label{fig:openssh-bench}
@@
 \input{graphs/openvpn-ftp-bench}
-\caption{FTP file transfer over an OpenVPN tunnel. The background stripped bars are the CPU usage.}
+\caption{FTP file transfer over an OpenVPN tunnel. The background stripped bars are the CPU usage.}{}
 \label{fig:openvpn-ftp-bench}
@@
 \input{graphs/ipsec-ftp-bench}
-\caption{FTP file transfer over an IPSec tunnel. The background stripped bars are the CPU usage.}
+\caption{FTP file transfer over an IPSec tunnel. The background stripped bars are the CPU usage.}{}
 \label{fig:ipsec-ftp-bench}
@@
 
+The GCM performance presented clearly shows a drop of throughput and an increase of CPU usage, illustrating the fact that those operations are hard on the software.
+With an hardware offload, we could expect not only a drastic drop of the CPU usage, but an increase of throughput as well, since it's CPU-limited in those results.
+Note that they are achieved usinf a C-based implementation of galois-field multiplications.
+As we saw in chapter~\ref{chap:theory}, modern processor designers tend to add specialized instruction set aimed at AES-GCM enhancement.
+Should further tests be conducted concerning IPSec paired with GCM, it would be wise to compare with an assembly implementation exploiting ARM NEON instruction set.
+Some are being developed~\cite{Conrado2013,Danilo2013}, but none have been committed to the Linux kernel repository yet.
+
 \subsection{Comparison}
@@
 \input{graphs/ftp-bench-comparison}
-\caption{Comparison of file transfer methods. The background stripped bars are the CPU usage.}
+\caption{Comparison of file transfer methods. The background stripped bars are the CPU usage.}{}
 \label{fig:ftp-bench-comparison}

### 2015-05-22 results.tex
ping bench as line plot

@@
 \chapter{Results and Analysis}\label{chap:results}
-
-\section{Response time -- latency}
-
-\subsection{IPSec}
-
-\begin{figure}[ht]
-\input{graphs/ipsec-ping-benchmark}
-\caption{Ping min/avg/max response time}{for different packet sizes using IPsec.
-For each packet size, 1000 requests were flooded to the board, that is \textit{"outputs packets as fast as they come back or one hundred times per second, whichever is more"}, according to the \texttt{ping} command manual.}
-\label{fig:ping-benchmark-ipsec}
-\end{figure}
 
@@
 As soon as the public key operations can be plugged into the crypto API, this use case should however be tested immediatly.
+
+\section{Response time -- latency}
+After the establishement of a connection.
+
+\subsection{IPSec}
+
+\begin{figure}[ht]
+\input{graphs/ipsec-ping-benchmark}
+\caption{Ping min/avg/max response time}{for different packet sizes using IPsec.
+For each packet size, 1000 requests were flooded to the board, that is \textit{"outputs packets as fast as they come back or one hundred times per second, whichever is more"}, according to the \texttt{ping} command manual.} %TODO move the end of the caption to the "test protocol" chapter.
+\label{fig:ping-benchmark-ipsec}
+\end{figure}
 

### 2015-05-23 graphs/openvpn-tls-bench.tex
raw results, tls benchmark

\begin{tikzpicture}

%%%%%%%%%%%%%%%%%%%%%%%%
% CPU in the background
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        width  = 0.7*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar=2pt,%space between the bars
        bar width=16pt,
        enlarge x limits={abs=1},
        ylabel = {CPU\#0 usage},
        hide x axis,
        axis y line*=right,
        ymin=0, ymax=100,
        symbolic x coords={RSA-1024, RSA-2048, RSA-4096},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.15)}, anchor=north,legend columns=2},
        enlarge x limits=0.1,
    ]

\addplot[style={black,fill=LimeGreen,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (RSA-1024, 40.32)
        (RSA-2048, 92.14)
        (RSA-4096, 81.97)
    };
    \label{software}

\addplot[style={black,fill=RedOrange,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (RSA-1024, 13.29)
        (RSA-2048, 4.82)
        (RSA-4096, 4.78)
    };
    \label{ba414e}
\legend{software, ba414e}
\end{axis}

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        title = {TLS connections through OpenVPN},
        width  = 0.7*\textwidth,
        height = 8cm,
        major x tick style = transparent,
        ybar=10pt,
        bar width=8pt,
        enlarge x limits={abs=1},
        ymajorgrids = true,
        ylabel = {Throughtput [MB/s]},
        xlabel = {},
        ymin=0,
        symbolic x coords={RSA-1024, RSA-2048, RSA-4096},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.25)}, anchor=north,legend columns=2},
        enlarge x limits=0.1,
    ]

\addplot[style={black,fill=ForestGreen,mark=none}]
    coordinates {
        (RSA-1024, 445.4)
        (RSA-2048, 155.6)
        (RSA-4096, 19.6)
    };
    \label{soft-tp}

\addplot[style={black,fill=BrickRed,mark=none}]
    coordinates {
        (RSA-1024, 509.3)
        (RSA-2048, 420.9)
        (RSA-4096, 115.5)
    };
    \label{ba411e-tp}
\legend{}
\end{axis}

\end{tikzpicture}

### 2015-05-23 results.tex
change ping value to avg
ping comparison
raw results, tls benchmark

@@
 
+\begin{figure}[ht]
+\center
+\input{graphs/openvpn-tls-bench}
+\caption{TLS connections per second}{The background stripped bars are the CPU usage. Raw data in table~\ref{tab:tls-con}.}
+\label{fig:openvpn-tls-bench}
+\end{figure}
+
 \section{Response time -- latency}
 After the establishement of a connection.
+
+\subsection{OpenVPN}
+
+\begin{figure}[ht]
+\input{graphs/openvpn-ping-benchmark}
+\caption{Ping min/avg/max response time}{for different packet sizes using OpenVPN}
+\label{fig:ping-benchmark-openvpn}
+\end{figure}
 
@@
 \label{fig:ping-benchmark-ipsec}
+\end{figure}
+
+\subsection{Comparison}
+
+\begin{figure}[ht]
+\input{graphs/ping-bench-comparison}
+\caption{Ping average comparison}{All values are for aes-256-cbc with sha-256}
+\label{fig:ping-benchmark-comparison}
 \end{figure}
@@
 \input{graphs/ftp-bench-comparison}
-\caption{Comparison of file transfer methods. The background stripped bars are the CPU usage.}{}
+\caption{Comparison of file transfer methods}{The background stripped bars are the CPU usage.}
 \label{fig:ftp-bench-comparison}

### 2015-05-23 implementation.tex
implementation introduction

@@
 \chapter{Implementation}
+This chapter will present how we implemented the protocols of chapter~\ref{chap:theory} on the board.
+The figure~\ref{fig:os-path-generic} show the generic data flow in the operating system through the user and kernel spaces.
+We will first detail the software part, with the OpenVPN and openSSH as the application, and OpenSSL as a cryptographic library on which both applications rely.
+Then will come strongswan, a user-space abstraction layer giving access to IPSec and we will see how different it is from an OpenVPN implementation.
+Lastly, the standard Linux cryptographic kernel modules and network drivers will be listed and briefly discussed.
+
+Before closing the chapter, we will present the two main IPs to which the cryptographic operations will be offloaded and the associated drivers.
 
@@
 }
-\caption{OS path generic}{}
+\caption{OS data paths}{(a) as a generic abstraction and (b) with some specific blocks replaced with custom implementations.}
 \label{fig:os-path-generic}
@@
 
+\subsection{OpenSSH}
+% Talk about the MAC=none patch
+% connection using SSL/TLS, then AES, but without fragmentation (unlike openvpn)
+
 \subsection{OpenSSL}
@@
 
-\subsection{OpenSSH}
-% Talk about the MAC=none patch
-% connection using SSL/TLS, then AES, but without fragmentation (unlike openvpn)
-
 \subsection{Strongswan}
@@
 The figure~\ref{fig:ipsec-workflow} illustrates the workflow of Alice communicating with Bob via an IPSec ESP tunnel.
-The XFRM, read "transform", framework is implementing IPSec and handles the incomming and outgoing packets for established VPNs~\cite{rosen2014}.
+The XFRM, read ``transform", framework is implementing IPSec and handles the incomming and outgoing packets for established VPNs~\cite{rosen2014}.
 Its name comes from the fact that the kernel transforms packet frames to incorporate IPSec security.

### 2015-05-24 implementation.tex
patch openssh, implem openssh and drivers.

@@
 \lstset{language=c}
-\lstset{commentstyle=\color{colComments}\textit,
-float=hbp,%
-basicstyle=\ttfamily\footnotesize, %
-identifierstyle=\color{colIdentifier}, %
-keywordstyle=\color{colKeys}, %
-stringstyle=\color{colString}, %
-columns=flexible, %
-tabsize=2, %
-extendedchars=true, %
-showspaces=false, %
-showstringspaces=false, %
-numbers=left, %
-numberstyle=\tiny, %
-breaklines=true, %
-breakautoindent=true, %
-captionpos=b}
 \begin{lstlisting}[caption=openvpn compress then encrypt -- sample from \texttt{forward.c}, label=list:openvpn-workflow]
@@
 % connection using SSL/TLS, then AES, but without fragmentation (unlike openvpn)
+OpenSSH relies on an external cryptographic library for all its security operations.
+Up until the version 6.7 published in october 2014, it had to be compiled against OpenSSL.
+However, after the infamous security vulnerability heartbleed in april 2014, the developpers took a step to move towards LibreSSL, a fork of OpenSSL managed by OpenBSD  developers.
+Still, there is no official support for any other cryptographic library.
+
+If OpenSSH does support most of OpenSSL ciphers by default, it takes some liberties such as disabling the CBC encryption mode and removing the support of no MAC during a transmission.
+
+\noindent The first liberty is a laste\footnote{A vulnerability note as been issued by Carnegie Mellon University Computer Emergency Response Team in early 2009 (last revision)~\cite{CERT2009}, in response to a research of University of London~\cite{Albrecht:2009} presenting a plaintext-recovering attack against SSH when CBC mode is used, but the OpenSSH update took only place in october 2014.}.
+Even if it is not recommended, the user can still enable this mode by explicitly configuring it in the \texttt{sshd\_config} options.
+
+\noindent The second liberty has been taken to prevent the user to strip himself from data authenticity and integrity.
+However, in the case of testing and benchmarking, leaving the MAC behind can be interresting, especially if it is not offloaded in hardware such as the encryption in our case.
+In order to add this feature to OpenSSH, we wrote a patch to apply on OpenSSH 6.7, available in appendix~\cite{chap:openssh-patch}.
 
@@
 It is to be noted that the present work uses an implementation with all the debug flags activated.
-If it does have an impact on the performance, it is minimal: in the worst case of the benchmark, the throughput drops only by 2.4\%.
+Figure~\ref{fig:openssl-speed-dbg-on-off} shows that if it does have an impact on the performance, it is minimal: in the worst case of the benchmark, the throughput drops only by 2.4\%.
 Moreover, the benchmark maximizes this difference by doing only OpenSSL operations.
@@
 \end{tikzpicture}
+\caption{OpenSSL debugging benchmark}{Software benchmark of Openssl speed for AES mode CBC, with 128- and 256-bit keys, debugging flags (de)activated at compilation (\texttt{-fno-inline -g -marm}).
+The throughput difference ranges from 0.2\% and 2.4\% , and is more marked for larger keysize, as more debugging data needs to be generated.}
 \label{fig:openssl-speed-dbg-on-off}
-\caption{Software benchmark of Openssl speed for AES mode CBC, with 128- and 256-bit keys, debugging flags (de)activated at compilation (\texttt{-fno-inline -g -marm}).}{}
 \end{figure}
@@
 % Mention the fact that we use the assembly implementation of AES, not the generic C one.
+% Need an lsmod here.
+Several kernel modules are needed to implement the various cryptographic algorithm in software.
+The GCM alone needs five different modules, and IPSec three others.
+The following description addresses all the kernel modules required to run all the use cases presented in this work.
+
+\begin{description}
+	\item[\texttt{aes\_arm}] Assembly implementation of AES. This version is optimized to use the ARMv7 instruction set.
+	\item[\texttt{sha256-generic}] C implementation of SHA-256.
+	\item[\texttt{gfmult128}] Multiplication in $GF(2^128)$, needed by the GCM mode.
+	\item[\texttt{ghash}] GHASH function needed by the authentication of the GCM mode.
+	\item[\texttt{seqiv}] Sequence IV number generator, needed by the CTR and GCM modes.
+	\item[\texttt{cbc}] C based CBC mode.
+	\item[\texttt{gcm}] C base GCM mode.
+	\item[\texttt{crypto\_null}] Null cipher. This basically does nothing on the plaintext, but it is still needed to propose the interface in the kernel.
+	\item[\texttt{tun}] TUN/TAP network device, needed by OpenVPN.
+	\item[\texttt{xfrm\_user}] XFRM operations.
+	\item[\texttt{esp4}] IPv4 ESP implementation. The AH counterparty is also available in the \texttt{ah4} module.
+	\item[\texttt{ipcomp}] IP compression module. Needed by IPSec if the option is activated in the configuration.
+	\item[\texttt{cryptodev}] Creates \texttt{/dev/crypto}, giving access to the crypto API in the user space.
+	\item[\texttt{uio}] User-space I/O driver, allowing to access the hardware memory from the user space, needed by our implementation of the BA414E driver.
+\end{description}
 
 \section{Offload}
+
+The table~\ref{tab:ip-ciphers} summarizes the ciphers supported by the two Barco Silex' IPs used in this work.
 
@@
 \subsection{BA414E Driver and Silex engine}
-At the moment of development, there is no asymetric cryptography interface in the crypto API.
+At the moment of development, there is no asymetric cryptography interface in the crypto API\footnote{A request for comment patch as been submitted to the Linux kernel mailing list~\cite{crypto-api-pk-encryption} in late April 2015, proposing a standard interface for public key encryption in the crypto API.}.
 The BA414E can thus not be accessed using the same path as the BA411E and a user-space driver will be needed, that is the leftern path of figure~\ref{fig:os-path-generic-b}.

### 2015-05-24 test-proto.tex
results: graphs
Move presentation to test protocol.

@@
 
-% Streamline
+\section{Experimental setup}
+The experimental environment is built around a standard x86 host and an ARM Cortex-A9 alongside an Altera Cyclone V FPGA as the target.
+Both are linked toghether through a network capped by 100Mbps switches.
+Both stations have gigabits ethernet interface and could hence be diretly connected to each other, but in that case the communication would be limited by the I/O transfers of the storage units -- a hard drive disk in one case, an micro-SD card in the second -- on which we can not depend to set a constant throughput limitation, as it is highly influenced by the data block size and general health of the support.
+
+\subsection{x86 host}
+The desktop host runs on Windows 7 Professional 64-bit, but a virtual machine using a Linux distribution is used for the developpement and testings.
+
+\begin{framed}
+\begin{description}
+	\item[OS] Ubuntu 12.04 LTS, kernel 3.16
+	\item[CPU] Intel Core-i3 ... (two logical core out of four)
+	\item[RAM] 1GB DDR3
+\end{description}
+\end{framed}
+
+\subsection{Altera Socrates SoCFPGA}
+
+\begin{framed}
+\begin{description}
+	\item[OS] Yocto project, kernel 3.14
+	\item[CPU] Dual core ARM Cortex-A9, 800MHz
+	\item[RAM] ...GB DDR3
+	\item[FPGA] Altera Cyclone V
+\end{description}
+\end{framed}
+
+\subsection{ARM DS-5 Streamline}
+% Use a daemon and a driver to get system info and send them to another machine. Has almost no impact on the performance and the load of the system.
+
+\section{File transfer}
+The file transfered is an un compressed block of 128MB of random data generated using the following command:
+\begin{lstlisting}[language=bash]
+  $ head -c $((1024*1024*128)) /dev/urandom > heavy.file
+\end{lstlisting}

### 2015-05-24 introduction.tex
openvpn stuff etc
One phrase. Yep.

@@
 \chapter{Introduction}
+% What is hardware offloading?
+
+Altera and Xilinx are trying to push those SoC plateforms to the market to popularize the FPGAs.
+
+Power is critical nowadays, esapcially in datacenters.
+Several solutions exists: abandonning general purpose processors altoghether and turn towards ARM cores, or add an FPGA to the processor for example.
+Intel proposed the latter solution in 2014~\cite{intel-xeon-fpga} and is on its way to expand that market with the merger with Altera~\cite{intel-altera-merger}.
 

### 2015-05-24 implementation.tex
Bunch of steps.

@@
 % By default, point-to-point, but since the version 2.0 can be configured in server mode, hence being able to manage hundreds of clients.
+% TODO talk about the cipher none patch for version < ...
+
+OpenVPN is widely used, and is can even integraed in some routers using packages such as DD-WRT and OpenWRT.
+OpenVPN is also used by the Dutch government to secure all its communications~\cite{openvpn-nl}.
+
+OpenVPN offers two different TUN/TAP virtual interfaces:
+\begin{description}
+	\item[TUN] Layer 3 IP tunnel
+	\item[TAP] Layer 2 ethernet tunnel.
+\end{description}
+
+Uses regular TCP/UDP network protocols, which can be an advantage over IPSec if the ISP decides to block the latter.
+
+Note that if one wants to use the \texttt{none} cipher with a version of OpenVPN ealier than 2.6.7, he would have to update it using a community patch~\cite{openvpn-patch-none}.
 
@@
 \subsection{OpenSSL}
+
+Offers an high-level interface called EVP to be used by other applications.
 

### 2015-05-24 technical-background.tex
ipsec transport/tunnel overhead figure.
implementation introduction
patch openssh, implem openssh and drivers.
Bunch of steps.

@@
-\chapter{Technical background}
+\chapter{Technical background}\label{chap:theory}
 
@@
 % Present the TCP/IP layering
+The RFC 1122~\cite{rfc1122} defines the TCP/IP and OSI stack as in the table~\ref{tab:tcp-ip-stack}.
+The main difference between the two is the application layer of the TCP/IP stack which corresponds to the three upper layers of the OSI model.
+Some references, such as~\citet{tanenbaum2011}, conceptually split the TCP/IP link layer into an additional physical layer.
+
+\begin{table}[ht]
+\center
+\begin{tabularx}{\textwidth}{|l|l|l|X|} \hline
+\multicolumn{2}{|c|}{TCP/IP layering} & \multicolumn{2}{c|}{OSI model} \\ \hline
+Layer & Protocols & Layer & Protocols \\ \hline
+\multirow{3}{*}{Application} & \multirow{3}{*}{FTP, SSH} & Application & FTP \\ \cline{3-4}
+ & & Presentation & ASCII, JPEG \\ \cline{3-4}
+ & & Session & RPC, PAP \\ \hline
+Transport & TCP, UDP & Transport & TCP, UDP \\ \hline
+Internet & IP, ICMP, IPSec & Network & IP, ICMP, IPSec \\ \hline
+\multirow{2}{*}{Data Link} & \multirow{2}{*}{PPP, MAC, Ethernet, L2TP} & Data link & PPP, MAC, Ethernet, L2TP \\ \cline{3-4}
+ & & Physical & USB, DSL, IEEE 802.11 \\ \hline
+\end{tabularx}
+\caption{TCP/IP and OSI model comparison}{They are globaly the same, except for the application layer of the TCP/IP stack which merge togheter the three upper layers of the OSI model.
+Between parenthsis are examples of protocols resting on each layer.}
+\label{tab:tcp-ip-stack}
+\end{table}
 
@@
 
-There exist several major implementations of VPN: SSL, IPSec and PPTP.
+There exist several major implementations of VPN: SSL, IPSec, L2TP and PPTP.
 The later was developped by a vendor consortium and proposed in the RFC 2637 and will not be discussed further.
@@
 % Question the security? Apparently, it does mac-then-encrypt, which is insecure regarding certain types of attacks \cite{cryptoeprint:2001:045}.
-% Plus, SSLv3 is to be deprecated, according to a queued RFC: http://www.rfc-editor.org/internet-drafts/draft-ietf-tls-sslv3-diediedie-03.txt
+% Plus, SSLv3 is to be deprecated, according to a queued RFC: http://www.rfc-editor.org/internet-drafts/draft-ietf-tls-sslv3-diediedie-03.txt, and is not supported anymore by many servers since poodle.
 
@@
 % IPSec as a protocol, strongswan come in the 'implementation' chapter.
+
+%TODO Figure IPSec frame structure.
+
+\begin{figure}[ht]
+\center
+\subfloat[\label{fig:ipsec-transport}Transport]{%
+	\large
+	\resizebox{.465\linewidth}{!}{%
+	\input{ipsec-transport}
+	}
+}
+\subfloat[\label{fig:ipsec-tunnel}Tunnel]{%
+	\large
+	\resizebox{.535\linewidth}{!}{%
+	\input{ipsec-tunnel}
+	}
+}
+\caption{IPSec transport (a) and tunnel (b) overheads}{Tunnel mode adds a custom IP header and moves the AH/ESP header in front of the original IP header.
+Note: ``trl" stands for ``trailer", ``hdr" for ``header"}
+\label{fig:ipsec-transport-tunnel}
+\end{figure}

### 2015-05-25 results.tex
Bunch of steps.
test protocol, openvpn TLS analysis

@@
 
-This benchmark is done only with OpenVPN.
-Since there is no standard support for those operation in the kernel yet, it would not have made sense to use IPSec, since it would have had to fallback to OpenSSL, then following the same path as OpenVPN do.
-As soon as the public key operations can be plugged into the crypto API, this use case should however be tested immediatly.
+If the ten clients could connect instantaneously to the server every second, the maximum number of connections would be 600 per minute.
+However, a certain connection time has to be taken into account.
+Those are summarized in table~\ref{tab:openvpn-con-time}.
+
+\begin{table}[ht]
+\center
+\small
+\begin{tabular}{ll|l|} \cline{3-3}
+ & & Connection time [s] \\ \hline
+\multicolumn{1}{|l|}{\multirow{2}{*}{RSA-1024}} & soft & 0.041921 \\ \cline{2-3}
+\multicolumn{1}{|l|}{} & BA411E & 0.020312 \\ \hline
+\multicolumn{1}{|l|}{\multirow{2}{*}{RSA-2048}} & soft & 0.202945 \\ \cline{2-3}
+\multicolumn{1}{|l|}{} & BA411E & 0.039965 \\ \hline
+\multicolumn{1}{|l|}{\multirow{2}{*}{RSA-4096}} & soft & 1.436743 \\ \cline{2-3}
+\multicolumn{1}{|l|}{} & BA411E & 0.183533 \\ \hline
+\end{tabular}
+\caption{OpenVPN connection time}{time necessary to establish an aes-256-cbc connection with DHE.}
+\label{tab:openvpn-con-time}
+\end{table}
+
+It already shows that the connection latency is divided by 2 for low security RSA, and up to by 7.8 for higher security parameters.
+The figure~\ref{fig:openvpn-tls-bench} shows the number of TLS connections per minute for three RSA exponent sizes: 1024-, 2048- and 4096-bit.
+The higher the exponent size, the higher the performance boost.
+
+\begin{table}[ht]
+\center
+\small
+\begin{tabular}{l|c|c|c|c|c|c|c|c|c|} \cline{2-10}
+ & \multicolumn{3}{c|}{RSA-1024} & \multicolumn{3}{c|}{RSA-2048} & \multicolumn{3}{c|}{RSA-4096} \\ \cline{2-10}
+ & \multicolumn{2}{c|}{Con.} & CPU & \multicolumn{2}{c|}{Con.} & CPU & \multicolumn{2}{|c|}{Con.} & CPU \\ \hline
+\multicolumn{1}{|c|}{Soft} & 445.4 & \multirow{2}{*}{x1.14} & 40.32 & 155.6& \multirow{2}{*}{x2.70}  & 92.14 & 19.6& \multirow{2}{*}{x5.89}  & 81.97 \\ \cline{1-2}\cline{4-5}\cline{7-8}\cline{10-10}
+\multicolumn{1}{|c|}{BA414E} & 509.3 & & 13.29 & 420.9 & & 4.82 & 115.5 & & 4.34 \\ \hline
+\end{tabular}
+\caption{TLS connections per minute}{measures obtained with ten clients concurently connecting to an OpenVPN server.}
+\label{tab:tls-con}
+\end{table}
 
@@
 \input{graphs/openvpn-tls-bench}
-\caption{TLS connections per second}{The background stripped bars are the CPU usage. Raw data in table~\ref{tab:tls-con}.}
+\caption{TLS connections per minute}{The background stripped bars are the CPU usage. Raw data in table~\ref{tab:tls-con}.}
 \label{fig:openvpn-tls-bench}
 \end{figure}
+
+\noindent For RSA-1024, the results are mitigated: a poor performance increase, but already less than half the CPU usage.
+It should be noted that at this point, the number of clients is probably too low to push the configuration to its limits.
+It is however an interesting comparison case with the next level of security: RSA-2048.
+
+\noindent RSA-2048 is a much more common configuration, espacially since the NIST deprecated RSA-1024 in 2013.
+The full software implementation is visibly affected by the increase of the exponent size: the CPU usage doubles and the server processes three time less connections.
+At the same time, the hardware loses less than 20\% connections for a third of the CPU usage.
+%TODO check if it is hardware limited or limited by the performances of the VM.
+It is not to be forgotten that it has to keep up with the server, even when the latter is helped by the high-performance hardware.
+My guess is that the VM is capping the perf, hence the exact same CPU usage for RSA-2048 and RSA-4096.
+
+\noindent The results obtained for RSA-4096 can be interpreted similarly to those of RSA-2048, except that the CPU usage is exactly the same for th hardware configuration.
+One way too look at those results is to directly compare the raw performance, and the hardware can then process almost six times more connections per minute than the software.
+However, this is only half of it, since it leaves the CPU usage drop aside.
+If we look at the efficiency, the software can process 0.24 connection per percentage of CPU usage, whilst the hardware can process 24 of them.
+The efficiency is thus multipled by a factor 1000.
+
+Such interesting results, particularly regarding the CPU usage, are possible because at least 87\% of the operations are RSA and Diffie-Hellman operations, which are entirely offloaded in hardware.
+Nevertheless, OpenVPN still needs to proceed to some extra computations (such as SHA-1 integrity), and the hardware operations are not instantaneous, so the performance gain can only be that high.
 

### 2015-05-25 test-proto.tex
test protocol, openvpn TLS analysis

@@
 
+\section{TLS Connections}
+This benchmark is done only with OpenVPN.
+Since there is no standard support for those operation in the kernel yet, it would not have made sense to use IPSec, since it would have had to fallback to OpenSSL, then following the same path as OpenVPN do.
+As soon as the public key operations can be plugged into the crypto API, this use case should however be tested.
+
+For this use case, the board is configured in server mode, so that it can accept connections from any client.
+The virtual machine will then execute then clients in parallel using the script~\ref{list:openvpn-client-script}.
+The only option differentiating the clients is their IP address and port number.
+Otherwise, all the clients share the same basic configuration file (see listing~\ref{list:openvpn-config-client}), which tell them to renogociate a new connection every second.
+Hence, if a connection could be made with no delay and if the processes scheduling were ideal, the server would have to address 600 connections per minute.
+
+\lstinputlisting[language=bash]{stress-openvpn.sh}
+
 \section{File transfer}

### 2015-05-25 stress-openvpn.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+
+timeout=120
+rsa=1024
+remote_ip=150.158.232.208
+
+for i in `seq 1 10`; do
+	timeout ${timeout} openvpn --config client_${i}_${rsa}.ovpn --verb 2 --remote ${remote_ip} &
+done

### 2015-05-25 article/graphs/ftp-bench-comparison.tex
Article draft, uses IEEE template.

\begin{tikzpicture}

%%%%%%%%%%%%%%%%%%%%%%%%
% CPU in the background
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        width  = 0.4*\textwidth,
        height = 5cm,
        major x tick style = transparent,
        ybar=2pt,%space between the bars
        bar width=10pt,
        enlarge x limits={abs=1},
        ylabel = {CPU\#0 usage},
        hide x axis,
        axis y line*=right,
        ymin=0, ymax=100,
        symbolic x coords={oot, openSSH, OpenVPN, IPSec},
        xtick = data,
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.45)}, anchor=north,legend columns=4},
        enlarge x limits=0.1,
    ]

\addplot[style={black,fill=LimeGreen,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (oot, 0)
        (openSSH, 89.29)
        (OpenVPN, 76.03)
        (IPSec, 74.64)
    };
    \label{software}

\addplot[style={black,fill=RedOrange,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (oot, 0)
        (openSSH, 68.29)
        (OpenVPN, 80.89)
        (IPSec, 17.25)
    };
    \label{hardware}

\addplot[style={black,fill=gray,postaction={pattern=north east lines},mark=none}]
    coordinates {
        (oot, 7.16)
        (openSSH, 0)
        (OpenVPN, 0)
        (IPSec, 0)
    };
    \label{oot}%"oot" for "out of tunnel"
\legend{software, hardware, oot}
\end{axis}

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        % title = {Comparison over file transfer methods},
        width  = 0.4*\textwidth,
        height = 5cm,
        major x tick style = transparent,
        ybar=6pt,
        bar width=6pt,
        enlarge x limits={abs=1},
        ymajorgrids = true,
        ylabel = {Throughtput [MB/s]},
        xlabel = {},
        ymin=0, ymax=12,
        symbolic x coords={oot, openSSH, OpenVPN, IPSec},
        xtick = data,
        x tick label style={rotate=45,anchor=east},
        scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        % legend style={at={(0.5,-0.25)}, anchor=north,legend columns=2},
        enlarge x limits=0.1,
    ]

\addplot[style={black,fill=ForestGreen,mark=none}]
    coordinates {
        (oot, 0)
        (openSSH, 8.19)
        (OpenVPN, 3.87)
        (IPSec, 6.47)
    };
    \label{soft-tp}

\addplot[style={black,fill=BrickRed,mark=none}]
    coordinates {
        (oot, 0)
        (openSSH, 10.39)
        (OpenVPN, 2.84)
        (IPSec, 5.80)
    };
    \label{hardware-tp}

\addplot[style={black,fill=black,mark=none}]
    coordinates {
        (oot, 11.39)
        (openSSH, 0)
        (OpenVPN, 0)
        (IPSec, 0)
    };
    \label{oot-tp}%"tp" for "throughput"
\legend{}
\end{axis}

\end{tikzpicture}

### 2015-05-25 article/graphs/openvpn-tls-bench.tex
Article draft, uses IEEE template.

--- 2015-05-23 graphs/openvpn-tls-bench.tex
@@
 \begin{axis}[
-        width  = 0.7*\textwidth,
-        height = 8cm,
+        width  = 0.4*\textwidth,
+        height = 5cm,
         major x tick style = transparent,
         ybar=2pt,%space between the bars
-        bar width=16pt,
+        bar width=10pt,
         enlarge x limits={abs=1},
@@
         scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
-        legend style={at={(0.5,-0.15)}, anchor=north,legend columns=2},
-        enlarge x limits=0.1,
+        legend style={at={(0.5,-0.20)}, anchor=north,legend columns=2},
+        enlarge x limits=0.4,
     ]
@@
     };
-    \label{ba414e}
-\legend{software, ba414e}
+    \label{hardware}
+\legend{software, hardware}
 \end{axis}
@@
         title = {TLS connections through OpenVPN},
-        width  = 0.7*\textwidth,
-        height = 8cm,
+        width  = 0.4*\textwidth,
+        height = 5cm,
         major x tick style = transparent,
-        ybar=10pt,
-        bar width=8pt,
+        ybar=6pt,
+        bar width=6pt,
         enlarge x limits={abs=1},
@@
         scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
-        legend style={at={(0.5,-0.25)}, anchor=north,legend columns=2},
-        enlarge x limits=0.1,
+        % legend style={at={(0.5,-0.25)}, anchor=north,legend columns=2},
+        enlarge x limits=0.4,
     ]
@@
     };
-    \label{ba411e-tp}
+    \label{hardware-tp}
 \legend{}

### 2015-05-25 implementation.tex
Theory and shit

--- 2015-05-24 implementation.tex
@@
 % By default, point-to-point, but since the version 2.0 can be configured in server mode, hence being able to manage hundreds of clients.
+% https://openvpn.net/index.php/open-source/339-why-ssl-vpn.html
+% https://openvpn.net/index.php/open-source/335-why-openvpn.html
+% https://openvpn.net/papers/BLUG-talk/index.html
+
+OpenVPN is also used by the Dutch government to secure all its communications~\cite{openvpn-nl}.
+
+Uses regular TCP/UDP network protocols, which can be an advantage over IPSec if the ISP decides to block the latter.
+
+OpenVPN offers two different TUN/TAP virtual interfaces:
+\begin{description}
+	\item[TUN] Virtual layer 3 IP tunnel, seen as a point-to-point interface from the kernel point of view.
+	\item[TAP] Virtual layer 2 ethernet tunnel.
+\end{description}
+
+Since OpenVPN aims at network security at the application level, it needs virtual network interface to which to push and from which to read the secured data.
+The \texttt{tun} adapter will be used when OpenVPN wants to establish a point-to-point connection between two client, while the \texttt{tap} device is used when there is one server acting as a bridge, hence managing lots of clients at the same time.
+
+OpenVPN as two choices when it wants to sends its data over the internet: either encapsulating its IP packet into TCP or UDP packets.
+The problem of encapsulating IP into TCP is that inside IP packets are already TCP frames.
+The reason is that IP has been designed to work on unreliable networks, and TCP includes in its standard protocols to retry and drop packets.
+Hence, encapsulating an IP packet into a TCP packet produces a redundancy by nesting reliability layers.
+UDP is the unreliable counterparty of TCP, offering a better alternative for encapsulation.
+If we refer to the figure~\ref{fig:openvpn-packet-flow}, \texttt{iperf} would send its regular \texttt{IP(TCP)} packet over the virtual \texttt{tun} device, which will then compress, fragment and encrypt the frame before encaspulating it into a new UDP packet that will finally be sent through the physical \texttt{eth0} interface.
+
+\begin{figure}[ht]
+\includegraphics[width=\textwidth]{OpenVPN-packetflow}
+\caption{OpenVPN packet flow}{it shows two hosts using \texttt{iperf} inside an OpenVPN tunnel.
+This figure comes from the OpenVPN documentation~\cite{openvpn-doc-workflow} and has inverted the encryption and fragmentation.}
+\label{fig:openvpn-packet-flow}
+\end{figure}
+
+\noindent At this point, it is important to note that if the figure~\ref{fig:openvpn-packet-flow} shows a workflow in which the encryption taking place before the fragmentation, a quick look at the source code proves the opposite.
+A sample of the said code has been writen down in the listing~\ref{list:openvpn-workflow}.
 
 \lstset{language=c}
-\begin{lstlisting}[caption=openvpn compress then encrypt -- sample from \texttt{forward.c}, label=list:openvpn-workflow]
+\begin{lstlisting}[caption={openvpn compress then encrypt -- sample from \texttt{forward.c}.
+It clearly shows that the order of operations in the packet workflow is compression, then fragmentation and finally encryption.}, label=list:openvpn-workflow, float]% The 'float' option makes the listing unbreakable.
 /* Compress, fragment, encrypt and HMAC-sign an outgoing packet. */
@@
 
-\begin{figure}[ht]
-\includegraphics[width=\textwidth]{OpenVPN-packetflow}
-\caption{OpenVPN packet flow as advertized on the openVPN wiki.}{}
-\label{fig:openvpn-packet-flow}
-\end{figure}
+An other way to look at this workflow is to completely isolate the OpenVPN part from both hosts, and exporting it to an independent device such as a router.
+The hosts would be stripped off the virtual interface -- \texttt{tun0} would become physical, and would communicate normally over the internet, at least from their point of view.
+All their communications would be routed by the router of their local area network, incomming on a physical \texttt{tun0} interface and outputed to the internet on the \texttt{eth0} interface.
+Such implementation are used when an adminstrator wishes to secure a whole local area network without bloating each client with a VPN application.
+In this case, there exist router firmware packages running OpenVPN, such as DD-WRT and OpenWRT.
+Whatever the method used, the crucial point is that the VPN tunnel usage is transparent to the application.\newline{}
+
+OpenVPN relies on a cryptographic library, OpenSSL in our case, for all the security operations: encryption, signature, certificate management, etc.
+
+\noindent Note that if one wants to use the \texttt{none} cipher with a version of OpenVPN ealier than 2.3.7, he would have to update it using a community patch~\cite{openvpn-patch-none}.
 
@@
 However, in the case of testing and benchmarking, leaving the MAC behind can be interresting, especially if it is not offloaded in hardware such as the encryption in our case.
-In order to add this feature to OpenSSH, we wrote a patch to apply on OpenSSH 6.7, available in appendix~\cite{chap:openssh-patch}.
+In order to add this feature to OpenSSH, we wrote a patch to apply on OpenSSH 6.7, available in appendix~\ref{chap:openssh-patch}.
 
 \subsection{OpenSSL}
+
+Offers an high-level interface called EVP to be used by other applications.
 

### 2015-05-25 technical-background.tex
Theory and shit

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Operating system}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{FPGA}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Cryptography}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Cryptography is the corner stone of security.
+The four main goals are the following, as defined in~\cite{Menezes1996}:
+\begin{description}
+	\item[Confidentiality] keeping information secret from all but those who are authorized to see it.
+	\item[Integrity] ensuring information has not been altered by unauthorized or unknown means.
+	\item[Source Authentication] corroborating the source of information.
+	\item[Non-repudiation] preventing the denial of previous commitments or actions.
+\end{description}
+
+In order to achieve those, four cryptographic primitives are needed: symmetric and asymmetric ciphers, message digests and digital signatures.
 
@@
 
-\subsubsection{SHA}
-Keyed signature algorithm, several versions in place. SHA-1 is depreciated, SHA-2 is widely used and SHA-3 is already defined and begins to be implemented.
+\begin{figure}
+\includegraphics[width=\textwidth]{nist-cbc}
+\caption{CBC encryption and decryption diagram}{taken from the NIST recommendation~\cite{nist-sp800-38A}.}
+\label{fig:cbc-encrypt-decrypt}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=\textwidth]{nist-gcm-encrypt}
+\caption{GCM encryption diagram}{taken from the NIST specification~\cite{mcgrew2005}.
+The ciphertext blocks are formed by \textit{xor}-ing the encrypted counter and the plaintext.
+The tag is generated by a chain of ciphertext \textit{xor}-ing with Galois field multiplicated data.
+The decryption works excatly the same way, except the plaintext and ciphertext are swapped.}
+\label{fig:gcm-encrypt}
+\end{figure}
+
+\subsection{Message digest}
+%Not the same as digital signature because we use the same key for both MAC values.
+A message digest is the result of a one-way mathematical function of a fixed size.
+Those hash functions are of two types~\cite{infof405}: manipulation detection codes (MDC) to guarantee integrity and message authentication codes (MAC) to guarantee both integrity and source authentication.
+
+An MDC $h(x)$ can follow an iterative construction for a message $x$ including $t$ blocks:
+\[
+\begin{dcases}
+	H_0 = \mbox{initial value}\\
+	H_i = f(H_{i-1}, x_i), \mbox{with } i \in [1,t]\\
+	h(x) = H_t\\
+\end{dcases}
+\]
+
+Based on this design and adding a key to the process, the RFC 2104~\cite{rfc2104} defines a MAC:
+\[
+	HMAC(k, x) = h((k\oplus opad)|h((k\oplus ipad)|x))
+\]
+with a key $k$, and two padding block added for security concerns: an outer pad $opad$ and an inner pad $ipad$.
+
+\noindent There exist a wide varety of MDCs, ranging from block cipher based such as Miyaguchi-Preneel, customized such as MD5, SHA-1 and SHA-2, or built using modular arithmetic such as MASH-1.
+
+In both schemes, data integrity can be guaranteed because the flip of one bit will irremediably change the digest.
+However, only a MAC can ensure source authentication since it is the only one based on a shared secret key.
+
+Now rises the question of what and when authenticating.
+\citet{Bellare2000} prooved that the most secure solution is to encrypt then compute the MAC from the ciphertext.
+We will see in section~\ref{sec:theory-network} that if IPSec follows this recommandation, SSL/TLS does not and MAC first the plaintext then encrypt the message.
 
@@
 
-\section{Network and VPN implementation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Network and VPN implementation}\label{sec:theory-network}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 There exist several major implementations of VPN: SSL, IPSec, L2TP and PPTP.
-The later was developped by a vendor consortium and proposed in the RFC 2637 and will not be discussed further.
+The later was developped by a vendor consortium leaded by Microsoft and proposed in the RFC 2637 and will not be discussed further.
 
@@
 % Plus, SSLv3 is to be deprecated, according to a queued RFC: http://www.rfc-editor.org/internet-drafts/draft-ietf-tls-sslv3-diediedie-03.txt, and is not supported anymore by many servers since poodle.
+Application level security.
 
@@
 % IPSec as a protocol, strongswan come in the 'implementation' chapter.
+Modification of the IP stack in the kernel space.
+IPSec is a network level security; it examines incomming IP packets and checks if there exists a security association with the destination, and decrypt it on-the-fly if necessary.
+
+One of the main disadvantage compared to a user-space VPN is the difficulty to traverse NAT.
 

### 2015-05-25 conclusion.tex
Bunch of stuff. AAAAAAAAAAAAAAAH

@@
 
-As we already remarked in \ref{sec:section-where-we-see-the-CPU-usage-of-ba411e-usage-with-active-polling}, the use cases involving IPSec were conducted using a previous revision of the driver still activelly polling the IP for its results.
+As we already remarked in \ref{sec:section-where-we-see-the-CPU-usage-of-ba411e-usage-with-active-polling}, the use cases involving IPsec were conducted using a previous revision of the driver still activelly polling the IP for its results.
 A better and cleaner way to proceed is to use interruption routines, as shown in \ref{sec:section-ba411e-with-irq}.

### 2015-05-25 implementation.tex
Bunch of stuff. AAAAAAAAAAAAAAAH

@@
 We will first detail the software part, with the OpenVPN and openSSH as the application, and OpenSSL as a cryptographic library on which both applications rely.
-Then will come strongswan, a user-space abstraction layer giving access to IPSec and we will see how different it is from an OpenVPN implementation.
+Then will come strongswan, a user-space abstraction layer giving access to IPsec and we will see how different it is from an OpenVPN implementation.
 Lastly, the standard Linux cryptographic kernel modules and network drivers will be listed and briefly discussed.
@@
 
-Uses regular TCP/UDP network protocols, which can be an advantage over IPSec if the ISP decides to block the latter.
+Uses regular TCP/UDP network protocols, which can be an advantage over IPsec if the ISP decides to block the latter.
 
@@
 % TODO talk about IKE and SA management
-Strongswan is a full implementation of IPSec relying on the kernel drivers for the networking part, on the crypto API for the cryptographic part, and on user space crypto libraries for the connection negociation.
+Strongswan is a full implementation of IPsec relying on the kernel drivers for the networking part, on the crypto API for the cryptographic part, and on user space crypto libraries for the connection negociation.
 An other popular implementation is ipsec-tools, but its development lags behind modern Linux and is not up-to-date with the 3.14 Linux kernel headers, making its cross-compilation painfull.
@@
 
-The figure~\ref{fig:ipsec-workflow} illustrates the workflow of Alice communicating with Bob via an IPSec ESP tunnel.
-The XFRM, read ``transform", framework is implementing IPSec and handles the incomming and outgoing packets for established VPNs~\cite{rosen2014}.
-Its name comes from the fact that the kernel transforms packet frames to incorporate IPSec security.
+The figure~\ref{fig:ipsec-workflow} illustrates the workflow of Alice communicating with Bob via an IPsec ESP tunnel.
+The XFRM, read ``transform", framework is implementing IPsec and handles the incomming and outgoing packets for established VPNs~\cite{rosen2014}.
+Its name comes from the fact that the kernel transforms packet frames to incorporate IPsec security.
 Depending on the configuration, XFRM uses the AH or ESP kernel module, which in turn calls the crypto API to encrypt and/or sign the IP packet.
 
-We can also clearly see one of the main advantages of IPSec: it works in the kernel space.
+We can also clearly see one of the main advantages of IPsec: it works in the kernel space.
 Since it does not require a virtual network interface like OpenVPN, the only transfer between the user/kernel space happens when the former wishes to send a packet on the network, passing it to the later -- or \textit{vice versa} for incomming packets.
@@
 }
-\caption{IPSec user/kernel space workflow, using \texttt{ping} as a test case.}{}
+\caption{IPsec user/kernel space workflow, using \texttt{ping} as a test case.}{}
 \label{fig:ipsec-workflow}
@@
 Several kernel modules are needed to implement the various cryptographic algorithm in software.
-The GCM alone needs five different modules, and IPSec three others.
+The GCM alone needs five different modules, and IPsec three others.
 The following description addresses all the kernel modules required to run all the use cases presented in this work.
@@
 	\item[\texttt{esp4}] IPv4 ESP implementation. The AH counterparty is also available in the \texttt{ah4} module.
-	\item[\texttt{ipcomp}] IP compression module. Needed by IPSec if the option is activated in the configuration.
+	\item[\texttt{ipcomp}] IP compression module. Needed by IPsec if the option is activated in the configuration.
 	\item[\texttt{cryptodev}] Creates \texttt{/dev/crypto}, giving access to the crypto API in the user space.

### 2015-05-25 main.tex
stuff
conclusion
daily update
openssl speed dbg on/off
openvpn stuff etc
diffie-hellman
diagrams, implem drivers & OS theory
FPGA stuff and new caption command.
raw results, tls benchmark
patch openssh, implem openssh and drivers.
Move presentation to test protocol.
Bunch of stuff. AAAAAAAAAAAAAAAH

@@
 \include{PageDeGarde}
+\setcounter{page}{1}
+\newgeometry{bottom=2.5cm}
+
+\include{aknowledgements}
 
@@
 
-\include{aknowledgements}
-
 \selectlanguage{english}
 \tableofcontents \pagebreak
+\listoffigures \pagebreak
 
@@
 
+\include{introduction}%=> What is the problematic of the thesis? What is the challenge? What do we want to show?
+
+\include{technical-background}%=> Background necessary to understand the problematic and its resolution
+
+\include{implementation}%=> How we will study the problematic, what tool are we going to use to make it clear?
+
+\include{test-proto}%=> How we obtained the results
+
+\include{results}%=> Results. Are they expected? Why are they like that? Analyse.
+
+\include{conclusion}
+
 \appendix
+
+% \include{raw-data}
+
+\include{openssh-patch}
+
+% \include{toolchain}
+
+% \include{cross-compilation}
+
+% \include{stuff}
+
+%\include{configuration}
 
@@
 
-% Bibliography
+% \nocite{*}
+\bibliographystyle{plainnat}
+\bibliography{bibliography}
 

### 2015-05-25 results.tex
Bunch of stuff. AAAAAAAAAAAAAAAH

@@
 
-\subsection{IPSec}
+\subsection{IPsec}
 
@@
 \section{File transfer}
-%FTP over openvpn and IPSec
+%FTP over openvpn and IPsec
 
@@
 \input{graphs/ipsec-ftp-bench}
-\caption{FTP file transfer over an IPSec tunnel. The background stripped bars are the CPU usage.}{}
+\caption{FTP file transfer over an IPsec tunnel. The background stripped bars are the CPU usage.}{}
 \label{fig:ipsec-ftp-bench}
@@
 As we saw in chapter~\ref{chap:theory}, modern processor designers tend to add specialized instruction set aimed at AES-GCM enhancement.
-Should further tests be conducted concerning IPSec paired with GCM, it would be wise to compare with an assembly implementation exploiting ARM NEON instruction set.
+Should further tests be conducted concerning IPsec paired with GCM, it would be wise to compare with an assembly implementation exploiting ARM NEON instruction set.
 Some are being developed~\cite{Conrado2013,Danilo2013}, but none have been committed to the Linux kernel repository yet.

### 2015-05-25 technical-background.tex
Bunch of stuff. AAAAAAAAAAAAAAAH

@@
 
-\subsection{Symetric cryptography}
-Talk about encryption, integrity and authentication.
-
-\subsubsection{AES}
-Many modes, CBC is mainly used, GCM is great.
-
-\begin{figure}
-\includegraphics[width=\textwidth]{nist-cbc}
-\caption{CBC encryption and decryption diagram}{taken from the NIST recommendation~\cite{nist-sp800-38A}.}
-\label{fig:cbc-encrypt-decrypt}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=\textwidth]{nist-gcm-encrypt}
-\caption{GCM encryption diagram}{taken from the NIST specification~\cite{mcgrew2005}.
-The ciphertext blocks are formed by \textit{xor}-ing the encrypted counter and the plaintext.
-The tag is generated by a chain of ciphertext \textit{xor}-ing with Galois field multiplicated data.
-The decryption works excatly the same way, except the plaintext and ciphertext are swapped.}
-\label{fig:gcm-encrypt}
-\end{figure}
-
 \subsection{Message digest}
@@
 \citet{Bellare2000} prooved that the most secure solution is to encrypt then compute the MAC from the ciphertext.
-We will see in section~\ref{sec:theory-network} that if IPSec follows this recommandation, SSL/TLS does not and MAC first the plaintext then encrypt the message.
+We will see in section~\ref{sec:theory-network} that if IPsec follows this recommandation, SSL/TLS does not and MAC first the plaintext then encrypt the message.
+
+\subsection{Symetric cryptography}
+Talk about encryption, integrity and authentication.
+
+Look in depth into AES, NIST approved five modes~\cite{nist-sp800-38A}: CBC (Cipher Block Chaining), ECB (Electronic CodeBook), CFB (Cipher FeedBack), OFB (Output FeedBack) and CTR (Counter).
+
+\begin{figure}
+\includegraphics[width=\textwidth]{nist-cbc}
+\caption{CBC encryption and decryption diagram}{taken from the NIST recommendation~\cite{nist-sp800-38A}.}
+\label{fig:cbc-encrypt-decrypt}
+\end{figure}
+
+AEAD => Encryption and authentication: GCM.
+
+\begin{figure}
+\includegraphics[width=\textwidth]{nist-gcm-encrypt}
+\caption{GCM encryption diagram}{taken from the NIST specification~\cite{mcgrew2005}.
+The ciphertext blocks are formed by \textit{xor}-ing the encrypted counter and the plaintext.
+The tag is generated by a chain of ciphertext \textit{xor}-ing with Galois field multiplicated data.
+The decryption works excatly the same way, except the plaintext and ciphertext are swapped.}
+\label{fig:gcm-encrypt}
+\end{figure}
 
@@
 Transport & TCP, UDP & Transport & TCP, UDP \\ \hline
-Internet & IP, ICMP, IPSec & Network & IP, ICMP, IPSec \\ \hline
+Internet & IP, ICMP, IPsec & Network & IP, ICMP, IPsec \\ \hline
 \multirow{2}{*}{Data Link} & \multirow{2}{*}{PPP, MAC, Ethernet, L2TP} & Data link & PPP, MAC, Ethernet, L2TP \\ \cline{3-4}
@@
 
-There exist several major implementations of VPN: SSL, IPSec, L2TP and PPTP.
+There exist several major implementations of VPN: SSL, IPsec, L2TP and PPTP.
 The later was developped by a vendor consortium leaded by Microsoft and proposed in the RFC 2637 and will not be discussed further.
@@
 
-\subsection{IPSec}
-% IPSec as a protocol, strongswan come in the 'implementation' chapter.
+\subsection{IPsec}
+% IPsec as a protocol, strongswan come in the 'implementation' chapter.
 Modification of the IP stack in the kernel space.
-IPSec is a network level security; it examines incomming IP packets and checks if there exists a security association with the destination, and decrypt it on-the-fly if necessary.
+IPsec is a network level security; it examines incomming IP packets and checks if there exists a security association with the destination, and decrypt it on-the-fly if necessary.
 
@@
 
-%TODO Figure IPSec frame structure.
+%TODO Figure IPsec frame structure.
 
@@
 }
-\caption{IPSec transport (a) and tunnel (b) overheads}{Tunnel mode adds a custom IP header and moves the AH/ESP header in front of the original IP header.
+\caption{IPsec transport (a) and tunnel (b) overheads}{Tunnel mode adds a custom IP header and moves the AH/ESP header in front of the original IP header.
 Note: ``trl" stands for ``trailer", ``hdr" for ``header"}
@@
 \end{figure}
+
+RFC 7321~\cite{rfc7321} defines the support for only three AES modes: CBC, CTR and GCM.
+
+%%%%%%%%%%%%%%%%%%
+% We can not invoke a verbatim environement from an other command. We thus save is inside a box and we will invoke said box later.
+\newsavebox\myv
+
+\begin{lrbox}{\myv}\begin{minipage}{\textwidth}
+\begin{verbatim}
+0                   1                   2                   3
+ 0 1 2 3 4 5 6 7 8 9 0 1 2 3 4 5 6 7 8 9 0 1 2 3 4 5 6 7 8 9 0 1
++-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+ -----------
+|               Security Parameters Index (SPI)                 |      ^Int.
++-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+      |Cov-
+|                      Sequence Number                          |      |ered
++-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+---   | ----
+|                    IV (optional)                              | ^ p  |   ^
++-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+ | a  |   |
+|                    Rest of Payload Data  (variable)           | | y  |Conf.
+~                                                               ~ | l  |Cov-
+|                                                               | | o  |ered*
++               +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+ | a  |   |
+|               |         TFC Padding * (optional, variable)    | v d  |   |
++-+-+-+-+-+-+-+-+         +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+---   |   |
+|                         |        Padding (0-255 bytes)        |      |   |
++-+-+-+-+-+-+-+-+-+-+-+-+-+     +-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+      |   |
+|                               |  Pad Length   | Next Header   |      v   v
++-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+------------
+|         Integrity Check Value-ICV   (variable)                |
+~                                                               ~
+|                                                               |
++-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+-+
+\end{verbatim}
+\end{minipage}\end{lrbox}
+
+\begin{figure}
+\center
+	\resizebox{.8\linewidth}{!}{%
+	\usebox\myv
+	}
+\caption{ESP packet structure}{as defined in RFC 4303~\cite{rfc4303}.}
+\label{fig:esp-packet-structure}
+\end{figure}
+
+Don't forget to talk about the IPsec overhead from~\cite{Xenakis20063225}.

### 2015-05-25 test-proto.tex
Theory and shit
Bunch of stuff. AAAAAAAAAAAAAAAH

@@
-\chapter{Test protocol and Prediction}
+\chapter{Test protocol}
 
@@
 	\item[OS] Ubuntu 12.04 LTS, kernel 3.16
-	\item[CPU] Intel Core-i3 ... (two logical core out of four)
+	\item[CPU] Intel Core-i5 (two logical core out of four)
 	\item[RAM] 1GB DDR3
@@
 	\item[CPU] Dual core ARM Cortex-A9, 800MHz
-	\item[RAM] ...GB DDR3
+	\item[RAM] 1GB DDR3
 	\item[FPGA] Altera Cyclone V
@@
 This benchmark is done only with OpenVPN.
-Since there is no standard support for those operation in the kernel yet, it would not have made sense to use IPSec, since it would have had to fallback to OpenSSL, then following the same path as OpenVPN do.
+Since there is no standard support for those operation in the kernel yet, it would not have made sense to use IPsec, since it would have had to fallback to OpenSSL, then following the same path as OpenVPN do.
 As soon as the public key operations can be plugged into the crypto API, this use case should however be tested.
@@
 
-\lstinputlisting[language=bash]{stress-openvpn.sh}
+\lstinputlisting[language=bash, label=list:openvpn-client-script, caption={Script starting ten clients in parallel who will stress the server.}]{stress-openvpn.sh}
 
@@
 \end{lstlisting}
+
+For the IPsec use case, the cipher/authentication pair null/null will be used to quantify the overhead of the encapsulation.
+It should however not be forgotten that it is not to be used as a production configuration.
+As the RFC 7321~\citep[pg. 7]{rfc7321} sates: ``Note that while authentication and encryption can each
+   be `NULL', they MUST NOT both be `NULL'".

### 2015-05-26 implementation.tex
moar theoretical works

@@
 This chapter will present how we implemented the protocols of chapter~\ref{chap:theory} on the board.
-The figure~\ref{fig:os-path-generic} show the generic data flow in the operating system through the user and kernel spaces.
-We will first detail the software part, with the OpenVPN and openSSH as the application, and OpenSSL as a cryptographic library on which both applications rely.
+The figure~\ref{fig:os-path-generic} shows the generic data flow in the operating system through the user and kernel spaces.
+We will first detail the software part, with the OpenVPN and openSSH as the applications, and OpenSSL as a cryptographic library on which both applications rely.
 Then will come strongswan, a user-space abstraction layer giving access to IPsec and we will see how different it is from an OpenVPN implementation.
@@
 % https://openvpn.net/papers/BLUG-talk/index.html
-
-OpenVPN is also used by the Dutch government to secure all its communications~\cite{openvpn-nl}.
+OpenVPN is a very popular VPN solution relying on the application layer of the TCP/IP stack by using the SSL/TLS protocols.
+% OpenVPN is also used by the Dutch government to secure all its communications~\cite{openvpn-nl}.
 

### 2015-05-26 technical-background.tex
moar theoretical works

@@
 
-Follow a quick presentation of FPGAs and how they can be driven from the operating system.
+\noindent Follow a quick presentation of FPGAs and how they can be driven from the operating system.
+
+\noindent Next comes An overview of the cryptographic field with the main primitives: message digests, symmetric ciphers and asymmetric ciphers.
+
+\ident Finally, the ground necessary to understand the inner working and challenges of some VPNs implementations will be covered with SSL/TLS and IPsec.
 
@@
 That memory can be directly mapped and accessed from the user-space using \texttt{/dev/mem}, or can use a direct memory access module (DMA).
-From the operation system, we build a bunch of scatterlist in ther kernel-space memory, then map those pages to memory descriptor that have a physical address on the DMA.
+From the operating system, we build a bunch of scatterlist in the kernel-space memory, then map those pages to memory descriptor that have a physical address on the DMA.
 They can be mapped three different ways: \texttt{DMA\_BIDIRECTIONAL}, \texttt{DMA\_TO\_DEVICE} or \texttt{DMA\_FROM\_DEVICE}.
-When the CPU write something in those descriptor and synchronize them with the DMA, it does not have to care about them anymore, the DMA in now in charge to send them to the device where registers are ready to read the incomming data.
+When the CPU write something in those descriptors and synchronize them with the DMA, it does not have to care about them anymore, the DMA in now in charge to send them to the device where registers are ready to read the incomming data.
 The same goes from the device to the CPU: when the device wants to communicate data to the OS, it writes it on the DMA that will transfer them to the CPU, triggering a flag on the way to notify it.
@@
 % IPsec as a protocol, strongswan come in the 'implementation' chapter.
-Modification of the IP stack in the kernel space.
-IPsec is a network level security; it examines incomming IP packets and checks if there exists a security association with the destination, and decrypt it on-the-fly if necessary.
+IPsec is a modification of the IP stack in the kernel space.
+It is a network/internet level security; it examines incomming IP packets and checks if there exists a security association with the destination, and decrypt it on-the-fly if necessary.
 
-One of the main disadvantage compared to a user-space VPN is the difficulty to traverse NAT.
+% One of the main disadvantage compared to a user-space VPN is the difficulty to traverse NAT.
 
-%TODO Figure IPsec frame structure.
+IPsec consist of three components~\cite{cryptoencyclopedia2011}, each defined in their respective RFC:
+\begin{description}
+	\item[Traffic protols] Encapsulating Security Payload (ESP,~\cite{rfc4303}) and Authentication Header (AH,~\cite{rfc4302}).
+	\item[Key management] Internet Key Exchange (IKEv2,~\cite{rfc7296}).
+	\item[Policy]Security Policy Database (SPD,~\cite{rfc4301}) and Security Association (SA~\cite{rfc4301}).
+\end{description}
+
+% ESP: TFC padding: traffic flow confidentiality, padding to hide the true nature of the packet, security mesure.
+% Link to the figure fig:esp-packet-structure
+Arguably, the features of AH can be offered by ESP.
+The existance of AH is actually historical; when the protocol was developped in the 1990s, there were laws in the United States and other countries preventing the export of product including encryption features.
+Having an authentication only counterparty allowed such product to be exported without encryption capabilities.\newline{}
+
+IPsec can be deployed in two modes: transport for end-to-end service and protecting the payload, and tunnel to typically connect gateways and protect the entire IP packet.
+The figure~\ref{fig:ipsec-transport-tunnel} shows the space overhead imposed by AH and ESP in transport and tunnel modes.
+
+Both protocols add 10 bytes of fixed size fields (SPI, SN, pad length and next header).
+The other fields are of variable length, or even optional, such as the optional authentication field of minimum 12 bytes, the optional IV of minimum 4 bytes and the padding between 0 and 255 bytes long.
+\citet{Xenakis20063225} reached the conclusion in their paper that the use of lightweight -- and deprecated -- encryption schemes (\textit{e.g.} HMAC-MD5 and DES) hardly had any impact on the throughput of the system and on the delay of the packets.
+Hence, the space overhead of IPsec is negligible when using more serious schemes (\textit{e.g.} AES-256-CBC with HMAC-SHA-256).
 
@@
 }
-\caption{IPsec transport (a) and tunnel (b) overheads}{Tunnel mode adds a custom IP header and moves the AH/ESP header in front of the original IP header.
-Note: ``trl" stands for ``trailer", ``hdr" for ``header"}
+\caption{IPsec (a) transport and (b) tunnel overheads}{Tunnel mode adds a custom IP header and moves the AH/ESP header in front of the original IP header.
+Note: ``trl" stands for ``trailer", ``hdr" for ``header", reproduced from~\cite{Xenakis20063225}}
 \label{fig:ipsec-transport-tunnel}
@@
 \end{figure}
-
-Don't forget to talk about the IPsec overhead from~\cite{Xenakis20063225}.

### 2015-05-26 technical-background.tex
Theory: IPsec

@@
 
-\ident Finally, the ground necessary to understand the inner working and challenges of some VPNs implementations will be covered with SSL/TLS and IPsec.
+\noindent Finally, the ground necessary to understand the inner working and challenges of some VPNs implementations will be covered with SSL/TLS and IPsec.
 
@@
 \begin{description}
-	\item[Traffic protols] Encapsulating Security Payload (ESP,~\cite{rfc4303}) and Authentication Header (AH,~\cite{rfc4302}).
+	\item[Traffic protocols] Encapsulating Security Payload (ESP,~\cite{rfc4303}) and Authentication Header (AH,~\cite{rfc4302}).
 	\item[Key management] Internet Key Exchange (IKEv2,~\cite{rfc7296}).
@@
 
-% ESP: TFC padding: traffic flow confidentiality, padding to hide the true nature of the packet, security mesure.
-% Link to the figure fig:esp-packet-structure
-Arguably, the features of AH can be offered by ESP.
+As~\citet{Paterson200672} present them, the SAs are used as repository for cryptographic parameters, and the SPD defines the policies to apply.
+If the SPD can be populated by hand, typically in a configuration file, the SA management should be left to an automated mechanism: the IKEv2.
+
+\noindent The IKEv2 protocol uses four different types of exchanges to fulfill its role:
+\begin{itemize}
+	\item \texttt{IKE\_SA\_INIT}: The two peers agree on cryptographic parameters to use. Among these is a Diffie-Hellman shared secret from which are derived symmetric keys used for a special IKE SA.
+	\item \texttt{IKE\_AUTH}: Once the peers are protected by the IKE SA, they can authenticate themselves to each other and add a first SA in the SA database (SADB).
+	The protocol supports three types of authentication methods:
+	\begin{itemize}
+		\item Digital signature using a PKI.
+		\item MAC using a pre-established secret key.
+		\item Extensible Authentication Protocol (EAP) defined in~\citet{rfc3748}.
+	\end{itemize}
+	At this point, the connection is fully established and ready to use.
+	\item \texttt{CREATE\_CHILD\_SA}: Used to create new SAs between the two peers. It may involve new generations of DH secrets to ensure perfect forward secrecy.
+	\item \texttt{INFORMATIONAL}: Exchange of management information.
+\end{itemize}\newline{}
+
+The figure~\ref{fig:esp-packet-structure} shows the structure of an ESP packet.
+The ESP header includes two fields: a Security Parameter Index (SPI) and a Sequence Number (SN)
+The SPI is a value on which both actors agreed on during the key exchange protocol, identifying the SA to use for the communication.
+The SN is an anti-replay feature, an integer incrementing with each packet processed by the SA.
+This way, the SA will reject any packet that does not have a SN in the current windows, avoiding an intruder to resend a previously captured packet on the network.
+As the SN is covered by the integrity, the opponent can not change it without having to change the Integrity Check Value (ICV, \textit{a.k.a.} the message digest).
+This authentication covers the whole packet, whilst the confidentiality excludes the ESP header.
+
+\noindent The ESP trailer includes a padding and two 1-byte fields: the padding length and the next header value, which can specify if the packet is dummy or real.
+The padding is needed when using block ciphers to obtain a payload which length is an integral multiple of the block length.
+
+\noindent The ESP payload includes an optional IV and an optional Traffic Flow Confidentiality (TFC) padding.
+The TFC padding allows to modify the length of the packet with dummy data, somewhat hiding the true nature of the payload.
+It differs from the standard padding because its length is not limited to 255 bytes like the latter.
+
+\noindent AH is the same as ESP, except is does not offer encryption features, only the authentication.
+Since ESP can use a null cipher, it arguably offers the same features as AH.
 The existance of AH is actually historical; when the protocol was developped in the 1990s, there were laws in the United States and other countries preventing the export of product including encryption features.
@@
 
-IPsec can be deployed in two modes: transport for end-to-end service and protecting the payload, and tunnel to typically connect gateways and protect the entire IP packet.
+IPsec can be deployed in two modes: 
+\begin{itemize}
+	\item Transport for end-to-end service and protecting the payload. IPsec examines the incomming payload, checks it has not been tampered with and pass it to the next layer.
+	\item Tunnel protects the entire IP packet by encapsulating it into a new IP packet.
+As the destination contained in the new IP header can be different from the original, this mode can be used to manage gateways.
+\end{itemize}
 The figure~\ref{fig:ipsec-transport-tunnel} shows the space overhead imposed by AH and ESP in transport and tunnel modes.

### 2015-05-26 results.tex
colors

@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{TLS connections}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 % Begin with the benchmark done by Bastien on the raw number of verif/s with openssl.
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Response time -- latency}
-After the establishement of a connection.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The following tests are conducted after the connection has been established, so as the clients do not need to undergo any new key negociation.
 
@@
 \input{graphs/openvpn-ping-benchmark}
-\caption{Ping min/avg/max response time}{for different packet sizes using OpenVPN}
+\caption{OpenVN: ping average response time}{}
 \label{fig:ping-benchmark-openvpn}
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{File transfer}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %FTP over openvpn and IPsec

### 2015-05-26 results.tex
Remove size 4000 from ping
Ping analysis

@@
 \subsection{OpenVPN}
+The figure~\ref{fig:ping-benchmark-openvpn} shows the results for different payload sizes.
+
+\noindent Firstly, is the impact of OpenVPN's operation on the delay: when no encryption nor authentication is used, the exchange is 33\% to 53\% longer.
+This is simply because the packet has to go through a virtual interface, hence even without the security time overhead, the latency is much longer.
+
+When combining encryption and authentication, the software and hardware implementation are neck and neck.
+Eventhough the MAC is not offload, the hardware implementation could have performed better.
+The reason is not only all the packets have to go through a virtual interface in the kernel, the payload also has to be sent to the hardware through the kernel.
+In the end, a packet will undergo two full round trips between the user and the kernel space, that is one more than a regular software implementation, plus a final transfer to the physical interface.
+
+\noindent All those transfer make the offload useless in such a case in term of raw performance.
+The CPU usage is not even worth mentioning as the operation is only a few milliseconds long.
 
@@
 \input{graphs/openvpn-ping-benchmark}
-\caption{OpenVN: ping average response time}{}
+\caption{OpenVN: ping average response time}{OpenVPN adds up to 53\% extra delay and offloading the encryption does not improve the latency.}
 \label{fig:ping-benchmark-openvpn}
@@
 \subsection{IPsec}
+The figure~\ref{fig:ping-benchmark-ipsec} studies the same payload size as OpenVPN, but with an extra experimental hardware implementation of AES mode GCM.
+
+\noindent In this case, the time overhead imposed by IPsec is not larger than 9\%, corroborating the results of~\citet{Xenakis20063225}.
+
+When combining encryption and authentication, the hardware implementation steadily takes the advantage over the software with the encrease of the payload length.
+If they both have the same latency at default payload size, the hardware is up to 13\% faster for 16000 bytes payloads.
+
+As for the GCM mode, the support in the BA411E driver is highly experimental and takes more of a proof of concept than a real use case.
+The current implementation could not support payload sizes larger than 1000 bytes, and at this size it is 5\% faster than the software.
+It is not much, but when compared to the 3\% improvement the BA411E offers at the same size for AES-256-CBC, we can expect an accelaration exceeding the 20\% for larger payload sizes.
+%TODO Instead of comparing with sha256, compare with none, no software holding back;
+Aside from the fact that the driver is not fully working yet, it could still be extensively improved.
+As an example, it still relies on the \texttt{seqiv} kernel module, as the software implementation does.
+Managing the generation of the IV directly in the driver would cut it loose from yet another software boulder.
 
@@
 \input{graphs/ipsec-ping-benchmark}
-\caption{Ping min/avg/max response time}{for different packet sizes using IPsec.
+\caption{IPsec: ping average response time}{for different packet sizes using IPsec.
 For each packet size, 1000 requests were flooded to the board, that is \textit{"outputs packets as fast as they come back or one hundred times per second, whichever is more"}, according to the \texttt{ping} command manual.} %TODO move the end of the caption to the "test protocol" chapter.
@@
 \subsection{Comparison}
+The figure~\ref{fig:ping-benchmark-comparison} summarizes the results for OpenVPN and IPsec for the most realistic use case: the combination of encryption and authentication.
+
+\noindent The first main difference is the overhead imposed by each method: 53\% for OpenVPN and only 9\% for IPsec.
+Even if those configurations are not realistic, it puts forward the advantage of directly working in the kernel as does IPsec.
+
+As for the combination of encryption and authentication, IPsec is between 13\% and 32\% faster than OpenVPN.
+IPsec loses its advantage with the increase of the payload size, the time lost when moving around the data between the user and the kernel space being less important compared to the securoty operations.
+
+\noindent In hardware, as OpenVPN could not take advantage of the accelaration of the encryption, IPsec is even faster, ranging from 25\% to 33\% faster than OpenVPN.
 
@@
 \input{graphs/ping-bench-comparison}
-\caption{Ping average comparison}{All values are for aes-256-cbc with sha-256}
+\caption{Ping average comparison}{All values are for AES-256-CBC with SHA-256.
+Globally, IPsec yields better results and propose the most significant hardware offload.
+Quick reading: the stripped bars are the IPsec results.}
 \label{fig:ping-benchmark-comparison}
@@
 %			- capped to software max: show CPU offload
+This use case studies the performance of a simple file transfer over three different secure implementations: OpenSSH, OpenVPN and IPsec.
+For each application, three encryption:authentication couples are considered: none:none, AES-256-CBC:none and AES-256-CBC:SHA-256.
 

### 2015-05-26 conclusion.tex
Results

@@
 %Note this is still under heavy criticism. If the hardware lags so far behind, why the relatively good score in openvpn?
-The figure \ref{fig:openssl-benchmark-cbc} clearly shows a threshold on the packet size from which the hardware has a clear advantage, and below which the user mode software implementation is to go for.
+The figure \ref{fig:openssl-speed} clearly shows a threshold on the packet size from which the hardware has a clear advantage, and below which the user mode software implementation is to go for.
 %TODO include a listing to where it could be a good place to branch on the packet size. I already found it.

### 2015-05-26 graphs/ipsec-ftp-bench.tex
Results

@@
 
-\addplot[style={black,fill=brown,postaction={pattern=north east lines},mark=none}]
+\addplot[style={black,fill=NavyBlue,postaction={pattern=north east lines},mark=none}]
     coordinates {
@@
 
-\addplot[style={black,fill=RawSienna,mark=none}]
+\addplot[style={black,fill=MidnightBlue,mark=none}]
     coordinates {

### 2015-05-26 graphs/openssh-bench.tex
Results

@@
 \begin{axis}[
-        title = {file transfer over SSH tunnel},
+        title = {File transfer over an SSH tunnel},
         width  = 0.95*\textwidth,

### 2015-05-26 graphs/openvpn-ftp-bench.tex
Results

@@
 \begin{axis}[
-        title = {FTP transfer inside OpenVPN tunnel},
+        title = {FTP transfer inside an OpenVPN tunnel},
         width  = 0.95*\textwidth,

### 2015-05-26 results.tex
Results

@@
 \chapter{Results and Analysis}\label{chap:results}
+
+This chapter gather the results of the use cases presented in chapter~\ref{chap:test-protocol}.
+Three use cases are considered: establishing TLS connections, exchanging ICMP requests to compute the latency and transfering a file over a secure channel.
+Depending on the case, several schemes are analysed: DHE-RSA, AES-CBC, AES-GCM and SHA-256.
+
+It is to be noted that all the CPU usage values are taken from the single one loaded core.
+Although the platform is dual-core, all applications are used single threaded, either by design, such as OpenVPN, or by choice to obtain comparable results.
 
@@
 
-\noindent Firstly, is the impact of OpenVPN's operation on the delay: when no encryption nor authentication is used, the exchange is 33\% to 53\% longer.
+\begin{figure}[ht]
+\input{graphs/openvpn-ping-benchmark}
+\caption{OpenVN: ping average response time}{OpenVPN adds up to 53\% extra delay and offloading the encryption does not improve the latency.}
+\label{fig:ping-benchmark-openvpn}
+\end{figure}
+
+Firstly, is the impact of OpenVPN's operation on the delay: when no encryption nor authentication is used, the exchange is 33\% to 53\% longer.
 This is simply because the packet has to go through a virtual interface, hence even without the security time overhead, the latency is much longer.
@@
 
-\begin{figure}[ht]
-\input{graphs/openvpn-ping-benchmark}
-\caption{OpenVN: ping average response time}{OpenVPN adds up to 53\% extra delay and offloading the encryption does not improve the latency.}
-\label{fig:ping-benchmark-openvpn}
-\end{figure}
-
 \subsection{IPsec}
@@
 
-\noindent In this case, the time overhead imposed by IPsec is not larger than 9\%, corroborating the results of~\citet{Xenakis20063225}.
+\begin{figure}[ht]
+\input{graphs/ipsec-ping-benchmark}
+\caption{IPsec: ping average response time}{for different packet sizes.}
+\label{fig:ping-benchmark-ipsec}
+\end{figure}
+
+In this case, the time overhead imposed by IPsec is not larger than 9\%, corroborating the results of~\citet{Xenakis20063225}.
 
@@
 
-\begin{figure}[ht]
-\input{graphs/ipsec-ping-benchmark}
-\caption{IPsec: ping average response time}{for different packet sizes using IPsec.
-For each packet size, 1000 requests were flooded to the board, that is \textit{"outputs packets as fast as they come back or one hundred times per second, whichever is more"}, according to the \texttt{ping} command manual.} %TODO move the end of the caption to the "test protocol" chapter.
-\label{fig:ping-benchmark-ipsec}
-\end{figure}
-
 \subsection{Comparison}
 The figure~\ref{fig:ping-benchmark-comparison} summarizes the results for OpenVPN and IPsec for the most realistic use case: the combination of encryption and authentication.
-
-\noindent The first main difference is the overhead imposed by each method: 53\% for OpenVPN and only 9\% for IPsec.
-Even if those configurations are not realistic, it puts forward the advantage of directly working in the kernel as does IPsec.
-
-As for the combination of encryption and authentication, IPsec is between 13\% and 32\% faster than OpenVPN.
-IPsec loses its advantage with the increase of the payload size, the time lost when moving around the data between the user and the kernel space being less important compared to the securoty operations.
-
-\noindent In hardware, as OpenVPN could not take advantage of the accelaration of the encryption, IPsec is even faster, ranging from 25\% to 33\% faster than OpenVPN.
 
@@
 \end{figure}
+
+The first main difference is the overhead imposed by each method: 53\% for OpenVPN and only 9\% for IPsec.
+Even if those configurations are not realistic, it puts forward the advantage of directly working in the kernel as does IPsec.
+
+As for the combination of encryption and authentication, IPsec is between 13\% and 32\% faster than OpenVPN.
+IPsec loses its advantage with the increase of the payload size, the time lost when moving around the data between the user and the kernel space being less important compared to the securoty operations.
+
+\noindent In hardware, as OpenVPN could not take advantage of the accelaration of the encryption, IPsec is even faster, ranging from 25\% to 33\% faster than OpenVPN.
 
@@
 
-\subsection{openSSH}
+\subsection{OpenSSH}
+The figure~\ref{fig:openssh-bench} shows the results for a file transfer over an SSH tunnel.
+For this application, there is no none:none couple case inside the tunnel, as it would be the same as a transfer outside the tunnel.
 
@@
 \input{graphs/openssh-bench}
-\caption{file transfer over an SSH tunnel. The background stripped bars are the CPU usage.}{}
+\caption{File transfer over an SSH tunnel.}{The background stripped bars are the CPU usage.}
 \label{fig:openssh-bench}
@@
 
+When only encryption is used, both implementation performs almost as well as outside the tunnel, but the software already saturates the CPU, whilst the hardware only uses half the same ressources.
+About 41\% of the operations using the CPU when accelarating with the hardware involve interuptions and kernel memory management.
+
+Adding the authentication makes the software performance drop by 25\%, as the CPU was already saturated without those extra operations.
+The hardware is also consistent, but as there were some ressources available, the throughput is merely affected, even if the MAC is not offloaded in hardware.
+
 \subsection{OpenVPN}
+The figure~\ref{fig:openvpn-ftp-bench} shows the performance of the file transfer over an OpenVPN tunnel.
 
@@
 \input{graphs/openvpn-ftp-bench}
-\caption{FTP file transfer over an OpenVPN tunnel. The background stripped bars are the CPU usage.}{}
+\caption{FTP file transfer over an OpenVPN tunnel.}{The results inside the tunnel with no encryption nor authentication highlights the overhead OpenVPN adds to the transfer.
+From there, the throughput only change by a small extent.
+The poor hardware results also shows the heaviness of OpenVPN on hardware offloading.
+The background stripped bars are the CPU usage.}
 \label{fig:openvpn-ftp-bench}
@@
 
-We can see that adding a MAC computation aside the encryption merely lowers the performance when using the hardware.
-Even though OpenSSL uses here an hihgly ARM-optimzed assembly implementation of SHA-256, it shows that the bottleneck is on the hardware side.
+As it was already the case for the ping benchmark, the overhead imposed by the manipulation of OpenVPN is extremely heavy: the CPU usage jumps from 7.16\% to 42.60\%, and the throughput drops by 55\%, from 11.39MB/s to 5.18MB/s.
 
-Indeed,
+\noindent Surprisingly enough, encrypting the data only changes the throughput by 8\%, but the CPU usage is almost doubled.
+A fair interpretation would be that the encryption is no the bottleneck, the transfer bewteen the user and kernel mode, as well as the fragmentation, are.
+However for the hardware, the performance collapse to 3.35MB/s.
+
+Adding a MAC computation aside the encryption lowers the performance in software and hardware by respectively 20\% and 15\%, for the same CPU usage as without authentication.
 
 \subsection{IPsec}
+The figure~\ref{fig:ipsec-ftp-bench} gathers the results of the file transfer over the IPsec tunnel.
+\begin{figure}[ht]
+\input{graphs/ipsec-ftp-bench}
+\caption{FTP file transfer over an IPsec tunnel.}{The advantage of the hardware is a lower CPU usage, at the cost of a few percents lower throughput.
+The GCM mode is also tested to discuss its implementation in hardware.
+The background stripped bars are the CPU usage.}
+\label{fig:ipsec-ftp-bench}
+\end{figure}
+
+In this implementation, the IPsec overhead is only of 10\%, close to the 9\% of the ping overhead, and the CPU usage only double to reach 14.68\%.
+
+When encrypting the data, the CPU utilization explodes in the case of the software implementation, but stays at the same level for the hardware.
+As for the throughput, it lowers respectively by 14\% and 17\%.
+What could have been expected for the hardware implementation is a more stable throughput, but an increase of CPU usage for the few operations undergone by the BA411E driver, especially considering the fact that it uses active polling on the hardware.
+Thus, even if the hardware is the bottleneck, the CPU usage should be higher.
+
+\noindent An explanation could be that the operating system prevent the driver to monopolizing the CPU with its polling and preempt it regularly, effectively lowering the ressource usage, but also limiting the performance.
+
+Adding the authentication yields expected results for : an increase of CPU usage and lower throughput.
+Both implementation lose 2MB/s and a CPU usage increase of 20\% and 16\%.
+
+In both cases, the hardware implementation exhibited slightly lower performance, but a CPU usage three to four times lower.
+On embedded platforms, this is as important as the raw performance, because lower CPU usage means underclocked CPU, resulting in a lower power consumption.
+
+The softaware GCM results allows to open a discussion on this mode used in conjunction with IPsec.
+The GCM performance presented clearly shows a drop of throughput and an increase of CPU usage, illustrating the fact that those operations are hard on the software.
+With an hardware offload, we could expect not only a drastic drop of the CPU usage, but an increase of throughput as well, since it's CPU-limited in those results.
+Note that they are achieved using a C-based implementation of galois-field multiplications.
+As we saw in chapter~\ref{chap:theory}, modern processor designers tend to add specialized instruction sets aimed at AES-GCM enhancement.
+Should further tests be conducted concerning IPsec paired with GCM, it would be wise to compare with an assembly implementation exploiting ARM NEON instruction set.
+Some are being developed~\cite{Conrado2013,Danilo2013}, but none have been committed to the Linux kernel repository yet.
+
+\subsection{Comparison}
+The figure~\ref{fig:ftp-bench-comparison} summarizes the file transfer performance for all the implementations, using the most realistic configuration tested: AES-256-CBC with SHA-256.
+\begin{figure}[ht]
+\input{graphs/ftp-bench-comparison}
+\caption{Comparison of file transfer methods}{Out of the three methods, OpenSSH is to be prefered is the raw performance is the objective, but IPsec offers an higher effeciency.
+At the same time, OpenVPN is worse on all fronts.
+The background stripped bars are the CPU usage.}
+\label{fig:ftp-bench-comparison}
+\end{figure}
+
+In this work, we have two objectives: the raw performance and the CPU usage.
+The first objetive is fulfilled by OpenSSH; even if it is a user-space application that needs to transfer the data through the kernel in order to use the hardware, it still outperforms the software by 27\%.
+
+\noindent The second objective is best fulfilled by IPsec with an efficiency of 0.34MB/s/\% of CPU for the hardware, and 0.09MB/s/\% of CPU for the software.
+
+OpenVPN is however losing on all fronts. Not only it has the lowest throughput, but the CPU usage of the hardware is out the roof, especially when compared to the performance.
+
+\noindent The table~\ref{tab:ftp-fragmentation} shows to what size are fragmented the packets by OpenVPN and IPsec.
+A look at the figure~\ref{fig:openssl-speed} shows that fragmenting at small size is a bad idea, especially when the hardware is reached from the user space.
+OpenVPN sends one third of its packets at sizes that are highly ineffective, whilst IPsec manages it better and limit the small packets to less than 10\%.
+
+\begin{table}
+\center
+\begin{tabular}{l|c|c|} \cline{2-3}
+ & Length & Frequency \\ \hline
+\multicolumn{1}{|l|}{\multirow{3}{*}{OpenVPN}} & 80 -- 159 & 33.04\% \\ \cline{2-3}
+\multicolumn{1}{|l|}{} & 1280 -- 2559 & 66.90\% \\ \cline{2-3}
+\multicolumn{1}{|l|}{} & other & 0.06\% \\ \hline
+\multicolumn{1}{|l|}{\multirow{3}{*}{IPsec}} & 80 -- 159 & 9.42\% \\ \cline{2-3}
+\multicolumn{1}{|l|}{} & 1280 -- 2559 & 89.84\% \\ \cline{2-3}
+\multicolumn{1}{|l|}{} & other & 0.74\% \\ \hline
+\end{tabular}
+\caption{Packet size frequency for an FTP transfer}{OpenVPN is less effective in its fragmentation, producing more packets of a smaller size, which perform worse on hardware.}
+\label{tab:ftp-fragmentation}
+\end{table}
 
 \begin{figure}[ht]
-\input{graphs/ipsec-ftp-bench}
-\caption{FTP file transfer over an IPsec tunnel. The background stripped bars are the CPU usage.}{}
-\label{fig:ipsec-ftp-bench}
+\input{graphs/openssl-speed}
+\caption{OpenSSL benchmark for AES-256-CBC}{compare a run of \texttt{openssl speed} for three implementation of AES: in software by OpenSSL (soft), in software by the standard Linux kernel module (soft-drv) and in hardware (hard).
+The last two have to go through cryptodev in order to be reached from the user space.}
+\label{fig:openssl-speed}
 \end{figure}
@@
 A more recent one using assembly instructions optimized for the NEON SIMD instruction set of the ARMv7 core could be used and would most probably yield better results.
-The CPU usage of the software implementation would drop -- even if not significantly -- as for the hardware, it would be less limited by the software MAC counter part, and if the CPU usage could stay at the same level, we could expect a better throughput.
-
-The GCM performance presented clearly shows a drop of throughput and an increase of CPU usage, illustrating the fact that those operations are hard on the software.
-With an hardware offload, we could expect not only a drastic drop of the CPU usage, but an increase of throughput as well, since it's CPU-limited in those results.
-Note that they are achieved usinf a C-based implementation of galois-field multiplications.
-As we saw in chapter~\ref{chap:theory}, modern processor designers tend to add specialized instruction set aimed at AES-GCM enhancement.
-Should further tests be conducted concerning IPsec paired with GCM, it would be wise to compare with an assembly implementation exploiting ARM NEON instruction set.
-Some are being developed~\cite{Conrado2013,Danilo2013}, but none have been committed to the Linux kernel repository yet.
-
-\subsection{Comparison}
-
-\begin{figure}[ht]
-\input{graphs/ftp-bench-comparison}
-\caption{Comparison of file transfer methods}{The background stripped bars are the CPU usage.}
-\label{fig:ftp-bench-comparison}
-\end{figure}
+The CPU usage of the software implementation would lower -- even if not significantly -- as for the hardware, it would be less limited by the software MAC counter part, and if the CPU usage would stay at the same level, we could expect a better throughput.
+%TODO table comparing the packet size (wireshark)

### 2015-05-26 test-proto.tex
moar theoretical works
Results

@@
-\chapter{Test protocol}
+\chapter{Test protocol}\label{chap:test-protocol}
 
@@
 
+As this experiment can be very unstable and vastly depends on the operating system scheduling, each test case has been repeated five times to ensure stable results.
+
 \lstinputlisting[language=bash, label=list:openvpn-client-script, caption={Script starting ten clients in parallel who will stress the server.}]{stress-openvpn.sh}
+
+The security parameter tested is a standard TLS-DHE-RSA, hence forcing the peers the renogociate a new shared secret and ephemeral Diffie-Hellman parameters at each connection attempt.
+
+\section{Response time -- latency}
+This use case exchanges ICMP request of various sizes via the \texttt{ping} command.
+The initiating peer sends an ICMP echo request to the remote peer, which then answers with an ICMP echo reply.
+
+\noindent For each packet size, 1000 requests were flooded to the board, that is \textit{"outputs packets as fast as they come back or one hundred times per second, whichever is more"}, according to the \texttt{ping} command manual.
+
+The following loop shows the options used for the test as well as the payload sizes:
+\begin{lstlisting}[language=bash]
+for i in 56 1000 8000 16000; do
+	sudo ping -f -s ${i} -c 1000 150.158.232.241
+done
+\end{lstlisting}
 
@@
    be `NULL', they MUST NOT both be `NULL'".
+
+We will also use two types of drivers for the BA411E: an interuption-based for OpenVPN and OpenSSH, and a polling-based for IPsec.
+
+All the tests will be conducted with the ESP protocol in tunnel mode, so that we have the worst case scenario; as AH imposes a smaller overhead, the perfomance could only be better.
+
+For OpenSSH, the commmand \texttt{scp} will be used to transfer the data securely over an SSH tunnel.
+As for OpenVPN and IPsec, a tunnel will be established beforehand, and then the simple \texttt{ftp} command will allow the client to fetch the file on the server.

### 2015-05-26 graphs/ipsec-ftp-bench.tex
raw results

@@
 \begin{axis}[
-        title = {FTP transfer insed IPSec tunnel},
+        title = {FTP transfer inside an IPSec tunnel},
         width  = 0.95*\textwidth,

### 2015-05-26 graphs/openvpn-tls-bench.tex
raw results

@@
         ymajorgrids = true,
-        ylabel = {Throughtput [MB/s]},
+        ylabel = {Connections per minute},
         xlabel = {},

### 2015-05-26 implementation.tex
raw results

@@
 
-\noindent The first liberty is a laste\footnote{A vulnerability note as been issued by Carnegie Mellon University Computer Emergency Response Team in early 2009 (last revision)~\cite{CERT2009}, in response to a research of University of London~\cite{Albrecht:2009} presenting a plaintext-recovering attack against SSH when CBC mode is used, but the OpenSSH update took only place in october 2014.}.
+\noindent The first liberty is a response to an old vulnarability\footnote{A vulnerability note as been issued by Carnegie Mellon University Computer Emergency Response Team in early 2009 (last revision)~\cite{CERT2009}, in response to a research of University of London~\cite{Albrecht:2009} presenting a plaintext-recovering attack against SSH when CBC mode is used, but the OpenSSH update took only place in october 2014.}.
 Even if it is not recommended, the user can still enable this mode by explicitly configuring it in the \texttt{sshd\_config} options.

### 2015-05-26 results.tex
raw results

@@
 \subsection{OpenVPN}
-The figure~\ref{fig:ping-benchmark-openvpn} shows the results for different payload sizes.
+The figure~\ref{fig:ping-benchmark-openvpn} shows the results for different payload sizes, which raw results are presented in table~\ref{tab:ping-benchmark-openvpn}
 
@@
 \end{figure}
+
+\begin{table}[ht]
+\center
+\small
+\begin{tabular}{l|c|c|c|c|} \cline{2-5}
+ & 56B & 1000B & 8000B & 16000B \\ \hline
+\multicolumn{1}{|l|}{raw} & 1.444 & 1.929 & 2.811 & 4.322 \\ \hline
+\multicolumn{1}{|l|}{none-none} & 2.066 & 2.561 & 4.293 & 6.117 \\ \hline
+\multicolumn{1}{|l|}{soft-cbc256-none} &  &  &  & \\ \hline
+\multicolumn{1}{|l|}{ba411e-cbc256-none} &  &  &  & \\ \hline
+\multicolumn{1}{|l|}{soft-cbc256-sha256} & 2.415 & 3.061 & 5.997 & 9.135 \\ \hline
+\multicolumn{1}{|l|}{ba411e-cbc256-sha256} & 2.416 & 3.052 & 5.963 & 9.207 \\ \hline
+\end{tabular}
+\caption{OpenVN: ping average response time}{the latency is given in milliseconds.}
+\label{tab:ping-benchmark-openvpn}
+\end{table}
 
@@
 The figure~\ref{fig:ping-benchmark-ipsec} studies the same payload size as OpenVPN, but with an extra experimental hardware implementation of AES mode GCM.
+The raw results are in table~\ref{tab:ping-benchmark-ipsec}.
 
@@
 \end{figure}
+
+\begin{table}[ht]
+\center
+\small
+\begin{tabular}{l|c|c|c|c|} \cline{2-5}
+ & 56B & 1000B & 8000B & 16000B \\ \hline
+\multicolumn{1}{|l|}{raw} & 1.444 & 1.929 & 2.811 & 4.322 \\ \hline
+\multicolumn{1}{|l|}{none-none} & 1.545 & 2.045 & 2.997 & 4.703 \\ \hline
+\multicolumn{1}{|l|}{soft-cbc256-none} &  &  &  &  \\ \hline
+\multicolumn{1}{|l|}{ba411e-cbc256-none} &  &  &  &  \\ \hline
+\multicolumn{1}{|l|}{soft-cbc256-sha256} & 1.645 & 2.322 & 4.762 & 7.975 \\ \hline
+\multicolumn{1}{|l|}{ba411e-cbc256-sha256} & 1.635 & 2.246 & 4.170 & 6.929 \\ \hline
+\multicolumn{1}{|l|}{soft-gcm256} & 1.651 & 2.388 & 5.383 & 9.241 \\ \hline
+\multicolumn{1}{|l|}{ba411e-gcm256} & 1.764 & 2.286 & -- & --  \\ \hline
+\end{tabular}
+\caption{IPsec: ping average response time}{the latency is given in milliseconds. Some results for the GCM mode in hardware are non-existant.}
+\label{tab:ping-benchmark-ipsec}
+\end{table}
 
@@
 For this application, there is no none:none couple case inside the tunnel, as it would be the same as a transfer outside the tunnel.
+The raw results are gathered in table~\ref{tab:openssh-bench}.
 
@@
 \input{graphs/openssh-bench}
-\caption{File transfer over an SSH tunnel.}{The background stripped bars are the CPU usage.}
+\caption{File transfer over an SSH tunnel}{The background stripped bars are the CPU usage.}
 \label{fig:openssh-bench}
 \end{figure}
+
+\begin{table}[ht]
+\center
+\small
+\begin{tabular}{l|c|c|c|c|c|c|} \cline{2-7}
+ & \multicolumn{2}{c|}{none:none} & \multicolumn{2}{c|}{aes256cbc:none} & \multicolumn{2}{c|}{aes256cbc:sha256} \\ \cline{2-7}
+ & Tp.  & CPU & Tp.  & CPU & Tp.  & CPU \\ \hline
+\multicolumn{1}{|c|}{Out-of-tunnel} & 11.39 & 7.16 & -- & -- & -- &  -- \\ \hline
+\multicolumn{1}{|c|}{Software} & -- & -- & 10.89 & 96.26 & 8.19 &  89.29 \\ \hline
+\multicolumn{1}{|c|}{BA414E} & -- & -- & 10.67 & 46.01 & 10.39 & 68.19 \\ \hline
+\end{tabular}
+\caption{File transfer over an SSH tunnel}{The throughput is in MB/s.}
+\label{tab:openssh-bench}
+\end{table}
 
@@
 The figure~\ref{fig:openvpn-ftp-bench} shows the performance of the file transfer over an OpenVPN tunnel.
+The data is gathered in table~\ref{tab:openvpn-ftp-bench}.
 
@@
 \input{graphs/openvpn-ftp-bench}
-\caption{FTP file transfer over an OpenVPN tunnel.}{The results inside the tunnel with no encryption nor authentication highlights the overhead OpenVPN adds to the transfer.
+\caption{FTP file transfer over an OpenVPN tunnel}{The results inside the tunnel with no encryption nor authentication highlights the overhead OpenVPN adds to the transfer.
 From there, the throughput only change by a small extent.
@@
 \end{figure}
+
+\begin{table}[ht]
+\center
+\small
+\begin{tabular}{l|c|c|c|c|c|c|} \cline{2-7}
+ & \multicolumn{2}{c|}{none:none} & \multicolumn{2}{c|}{aes256cbc:none} & \multicolumn{2}{c|}{aes256cbc:sha256} \\ \cline{2-7}
+ & Tp.  & CPU & Tp.  & CPU & Tp.  & CPU \\ \hline
+\multicolumn{1}{|c|}{Out-of-tunnel} & 11.39 & 7.16 & -- & -- & -- & -- \\ \hline
+\multicolumn{1}{|c|}{Inside tunnel} & 5.18 & 42.60 & -- & -- & -- & -- \\ \hline
+\multicolumn{1}{|c|}{Software} & -- & -- & 4.78 & 76.60 & 3.87 & 76.03 \\ \hline
+\multicolumn{1}{|c|}{BA414E} & -- & -- & 3.35 & 83.74 & 2.84 & 80.89 \\ \hline
+\end{tabular}
+\caption{FTP file transfer over an OpenVPN tunnel}{The throughput is in MB/s.}
+\label{tab:openvpn-ftp-bench}
+\end{table}
 
@@
 \input{graphs/ipsec-ftp-bench}
-\caption{FTP file transfer over an IPsec tunnel.}{The advantage of the hardware is a lower CPU usage, at the cost of a few percents lower throughput.
+\caption{FTP file transfer over an IPsec tunnel}{The advantage of the hardware is a lower CPU usage, at the cost of a few percents lower throughput.
 The GCM mode is also tested to discuss its implementation in hardware.
@@
 \end{figure}
+
+\begin{table}[ht]
+\center
+\small
+\begin{tabular}{l|c|c|c|c|c|c|c|c|} \cline{2-9}
+ & \multicolumn{2}{c|}{none:none} & \multicolumn{2}{c|}{aes256cbc:none} & \multicolumn{2}{c|}{aes256cbc:sha256} & \multicolumn{2}{c|}{aes256gcm} \\ \cline{2-9}
+ & Tp.  & CPU & Tp.  & CPU & Tp.  & CPU & Tp.  & CPU \\ \hline
+\multicolumn{1}{|c|}{Out-of-tunnel} & 11.39 & 7.16 & -- & -- & -- & -- & -- & -- \\ \hline
+\multicolumn{1}{|c|}{Inside tunnel} & 10.21 & 14.68 & -- & -- & -- & -- & -- & -- \\ \hline
+\multicolumn{1}{|c|}{Software} & -- & -- & 8.83 & 63.74 & 6.47 & 74.64 & 5.09 & 89.66 \\ \hline
+\multicolumn{1}{|c|}{BA414E} & -- & -- & 8.52 & 14.87 & 5.80 & 17.25 & -- & -- \\ \hline
+\end{tabular}
+\caption{FTP file transfer over an IPsec tunnel}{The throughput is in MB/s.}
+\label{tab:ipsec-ftp-bench}
+\end{table}
 

### 2015-05-27 test-proto.tex
test protocols

@@
 \subsection{ARM DS-5 Streamline}
-% Use a daemon and a driver to get system info and send them to another machine. Has almost no impact on the performance and the load of the system.
+Streamline is a toolsuite included in ARM Development Studio.
+By installing a kernel module and running a daemon on the board, it can fetch a huge amount of data without over-loading the system.
+This tool is able to gather information on the CPU utilisation, the amount of context switches, of interrupts, of memory used, and much more.
+If the applications, libraries and kernel modules have been compiled with the required debugging options, it can also build the call path of a run, giving extensive statistics on the most ressource-hungry functions.
+
+If it has close to no impact on the performance of the board, the same does not go for the hosting machine.
+A serious drawback is his memory consumption that, when combined with the ressources used by the virtual machine, bring the hosting machine to its knees.
+
+\noindent The user should also be aware that if Streamline if a very powerfull and versatile tool, it is not reliable at very high resolution.
+The data should not be analyzed in windows smaller than a millisecond.
+
+\section{Test cases}
+The aim of this work is to test the implementations under realistic situations.
+To reach this goal, three test cases will be used:
+\begin{itemize}
+	\item Lots of connections without fetching any data.
+	\item Latency of the connection. This test can also illustrate the transfer of small amount of data.
+	\item Large file transfer on an established connection.
+\end{itemize}
+
+Each of those tests will go from a raw run, without any encryption nor authentication serving as a comparison ground, gradually increasing security up to a complete encryption/authentication scheme.
+
+\noindent As AES mode CBC and SHA-2 are widely used, they will be used with a serious security paramter using keys of 256-bit.
+Some test case will also experiment some implementation of AES-GCM to compare its performance and justify the need for its offloading in hardware.
+
+For the IPsec and OpenVPN implementation, the cipher/authentication pair null/null will be used to quantify the overhead of the encapsulation.
+It should however not be forgotten that it is not to be used as a production configuration.
+As the RFC 7321~\citep[pg. 7]{rfc7321} sates for IPsec: ``Note that while authentication and encryption can each
+   be `NULL', they MUST NOT both be `NULL'".
+
+We will use two types of drivers for the BA411E: an interuption-based for OpenVPN and OpenSSH, and a polling-based for IPsec.
+
+All the tests will be conducted with the ESP protocol in tunnel mode, so that we have the worst case scenario; as AH imposes a smaller overhead, the perfomance could only be better.
 
@@
 
+\lstinputlisting[language=bash, label=list:openvpn-client-script, caption={Script starting ten clients in parallel who will stress the server.}]{stress-openvpn.sh}
+
 As this experiment can be very unstable and vastly depends on the operating system scheduling, each test case has been repeated five times to ensure stable results.
-
-\lstinputlisting[language=bash, label=list:openvpn-client-script, caption={Script starting ten clients in parallel who will stress the server.}]{stress-openvpn.sh}
 
@@
 
-For the IPsec use case, the cipher/authentication pair null/null will be used to quantify the overhead of the encapsulation.
-It should however not be forgotten that it is not to be used as a production configuration.
-As the RFC 7321~\citep[pg. 7]{rfc7321} sates: ``Note that while authentication and encryption can each
-   be `NULL', they MUST NOT both be `NULL'".
-
-We will also use two types of drivers for the BA411E: an interuption-based for OpenVPN and OpenSSH, and a polling-based for IPsec.
-
-All the tests will be conducted with the ESP protocol in tunnel mode, so that we have the worst case scenario; as AH imposes a smaller overhead, the perfomance could only be better.
-
 For OpenSSH, the commmand \texttt{scp} will be used to transfer the data securely over an SSH tunnel.
 As for OpenVPN and IPsec, a tunnel will be established beforehand, and then the simple \texttt{ftp} command will allow the client to fetch the file on the server.
+
+All the results on the throughput and CPU usage take into account only the file transfer.
+The server authentication, the connection establishement and any other securoty negociation is ignored in this particular test case.

### 2015-05-27 technical-background.tex
colors
something on the OSI model

@@
 % Present the TCP/IP layering
-The RFC 1122~\cite{rfc1122} defines the TCP/IP and OSI stack as in the table~\ref{tab:tcp-ip-stack}.
+Computer networks are described by two models: the TCP/IP stack and the OSI model, each splitting the network workflow into more or less abstract layers.
+The RFC 1122~\cite{rfc1122} defines the TCP/IP, and the international standard ISO/IEC 7498-1~\cite{ISOIEC7498} defines the OSI stack as in the table~\ref{tab:tcp-ip-stack}.
 The main difference between the two is the application layer of the TCP/IP stack which corresponds to the three upper layers of the OSI model.
@@
 	\item \texttt{INFORMATIONAL}: Exchange of management information.
-\end{itemize}\newline{}
+\end{itemize}
+~\newline{}
 

### 2015-05-27 implementation.tex
Bastien's remarks

--- 2015-05-26 implementation.tex
@@
 \chapter{Implementation}
-This chapter will present how we implemented the protocols of chapter~\ref{chap:theory} on the board.
+This chapter will present how we implemented the protocols presented in chapter~\ref{chap:theory} on board.
 The figure~\ref{fig:os-path-generic} shows the generic data flow in the operating system through the user and kernel spaces.
@@
 
-Before closing the chapter, we will present the two main IPs to which the cryptographic operations will be offloaded and the associated drivers.
+Before closing the chapter, we will present the two main IP cores to which the cryptographic operations will be offloaded and the associated drivers.
 
@@
 
-Uses regular TCP/UDP network protocols, which can be an advantage over IPsec if the ISP decides to block the latter.
+It uses regular TCP/UDP network protocols, which can be an advantage over IPsec if the ISP decides to block the latter.
 
@@
 
-Since OpenVPN aims at network security at the application level, it needs virtual network interface to which to push and from which to read the secured data.
+Since OpenVPN aims at network security at the application level, it needs virtual network interface to read and write the data.
 The \texttt{tun} adapter will be used when OpenVPN wants to establish a point-to-point connection between two client, while the \texttt{tap} device is used when there is one server acting as a bridge, hence managing lots of clients at the same time.
@@
 OpenVPN as two choices when it wants to sends its data over the internet: either encapsulating its IP packet into TCP or UDP packets.
-The problem of encapsulating IP into TCP is that inside IP packets are already TCP frames.
+The problem of encapsulating IP into TCP is that inside IP packets, there already are TCP frames.
 The reason is that IP has been designed to work on unreliable networks, and TCP includes in its standard protocols to retry and drop packets.
@@
 
-An other way to look at this workflow is to completely isolate the OpenVPN part from both hosts, and exporting it to an independent device such as a router.
+An other way to look at this workflow is to completely isolate the OpenVPN block from both hosts, and exporting it to an independent device such as a router.
 The hosts would be stripped off the virtual interface -- \texttt{tun0} would become physical, and would communicate normally over the internet, at least from their point of view.
@@
 
-\noindent The first liberty is a laste\footnote{A vulnerability note as been issued by Carnegie Mellon University Computer Emergency Response Team in early 2009 (last revision)~\cite{CERT2009}, in response to a research of University of London~\cite{Albrecht:2009} presenting a plaintext-recovering attack against SSH when CBC mode is used, but the OpenSSH update took only place in october 2014.}.
+\noindent The first customization is a response to an old vulnarability\footnote{A vulnerability note as been issued by Carnegie Mellon University Computer Emergency Response Team in early 2009 (last revision)~\cite{CERT2009}, in response to a research of University of London~\cite{Albrecht:2009} presenting a plaintext-recovering attack against SSH when CBC mode is used, but the OpenSSH update took only place in october 2014.}.
 Even if it is not recommended, the user can still enable this mode by explicitly configuring it in the \texttt{sshd\_config} options.
 
-\noindent The second liberty has been taken to prevent the user to strip himself from data authenticity and integrity.
-However, in the case of testing and benchmarking, leaving the MAC behind can be interresting, especially if it is not offloaded in hardware such as the encryption in our case.
-In order to add this feature to OpenSSH, we wrote a patch to apply on OpenSSH 6.7, available in appendix~\ref{chap:openssh-patch}.
+\noindent The second has been taken to prevent the user to strip himself from data authenticity and integrity.
+However, in the case of testing and benchmarking, skipping the MAC can be interresting, especially if it is not offloaded in hardware such as the encryption in our case.
+In order to overcome this limitation, we wrote a patch to apply on OpenSSH 6.7, available in appendix~\ref{chap:openssh-patch}.
 
@@
 
-Offers an high-level interface called EVP to be used by other applications.
+OpenSSL offers an high-level interface called EVP to be used by other applications.
 
 It is to be noted that the present work uses an implementation with all the debug flags activated.
-Figure~\ref{fig:openssl-speed-dbg-on-off} shows that if it does have an impact on the performance, it is minimal: in the worst case of the benchmark, the throughput drops only by 2.4\%.
+Figure~\ref{fig:openssl-speed-dbg-on-off} shows that if it does have an impact on the performance, it is negligible for our tests: in the worst case of the benchmark, the throughput drops only by 2.4\%.
 Moreover, the benchmark maximizes this difference by doing only OpenSSL operations.
@@
 % TODO talk about IKE and SA management
-Strongswan is a full implementation of IPsec relying on the kernel drivers for the networking part, on the crypto API for the cryptographic part, and on user space crypto libraries for the connection negociation.
-An other popular implementation is ipsec-tools, but its development lags behind modern Linux and is not up-to-date with the 3.14 Linux kernel headers, making its cross-compilation painfull.
+Strongswan is a full implementation of IPsec relying on the kernel drivers for the networking part, on the Linux Crypto API for the cryptographic part, and on user space crypto libraries for the connection negociation.
+An other popular implementation is ipsec-tools, but its development lags behind modern Linux and is not up-to-date with the 3.14 Linux kernel headers, making its cross-compilation difficult.
 %Use strongswan 5.3.0
-Strongswan as two advantages: it has a tremendous and exhaustive documentation, and its uer interface is straightforward.
-Once configured, a simple \texttt{ipsec start \&\& ipsec up <connection>} on both sides is enough to create a ready to use VPN.
+Strongswan as two advantages: it has a tremendous and exhaustive documentation, and its user interface is straightforward.
+Once configured, a simple \texttt{ipsec start \&\& ipsec up <connection>} on both sides is enough to create a ready-to-use VPN.
 
 The figure~\ref{fig:ipsec-workflow} illustrates the workflow of Alice communicating with Bob via an IPsec ESP tunnel.
-The XFRM, read ``transform", framework is implementing IPsec and handles the incomming and outgoing packets for established VPNs~\cite{rosen2014}.
+The XFRM (read ``transform") framework is implementing IPsec and handles the incomming and outgoing packets for established VPN connections~\cite{rosen2014}.
 Its name comes from the fact that the kernel transforms packet frames to incorporate IPsec security.
@@
 
-The table~\ref{tab:ip-ciphers} summarizes the ciphers supported by the two Barco Silex' IPs used in this work.
+The table~\ref{tab:ip-ciphers} summarizes the ciphers supported by the two Barco Silex' IP cores used in this work.
 
@@
 \begin{tabular}{|l|c|}\hline
-IP & Ciphers \\ \hline
-BA414E & RSA, DH, DSA, EC \\ 
-BA411E & AES modes CBC, CTR, GCM, CCM, CFB, OFB, CTS, ECB \\ \hline
+IP core & Ciphers \\ \hline
+BA414E & RSA, DH, DSA, ECDSA, ECDH \\ 
+BA411E & AES modes CBC, CTR, GCM, CCM, CFB, OFB, XTS, ECB \\ \hline
 \end{tabular}
-\caption{Summary of the ciphers supported by two of Barco Silex' IPs.}{}
+\caption{Summary of the ciphers supported by two of Barco Silex' IP cores.}{}
 \label{tab:ip-ciphers}
@@
 
+For this tests, only one instance of each IP core will be programmed on the FPGA.
+As all the tests will be single-threaded, having two or more instances of the same IP core would not improve the performance, as they could not be used in parallel anyway.
+
 \subsection{BA411E Driver}
-Takes the rightern path, which is the cleanest because the most versatile.
-By pluging the driver into the crypto API, we offer a standard kernel interface that can be used by any other kernel driver, such as the ESP driver, or user-space application via the cryptodev driver and OpenSSL engine.
+This driver takes the rightern path of figure~\ref{fig:os-path-generic-b}, which is the best because the most versatile.
+By pluging the driver into the Linux Crypto API, we offer a standard kernel interface that can be used by any other kernel driver, such as the ESP driver, or user-space application via the cryptodev driver and OpenSSL engine.
 
@@
 
-Be it IRQ or polling, the inner workings are the same and follow the exact same canvas as the default software implementation the different AES modes.
+Whichever method is used to fetch the state of the hardware, the rest of the driver follow the exact same canvas as the default software implementation of the different AES modes.
 
 \subsection{BA414E Driver and Silex engine}
-At the moment of development, there is no asymetric cryptography interface in the crypto API\footnote{A request for comment patch as been submitted to the Linux kernel mailing list~\cite{crypto-api-pk-encryption} in late April 2015, proposing a standard interface for public key encryption in the crypto API.}.
+At the moment of development, there is no asymmetric cryptography interface in the Linux Crypto API\footnote{A request for comment patch as been submitted to the Linux kernel mailing list~\cite{crypto-api-pk-encryption} in late April 2015, proposing a standard interface for public key encryption in the crypto API.}.
 The BA414E can thus not be accessed using the same path as the BA411E and a user-space driver will be needed, that is the leftern path of figure~\ref{fig:os-path-generic-b}.

### 2015-05-27 technical-background.tex
Bastien's remarks

@@
 
-This chapter will address the technical ground inherent to this work.
-First comes an overview of Linux operating systems, the distinction between user and kernel mode, and the design of device drivers.
+This chapter will address the technical ground necessary to this work.
+First comes a quick presentation of FPGAs and how they can be driven from the operating system.
 
-\noindent Follow a quick presentation of FPGAs and how they can be driven from the operating system.
+\noindent Follows an overview of Linux operating systems, the distinction between user and kernel mode, and the design of device drivers.
 
-\noindent Next comes An overview of the cryptographic field with the main primitives: message digests, symmetric ciphers and asymmetric ciphers.
+\noindent Next comes an overview of the cryptographic field with the main primitives: message digests, symmetric ciphers and asymmetric ciphers.
 
-\noindent Finally, the ground necessary to understand the inner working and challenges of some VPNs implementations will be covered with SSL/TLS and IPsec.
+\noindent Finally, the ground necessary to understand the inner working and challenges of some VPNs implementations using SSL/TLS and IPsec will be covered.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{FPGA}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+Driving from the OS: basically, they will need to share some memory.
+That memory can be directly mapped and accessed from the user-space using \texttt{/dev/mem}, or can use a direct memory access module (DMA).
+From the operating system, we build a bunch of scatterlist in the kernel-space memory, then map those pages to memory descriptor that have a physical address on the DMA.
+They can be mapped three different ways: \texttt{DMA\_BIDIRECTIONAL}, \texttt{DMA\_TO\_DEVICE} or \texttt{DMA\_FROM\_DEVICE}.
+When the CPU write something in those descriptors and synchronize them with the DMA, it does not have to care about them anymore, the DMA in now in charge to send them to the device where registers are ready to read the incomming data.
+The same goes from the device to the CPU: when the device wants to communicate data to the OS, it writes it on the DMA that will transfer them to the CPU, triggering a flag on the way to notify it.
+
++ talk about the slave/master and the two data paths: data and control.
 
@@
 
-There are two ways to get the results from a hardware device: either by using interruption, or by actively polling the device, relentlessly asking it if it finished its operations.
+There are two ways to get the status of a hardware device: either by using interruption, or by actively polling the device, relentlessly asking if it finished its operations.
 The first case is the cleanest and the most common: when the device has something to send to the driver, or if anything unexpected happened, it sends an interruption request (IRQ) to the processor, which will in turn execute the interruption routine registered by the driver~\citep[chap.
@@
 The second case is the easiest and is always guaranteed to work, but won't let go off the processor willingly, loading it at 100\%, and avoiding any other task to be executed.
-Hopefully, modern monolithic kernels such as the Linux kernel from 2.6 provide preemptive scheduling~\cite{Santhanam2003}, that is the scheduler interrupts the running task and assigns the processor ressources it used to an other one.
+Fortunately, modern monolithic kernels such as the Linux kernel from 2.6 provide preemptive scheduling~\cite{Santhanam2003}, that is the scheduler interrupts the running task and assigns the processor ressources it used to an other one.
 Hence, systems with a lot of processes in need for CPU ressources would not be stalled, but it would not change anything if the only process heavily requesting processor time is the one using the driver.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{FPGA}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-Driving from the OS: basically, they will need to share some memory.
-That memory can be directly mapped and accessed from the user-space using \texttt{/dev/mem}, or can use a direct memory access module (DMA).
-From the operating system, we build a bunch of scatterlist in the kernel-space memory, then map those pages to memory descriptor that have a physical address on the DMA.
-They can be mapped three different ways: \texttt{DMA\_BIDIRECTIONAL}, \texttt{DMA\_TO\_DEVICE} or \texttt{DMA\_FROM\_DEVICE}.
-When the CPU write something in those descriptors and synchronize them with the DMA, it does not have to care about them anymore, the DMA in now in charge to send them to the device where registers are ready to read the incomming data.
-The same goes from the device to the CPU: when the device wants to communicate data to the OS, it writes it on the DMA that will transfer them to the CPU, triggering a flag on the way to notify it.
 
@@
 
-Cryptography is the corner stone of security.
-The four main goals are the following, as defined in~\cite{Menezes1996}:
+Cryptography is the keystone of security.
+As \citet{Menezes1996} defines it: ``Cryptography is the study of mathematical techniques related to aspects of information security such as confidentiality, data integrity, entity authentication, and data origin authentication."
+
+The four main goals are:
 \begin{description}
@@
 
-In order to achieve those, four cryptographic primitives are needed: symmetric and asymmetric ciphers, message digests and digital signatures.
+In order to achieve those, four cryptographic primitives are needed: message digests, symmetric and asymmetric ciphers, and digital signatures.
 
@@
 A message digest is the result of a one-way mathematical function of a fixed size.
+For a given function, each message has a unique digest, comparable to a fingerprint.
 Those hash functions are of two types~\cite{infof405}: manipulation detection codes (MDC) to guarantee integrity and message authentication codes (MAC) to guarantee both integrity and source authentication.
@@
 
-Based on this design and adding a key to the process, the RFC 2104~\cite{rfc2104} defines a MAC:
+In order to create a MAC, we can add a secret key $k$ to the process, $H_i$ becoming:
+\[
+	H_i = f(H_{i-1}, x_i, k), \mbox{with } i \in [1,t]\\
+\]
+
+HMAC is such function $f()$, as defined in RFC 2104~\cite{rfc2104}
 \[
@@
 
-\subsection{Symetric cryptography}
+\subsection{Symmetric cryptography}
 Talk about encryption, integrity and authentication.
@@
 
-\subsection{Asymetric cryptography}
+\subsection{Asymmetric cryptography}
 % Diffie-Hellman: show the math about the shared key, maybe from the course INFOH405 or RFC2631, and point which operation will be offloaded.
-Asymetric cryptography relies on a pair of keys: one private known only to the owner of the certificate, and one public available to anyone.
-Such cryptography uses two kinds of operations: encryption using the public key of the recipient and digital signature, which is an ecnryption using the private key of the sender.
+Asymmetric cryptography relies on a pair of keys: one private known only to the owner of the certificate, and one public available to anyone.
+Such cryptography uses two kinds of operations: encryption using the public key of the recipient and digital signature, which is an encryption using the private key of the sender.
 
 \subsubsection{RSA}
-RSA is a public-key scheme proposed in 1978 by three MIT researchers who gave it their name~\cite{Rivest:1978:MOD:359340.359342}.
+RSA is a public-key scheme proposed in 1978 by three MIT researchers (who gave it their name)~\cite{Rivest:1978:MOD:359340.359342}.
 A few years later, they founded RSA Laboratories, which is now in charge of maintaining its standards, alongside many others, as the first Public-Key Cryptography Standards, \textit{aka} PKCS \#1.
@@
 	\item randomy choose two large primes $p$ and $q$;
-	\item compute the modulus $n = p q$, and consequently we have $\phi(n) = (p-1)(q-1)$, with $\phi(n)$ as the Euler function;
+	\item compute the modulus $n = p q$, and consequently we have \[\phi(n) = (p-1)(q-1)\]with $\phi(n)$ as the Euler function;
 	\item randomly choose the public exponent $e \in ]1,\phi(n)[\ s.t.\ GCD(e,\phi(n)) = 1$;
@@
 
+The scheme is based on modular artihmetic, which is secure because of the factoring problem.
+For an opponent to decrypt a ciphertext $c$ wihtout knowing the private key $(d, n)$, he would need to compute the $e^{th}$ roots of $c$, modulo $n$.
+This operation is computationally hard, since an efficient solution is yet to be found for sufficiently large key size.
+
 \subsubsection{Diffie-Hellman}
-Diffie-Hellman is a secret key exchange protocol: two parties compute a shared secret $ZZ$ that can be used as a symetric key during the following exchanges.
+Diffie-Hellman is a secret key exchange protocol: two parties compute a shared secret $ZZ$ that can be used as a symmetric key during the following exchanges.
 It uses the same kind of operation as RSA, that is modular exponentiation.
-The protocol can be one of two type (\cite{rfc2631}, \cite{Frankel:2005:SGI:2206289}):
+The protocol can be one of three types (\cite{rfc2631}, \cite{Frankel:2005:SGI:2206289}):
 \begin{itemize}
-	\item Static: the actors use their authenticated certificate to compute the shared secret.
-	\item Ephemeral: the actors create a new pair of public/private keys from which the secret key is derived.
-	\item Anonymous: same as ephemeral, but without signing anything, hence not identifying neither of the actors. This mode is not advisable since it's vulnerable to main-in-the-middle attack.
+	\item Static (DH): the actors use their authenticated certificate to compute the shared secret.
+	\item Ephemeral (EDH): the actors create a new pair of public/private keys from which the secret key is derived.
+	\item Anonymous (ADH): same as ephemeral, but without signing anything, hence not identifying neither of the actors, hence not providing any authentication.
+This mode is not advisable since it's vulnerable to main-in-the-middle attack.
 \end{itemize}
@@
 If that secret has already been used, he can decipher all data transfered during past connections.
-However, if the secret is new for every new connection, the compromission of the shared secret des not jeopardize past communications.
-This is perfect forward secrecy: using a new key to protect the past.
+However, if the secret is new for every new connection, the compromission of the shared secret does not jeopardize past communications.
+This is perfect forward secrecy: using a new key to protect the previous messages.
 
-Hereunder is the generation of an ephemeral shared secret.
+Hereunder is the generation algorithm of an ephemeral shared secret.
 For a static secret, Alice and Bob will simply use their static certificate, sparing the modular exponentiation of the ephemeral public key generation.
@@
 	\item Bob checks the signature and picks $x_b$.
-	\item Bob computes $y_a^{x_b} mod\ p = g^{x_a x_b} mod\ p = ZZ$, the shared secret to use as a premaster key from which will be derived the symetric key for further communications.
+	\item Bob computes $y_a^{x_b} mod\ p = g^{x_a x_b} mod\ p = ZZ$, the shared secret to use as a premaster key from which will be derived the symmetric key for further communications.
 	\item Bob sends $y_b = g^{x_b} mod\ p$, signing everything with his private certificate.
@@
 
-If the server is Alice, it has to do at least one signature, one signature verification and two modular exponentiations.
+If the server is Alice, it has to do at least one signature generation, one signature verification and two modular exponentiations.
 Note that the client, B in our case, could have one signature less because RFC 5246~\cite{rfc5246} leave it as an optional feature, and the server would then have one verification less.
 However, any sane configuration will have both actors signing their ephemeral public key.
-If the certificate use RSA, we end up with four modular exponentiations, which can become quite heavy computing wise for certain sizes of prime numbers.
-We will see in chapter~\ref{chap:results} that while a 1024-bit prime is easily manageable by full software implementation, hardware offloading become a necessity for 4096-bit primes.
+If the certificate use RSA, we end up with four modular exponentiations, which can become quite heavy computing wise for certain sizes of key.
+We will see in chapter~\ref{chap:results} that while a 1024-bit key size is easily manageable by full software implementation, hardware offloading become a necessity for 4096-bit key sizes.
 Moreover, 1024-bit parameter size, both RSA and Diffie-Hellman, are disallowed by the NIST recommandations since 2013~\cite{nist-sp800-131A}.
@@
 Internet & IP, ICMP, IPsec & Network & IP, ICMP, IPsec \\ \hline
-\multirow{2}{*}{Data Link} & \multirow{2}{*}{PPP, MAC, Ethernet, L2TP} & Data link & PPP, MAC, Ethernet, L2TP \\ \cline{3-4}
+\multirow{2}{*}{Link} & \multirow{2}{*}{PPP, MAC, Ethernet, L2TP} & Data link & PPP, MAC, Ethernet, L2TP \\ \cline{3-4}
  & & Physical & USB, DSL, IEEE 802.11 \\ \hline
@@
 
+\noindent L2TP (Layer 2 Tunneling Protocol) is a protocol that simply offers a tunnel without providing confidentiality~\cite{rfc3931}.
+A common use case is to combine IPsec and L2TP, as defined in RFC 3193~\cite{rfc3193}.
+
 \subsection{SSL/TLS}
@@
 % IPsec as a protocol, strongswan come in the 'implementation' chapter.
-IPsec is a modification of the IP stack in the kernel space.
+IPsec is a protocol whose implementation is a modification of the IP stack in the kernel space.
 It is a network/internet level security; it examines incomming IP packets and checks if there exists a security association with the destination, and decrypt it on-the-fly if necessary.

### 2015-05-27 test-proto.tex
Bastien's remarks

@@
 Both are linked toghether through a network capped by 100Mbps switches.
-Both stations have gigabits ethernet interface and could hence be diretly connected to each other, but in that case the communication would be limited by the I/O transfers of the storage units -- a hard drive disk in one case, an micro-SD card in the second -- on which we can not depend to set a constant throughput limitation, as it is highly influenced by the data block size and general health of the support.
+Both stations have gigabits ethernet interface and could hence be directly connected to each other, but in that case the communication would be limited by the I/O transfers of the storage units -- a hard drive disk in one case, an micro-SD card in the second -- on which we can not depend to set a constant throughput limitation, as it is highly influenced by the data block size and general health of the support.
 
@@
 
+The table~\ref{tab:software-version} summarizes the softwares used and tjeir respective version.
+
+\begin{table}
+\center
+\begin{tabular}{|l|l|} \hline
+OpenVPN & 2.3.6 \\ \hline
+OpenSSL & 1.0.2a \\ \hline
+Strongswan & 5.3.0 \\ \hline
+OpenSSH & 6.7p1 \\ \hline
+\end{tabular}
+\caption{Software used for the test}{note that OpenSSH was updated with the patch in appendix~\ref{chap:openssh-patch}.}
+\label{tab:software-version}
+\end{table}
+
 \noindent As AES mode CBC and SHA-2 are widely used, they will be used with a serious security paramter using keys of 256-bit.
@@
 \section{TLS Connections}
-This benchmark is done only with OpenVPN.
+This benchmark is done only with OpenVPN 2.3.6.
 Since there is no standard support for those operation in the kernel yet, it would not have made sense to use IPsec, since it would have had to fallback to OpenSSL, then following the same path as OpenVPN do.
-As soon as the public key operations can be plugged into the crypto API, this use case should however be tested.
+As soon as the public key operations can be plugged into the Linux Crypto API, this use case should however be tested.
 
 For this use case, the board is configured in server mode, so that it can accept connections from any client.
+The server does not connect with a specific client and waits for incomming connection requests (see listing~\ref{list:openvpn-server}).
 The virtual machine will then execute then clients in parallel using the script~\ref{list:openvpn-client-script}.
 The only option differentiating the clients is their IP address and port number.
-Otherwise, all the clients share the same basic configuration file (see listing~\ref{list:openvpn-config-client}), which tell them to renogociate a new connection every second.
+Otherwise, all the clients share the same basic configuration file (see listing~\ref{list:openvpn-client}), which tell them to renogociate a new connection every second.
 Hence, if a connection could be made with no delay and if the processes scheduling were ideal, the server would have to address 600 connections per minute.
@@
 \section{File transfer}
-The file transfered is an un compressed block of 128MB of random data generated using the following command:
+The file transfered is an uncompressed block of 128MB of random data generated using the following command:
 \begin{lstlisting}[language=bash]
   $ head -c $((1024*1024*128)) /dev/urandom > heavy.file
-\end{lstlisting}
+\end{lstlisting}~\newline{}
 
@@
 As for OpenVPN and IPsec, a tunnel will be established beforehand, and then the simple \texttt{ftp} command will allow the client to fetch the file on the server.
+In order to use custom temporary security configuration with \texttt{scp}, the following command can be used:
+\begin{lstlisting}[language=bash]
+  $ ./scp -S /path/to/patched_openssh/bin/ssh -o Ciphers=aes256-cbc -o MACs=none@barco.com root@150.158.232.241:heavy.file .
+\end{lstlisting}~\newline{}
 

### 2015-05-27 results.tex
Bastien's remarks
barco ip
updated results

@@
 
-It already shows that the connection latency is divided by 2 for low security RSA, and up to by 7.8 for higher security parameters.
+It already shows that when using the hardware, the connection latency is divided by 2 for low security RSA, and up to by 7.8 for higher security parameters.
 The figure~\ref{fig:openvpn-tls-bench} shows the number of TLS connections per minute for three RSA exponent sizes: 1024-, 2048- and 4096-bit.
@@
 \subsection{OpenVPN}
-The figure~\ref{fig:ping-benchmark-openvpn} shows the results for different payload sizes, which raw results are presented in table~\ref{tab:ping-benchmark-openvpn}
+The figures~\ref{fig:ping-benchmark-openvpn}~and~\ref{fig:ping-benchmark-openvpn-soft} shows the results for different payload sizes, which raw results are presented in table~\ref{tab:ping-benchmark-openvpn}
+
+\begin{figure}[ht]
+\input{graphs/openvpn-ping-benchmark-soft}
+\caption{OpenVN: ping average response time -- software}{Compares the response for different packet size when using a full-software implementation.
+OpenVPN adds up to 53\% extra delay when not using encryption nor authentication (none:none).}
+\label{fig:ping-benchmark-openvpn-soft}
+\end{figure}
 
@@
 \input{graphs/openvpn-ping-benchmark}
-\caption{OpenVN: ping average response time}{OpenVPN adds up to 53\% extra delay and offloading the encryption does not improve the latency.}
+\caption{OpenVN: ping average response time -- Software/Hardware}{When using encryption only, the transfers between the user space and the hardware make the hardware implementation perform worse than the software.
+Adding the authentication fills the gap by making the encryption synchronize with the authentication.}
 \label{fig:ping-benchmark-openvpn}
@@
 \multicolumn{1}{|l|}{none-none} & 2.066 & 2.561 & 4.293 & 6.117 \\ \hline
-\multicolumn{1}{|l|}{soft-cbc256-none} &  &  &  & \\ \hline
-\multicolumn{1}{|l|}{ba411e-cbc256-none} &  &  &  & \\ \hline
+\multicolumn{1}{|l|}{soft-cbc256-none} & 2.044 & 2.760 & 5.092 & 7.166\\ \hline
+\multicolumn{1}{|l|}{ba411e-cbc256-none} & 2.301 & 2.783 & 5.377 & 8.009\\ \hline
 \multicolumn{1}{|l|}{soft-cbc256-sha256} & 2.415 & 3.061 & 5.997 & 9.135 \\ \hline
@@
 
+\noindent It is interresting to note that for packets of 16000 bytes, the hardware adds an extra 12\% to the latency when using only encryption, but only 1\% when adding the authentication.
+In the second case, the synchronization between the confidentiality and the authentication slows the process down up to the point that the user/kernel space transfers are less an overhead than with confidentiality alone.
+If the MAC were to be done in hardware as well, the extra delay would probably be even larger as the data would have to be transfered to two different IP cores via two different drivers\footnote{A unifed driver is under development, gathering the encryption and authentication in the same kernel module.}.
+
 \noindent All those transfer make the offload useless in such a case in term of raw performance.
@@
 
-\subsection{IPsec}
-The figure~\ref{fig:ping-benchmark-ipsec} studies the same payload size as OpenVPN, but with an extra experimental hardware implementation of AES mode GCM.
+\subsection{IPsec}\label{sec:lantecy-ipsec}
+The figures~\ref{fig:ping-benchmark-ipsec} and \ref{fig:ping-benchmark-ipsec-soft-hard}  studies the same payload size as OpenVPN, but with an extra software implementation of AES mode GCM.
 The raw results are in table~\ref{tab:ping-benchmark-ipsec}.
@@
 \begin{figure}[ht]
+\input{graphs/ipsec-ping-benchmark-soft}
+\caption{IPsec: ping average response time -- Software}{The overhead imposed by IPsec when not using encryption nor authentication is 9\%.}
+\label{fig:ping-benchmark-ipsec-soft}
+\end{figure}
+
+\begin{figure}[ht]
+\center
+\subfloat[\label{fig:ping-benchmark-ipsec-gcm}GCM]{%
+	\input{graphs/ipsec-ping-benchmark-gcm}
+}
+\subfloat[\label{fig:ping-benchmark-ipsec}CBC]{%
 	\input{graphs/ipsec-ping-benchmark}
-\caption{IPsec: ping average response time}{for different packet sizes.}
-\label{fig:ping-benchmark-ipsec}
+}
+\caption{IPsec: ping average response time -- software/Hardware}{\newline{}(a) Processing GCM in hardware yields at least as good results as CBC.
+Comparing the CBC mode without authentication in hardware with the GCM mode in software thus makes sense.\newline{}(b) With or without authentication, the better results by the same margin.}
+\label{fig:ping-benchmark-ipsec-soft-hard}
 \end{figure}
@@
 \multicolumn{1}{|l|}{none-none} & 1.545 & 2.045 & 2.997 & 4.703 \\ \hline
-\multicolumn{1}{|l|}{soft-cbc256-none} &  &  &  &  \\ \hline
-\multicolumn{1}{|l|}{ba411e-cbc256-none} &  &  &  &  \\ \hline
+\multicolumn{1}{|l|}{soft-cbc256-none} & 1.581 & 2.134 & 3.910 & 6.426 \\ \hline
+\multicolumn{1}{|l|}{ba411e-cbc256-none} & 1.639 & 2.080 & 3.445 & 5.345 \\ \hline
 \multicolumn{1}{|l|}{soft-cbc256-sha256} & 1.645 & 2.322 & 4.762 & 7.975 \\ \hline
@@
 \multicolumn{1}{|l|}{soft-gcm256} & 1.651 & 2.388 & 5.383 & 9.241 \\ \hline
-\multicolumn{1}{|l|}{ba411e-gcm256} & 1.764 & 2.286 & -- & --  \\ \hline
+% \multicolumn{1}{|l|}{ba411e-gcm256} & 1.764 & 2.286 & -- & --  \\ \hline
 \end{tabular}
@@
 
-As for the GCM mode, the support in the BA411E driver is highly experimental and takes more of a proof of concept than a real use case.
-The current implementation could not support payload sizes larger than 1000 bytes, and at this size it is 5\% faster than the software.
-It is not much, but when compared to the 3\% improvement the BA411E offers at the same size for AES-256-CBC, we can expect an accelaration exceeding the 20\% for larger payload sizes.
-%TODO Instead of comparing with sha256, compare with none, no software holding back;
-Aside from the fact that the driver is not fully working yet, it could still be extensively improved.
-As an example, it still relies on the \texttt{seqiv} kernel module, as the software implementation does.
-Managing the generation of the IV directly in the driver would cut it loose from yet another software boulder.
+As for the GCM mode, the support in the BA411E driver is highly experimental and is not functional enough to be integrated in those results.
+However, the process of GCM packets in hardware should be at least as fast as CBC packets.
+We can thus compare the software implementation of the GCM mode with a hardware implementation of CBC only (wihtout authentication).
+The figure~\ref{fig:ping-benchmark-ipsec-gcm} shows an improved delay by up to 42\% when using the hardware.
 
@@
 \caption{Ping average comparison}{All values are for AES-256-CBC with SHA-256.
+this graph shows two results: the comparison between IPsec and OpenVPN overhead without encryption nor authentication, and the performance of IPsec and OpenVPN in software and hardware.
 Globally, IPsec yields better results and propose the most significant hardware offload.
@@
 In both cases, the hardware implementation exhibited slightly lower performance, but a CPU usage three to four times lower.
-On embedded platforms, this is as important as the raw performance, because lower CPU usage means underclocked CPU, resulting in a lower power consumption.
+On embedded platforms, this is as important as the raw performance, because lower CPU usage means underclocked CPU, resulting in a lower power consumption.\newline{}
 
@@
 With an hardware offload, we could expect not only a drastic drop of the CPU usage, but an increase of throughput as well, since it's CPU-limited in those results.
-Note that they are achieved using a C-based implementation of galois-field multiplications.
+As we already discussed with the latency results in section~\ref{sec:lantecy-ipsec}, the results of a hardware implementation of GCM should be at least as good as those of AES-CBC wihtout authentication.
+If it were the case, we would increase the throughput by at least 67\% and the decrease the CPU usage by at least 84\%.
+
+Note that the software results for the GCM are achieved using a C-based implementation of galois-field multiplications.
 As we saw in chapter~\ref{chap:theory}, modern processor designers tend to add specialized instruction sets aimed at AES-GCM enhancement.

### 2015-05-27 technical-background.tex
barco ip
FPGA theory

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The aim of this work is to study the impact of offloading cryptographic operations in hardware.
+In order to reach that goal, an FPGA will be used.
 
-Driving from the OS: basically, they will need to share some memory.
+A Field-Prgrammable Gate Array (FPGA) is an integrated circuit that can be programmed after the manufacturing, using a semiconductor intellectual property core (IP core).
+In this work, we will use some of Barco Silex's crypto IP cores.
+
+In order for the operating system to comunicate with the device, they will need to share some memory.
 That memory can be directly mapped and accessed from the user-space using \texttt{/dev/mem}, or can use a direct memory access module (DMA).
@@
 + talk about the slave/master and the two data paths: data and control.
+figure~\ref{fig:barco-ba411e-ipcore} shows a schematic view of a Barco Silex' crypto IP core
+\begin{figure}
+\includegraphics[width=\linewidth]{barco-ba411e-ipcore}
+\caption{Barco Silex' AES IP Core}{There are two channels to comunicate with the IP core: the data and control paths~\cite{barco-ba411e}.}
+\label{fig:barco-ba411e-ipcore}
+\end{figure}
+
+There are two types of interface with such a device: master or slave.
+The figure~\ref{fig:barco-ba411e-ipcore}  represent one of Barco Silex' crypto IP cores that proposes two interfaces: one of each type.
+\begin{itemize}
+	\item The data path is a master.
+That is the operating system will not have a direct access to the memory registers of the hardware devide.
+Instead, they will communicate via a DMA bus: one peer places data on the DMA and notifies the other that he can fetch it.
+	\item The control (ctrl) path is a slave. The operatin system can directly read and write on this part of the device memory.
+\end{itemize}
+
+A master interface has the advantage to offload a part of the memory management to the device, but there is a space overhead to connect the DMA as well as an extra delay to process its data.
+A slave interface is usually used for sporadic data, avoiding the overhead of a DMA.
+As an example, the data path of Barco Silex' symmetric cryptography IP core is a master, as there is a lot of data to process, whilst the same data path of their asymmetric cryptography IP core is a slave, for the operands are larger and the computation takes more time.
 

### 2015-05-27 conclusion.tex
Glossary

@@
 
-\paragraph{Registering public key verification with the crypto API}
+\paragraph{Registering public key verification with the Linux Crypto API}
 As we saw in \ref{sec:section-with-RSA-BA414}, the driver is already capable of offloading a large portion of public key operations to the IP, but only with very specific libraries at the time being -- openSSL in our case.
-The next step is to register the very same operations with the crypto API so it can be used without having to rely on a custom openSSL engine.
+The next step is to register the very same operations with the Linux Crypto API so it can be used without having to rely on a custom openSSL engine.
 This feature would require to work very closely to the linux kernel developpement.

### 2015-05-27 technical-background.tex
Glossary

@@
 
-+ talk about the slave/master and the two data paths: data and control.
-figure~\ref{fig:barco-ba411e-ipcore} shows a schematic view of a Barco Silex' crypto IP core
 \begin{figure}
@@
 	\item Static (DH): the actors use their authenticated certificate to compute the shared secret.
-	\item Ephemeral (EDH): the actors create a new pair of public/private keys from which the secret key is derived.
-	\item Anonymous (ADH): same as ephemeral, but without signing anything, hence not identifying neither of the actors, hence not providing any authentication.
+	\item Ephemeral (DHE): the actors create a new pair of public/private keys from which the secret key is derived.
+	\item Anonymous: same as ephemeral, but without signing anything, hence not identifying neither of the actors, hence not providing any authentication.
 This mode is not advisable since it's vulnerable to main-in-the-middle attack.

### 2015-05-27 test-proto.tex
Glossary

@@
 \chapter{Test protocol}\label{chap:test-protocol}
+This chapter presents the platform on which the tests will be undertaken, as well as the tests themselves.
 
@@
 \subsection{x86 host}
-The desktop host runs on Windows 7 Professional 64-bit, but a virtual machine using a Linux distribution is used for the developpement and testings.
+The desktop host runs on Windows 7 Professional 64-bit, but a virtual machine using a Linux distribution is used for the developement and testings.
 
@@
 	\item[OS] Ubuntu 12.04 LTS, kernel 3.16
-	\item[CPU] Intel Core-i5 (two logical core out of four)
+	\item[CPU] Intel Core-i5 M560, 2.67GHz (two logical core out of four)
 	\item[RAM] 1GB DDR3

### 2015-05-27 technical-background.tex
Symmetric cipher theory

@@
 \subsection{Symmetric cryptography}
-Talk about encryption, integrity and authentication.
+Consider a cryptosystem $(P,C,K,E,D)$, with $P$ and $C$ the spaces of plaintexts and ciphertexts, $E$ and $D$ the encryption and decryption algorithms and a keyspace $K$.
+\citet{Menezes1996} and \citet{infof405} define a symmetric scheme as a cryptosystem with two keys $e \in K$ and $d \in K$ respectively associated with the encryption and decryption algorithms.
+When using the key pair ($e$, $d$), it should be computationally easy to compute $d$ knowing only $e$, and $e$ from $d$.
+In the rest of this work, we will consider the most common case: $e = d$, having thus a single key shared between the peers.
 
-Look in depth into AES, NIST approved five modes~\cite{nist-sp800-38A}: CBC (Cipher Block Chaining), ECB (Electronic CodeBook), CFB (Cipher FeedBack), OFB (Output FeedBack) and CTR (Counter).
+There exists numerous cipher types, and one of the most widespread is the block cipher.
+A block cipher splits the message into blocks of fixed size and work on them separately.
+An example of such a cipher is illustrated on the figure~\ref{fig:cbc-encrypt} with the Cipher Block Chaining (CBC) mode.
+This particular mode is easy to implement and has been robust for a long time, before the rise of the padding oracle attack~\cite{vaudenay2002}.
+
+\indent An other interresting block cipher mode is the Counter (CTR).
+Each block of data is treated independently from the others (see figure~\ref{fig:ctr-encrypt}, making the scheme parallelizable and thus more interresting on hardware.
 
@@
 \includegraphics[width=\textwidth]{nist-cbc}
-\caption{CBC encryption and decryption diagram}{taken from the NIST recommendation~\cite{nist-sp800-38A}.}
-\label{fig:cbc-encrypt-decrypt}
+\caption{CBC encryption diagram}{The output of one block is used as the input of the next one. The decryption is very similar. Taken from the NIST recommendation~\cite{nist-sp800-38A}.}
+\label{fig:cbc-encrypt}
 \end{figure}
 
-AEAD => Encryption and authentication: GCM.
+\begin{figure}
+\includegraphics[width=\textwidth]{nist-ctr}
+\caption{CTR encryption diagram}{Each block is processed independently from the others. The decryption is very similar. Taken from the NIST recommendation~\cite{nist-sp800-38A}.}
+\label{fig:ctr-encrypt}
+\end{figure}
+
+Those modes can not work alone and have to be associated with a cipher (``CIPH" on figures~\ref{fig:cbc-encrypt}~and~\ref{fig:ctr-encrypt}).
+In 1997, the NIST hosted a competition to choose the successor of the Data Encryption Standard (DES).
+In 2000, Rijndael emerged victorious and was renamed Advanced Encryption Standard (AES).
+It is still nowadays the most used symmetric block cipher used in a large variety of applications.\newline{}
+
+Most applications want to use confidential and authenticated channels for their sensible communication.
+In order to do so, they need to combine a symmetric cipher and a MAC.
+
+\noindent New block cipher have emerged to answer the need for such combinations and formed the Authenticated Encryption with Associated Data (AEAD) mode of operation.
+The most widespread mode in network applications is the Galois Counter Mode (GCM, see figure~\ref{fig:gcm-encrypt}), combining the CTR block cipher mode and a GHASH function using multiplications in the Galois field.
+The advantage of this mode is that the authentication can be computed in parallel with the encryption, itself parallelizable.
+This mode can thus be highly optimized in hardware, as proved Barco Silex with their BA415 IP core reaching up to 100Gbps~\cite{barco-ba415}.
 

### 2015-05-28 technical-background.tex
theory: OS

@@
 
+Modern operating systems run on several levels of privilege.
+Under Unix, the kernel runs under the highest level and it has no restriction, whereas applications run on the lowest level.
+We call those two levels kernel space and user space, segragating the privilege and the memory between the two modes (\citet{tanenbaum2014} and \citet[chap. 2]{Corbet:2005:LDD:1209083}).
+The figure~\ref{fig:user-kernel} shows this separation.
+In order for the applications in the user space to reach the hardware, they need to go through the kernel and use the device drivers that are registered there.
+
+\begin{figure}[ht]
+\center
+\input{user-kernel}
+\caption{User and kernel space separation}{The applications and libraries from the the user space communicates with the kernel space via the system calls.}
+\label{fig:user-kernel}
+\end{figure}
+
 \subsection{Device Driver}\label{sec:theory-driver}
 % General idea of it works
-% device driver in userspace: linux userspace i/o (/dev/mem)
-Userspace I/O~\cite{koch2011}
+Device drivers allows the operating system to interact with hardware devices.
+On a Linux operating system, a driver can be loaded into the system while it is running, simply registering itself into the kernel.
+This dynamic loading and unloading offers a phenomenal modularity to the system, and consequently, a device driver is also called a kernel module.\newline{}
 
 There are two ways to get the status of a hardware device: either by using interruption, or by actively polling the device, relentlessly asking if it finished its operations.
-The first case is the cleanest and the most common: when the device has something to send to the driver, or if anything unexpected happened, it sends an interruption request (IRQ) to the processor, which will in turn execute the interruption routine registered by the driver~\citep[chap.
+The first case is the cleanest and the most common: when the device has something to send to the driver, or if anything unexpected happened, it sends an interruption request (IRQ) to the processor, which will in turn execute the interruption service routine (ISR) registered by the driver~\citep[chap.
 10]{Corbet:2005:LDD:1209083}.
-The second case is the easiest and is always guaranteed to work, but won't let go off the processor willingly, loading it at 100\%, and avoiding any other task to be executed.
+The second case is the easiest and is always guaranteed to work, but won't let go of the processor willingly, loading it at 100\%, and avoiding any other task to be executed.
 Fortunately, modern monolithic kernels such as the Linux kernel from 2.6 provide preemptive scheduling~\cite{Santhanam2003}, that is the scheduler interrupts the running task and assigns the processor ressources it used to an other one.
-Hence, systems with a lot of processes in need for CPU ressources would not be stalled, but it would not change anything if the only process heavily requesting processor time is the one using the driver.
+Hence, systems with a lot of processes in need for CPU ressources would not be stalled, but it would not change anything if the only process heavily requesting processor time is the one using the driver.\newline{}
+
+The vast majority of drivers are meant to be run in kernel space, but some use cases may require the user space to directly access the hardware.
+In those cases, one can use the Userspace I/O framework.
+
+\citet{koch2011} explains that a device driver has two roles: accessing the device memory and dealing with its interruptions.
+The device memory can already be accessed from the user space without the need for UIO via \texttt{/dev/mem}, but UIO improves this method by preventing the user space from mapping memory that does not belong to the device.
+
+\noindent As for the interruptions, the user space alone is not enough.
+When the device sends an IRQ, the system will react by executing an ISR.
+The IRQ flag however needs to be deactivated at the end of the routine, and this must be done in kernel mode.
+Hence, a small kernel module is still needed to handle this ISR.
+
+\noindent Of course, this type of driver can also actively poll the status of the device, removing the need for a kernel module as it would not rely on interruptions.
 
@@
 
-\subsection{Symmetric cryptography}
+\subsection{Symmetric cryptography}\label{sec:symmetric}
 Consider a cryptosystem $(P,C,K,E,D)$, with $P$ and $C$ the spaces of plaintexts and ciphertexts, $E$ and $D$ the encryption and decryption algorithms and a keyspace $K$.
 \citet{Menezes1996} and \citet{infof405} define a symmetric scheme as a cryptosystem with two keys $e \in K$ and $d \in K$ respectively associated with the encryption and decryption algorithms.
+Hence, for all messages $x$ in the message space $M$: $D_d(E_e(x)) = x$.
 When using the key pair ($e$, $d$), it should be computationally easy to compute $d$ knowing only $e$, and $e$ from $d$.
@@
 \subsection{Asymmetric cryptography}
-% Diffie-Hellman: show the math about the shared key, maybe from the course INFOH405 or RFC2631, and point which operation will be offloaded.
-Asymmetric cryptography relies on a pair of keys: one private known only to the owner of the certificate, and one public available to anyone.
-Such cryptography uses two kinds of operations: encryption using the public key of the recipient and digital signature, which is an encryption using the private key of the sender.
+Asymmetric cryptography relies on a pair of keys: one private known only to its owner, and one public available to anyone.
+If we use the same notation for the cryptosystem as in section~\ref{sec:symmetric}, the public key would be $e$, and the private key $d$.
+Such cryptography uses two kinds of operations: 
+\begin{itemize}
+	\item Encryption: the sender uses the public key of the recipient to encrypt the message $m$ into a ciphertext $c = E_e(m)$, and the recipient uses his private key to decrypt it: $m = D_d(m)$.
+	\item Digital signature: the sender generates the signature $s$ of the message with his private key: $s = D_d(m)$, and the recipient uses the public key of the sender to verify the signature: $m = E_e(s)$.
+	Since the sender can not deny having signed the message with his private key, this adds a non-repudiation of origin property to the scheme.
+\end{itemize}
+
+Eventhough the signature algorithms are often more efficient than the usual MACs and the key pairs key stay in use for very long period of time, asymmetric cryptography can not supplant its symmetric counterpart.
+The throughput rates of modern asymmetric schemes are still well under the symmetric schemes, plus the key sizes are much larger.
+Whereas the most efficient search for the symmetric key (ignoring the weaknesses specific to certain modes) is an exhaustive search, all public-key systems are vulnerable to more efficient attacks (\textit{e.g.} factorization or discrete logarithm).
+
+In practice, public-key cryptography is used to estalish a secure channel between two or more peers and exchange a shared secret key that will be used the encrypt the subsequent communications.
 
@@
 % First explain what a VPN is.
+\citet{Frankel:2005:SGI:2206289} define a VPN as: ``virtual network built on top of existing physical networks that can provide a
+secure communications mechanism for data and control information transmitted between networks."
 

### 2015-05-28 technical-background.tex
Theory: SSL/TLS.

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Network and VPN implementation}\label{sec:theory-network}
+\section{Network and VPN}\label{sec:theory-network}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \noindent L2TP (Layer 2 Tunneling Protocol) is a protocol that simply offers a tunnel without providing confidentiality~\cite{rfc3931}.
-A common use case is to combine IPsec and L2TP, as defined in RFC 3193~\cite{rfc3193}.
+A common use case is to combine IPsec and L2TP, as defined in RFC 3193~\cite{rfc3193}.\newline{}
+
+The present work will focus and SSL/TLS and IPsec, the figure~\ref{fig:ipsec-tls} showing their respective integration in the network stack.
+
+\begin{figure}[ht]
+\center
+\input{tls-ipsec}
+\caption{}{}
+\label{fig:tls-ipsec}
+\end{figure}
 
@@
 % Plus, SSLv3 is to be deprecated, according to a queued RFC: http://www.rfc-editor.org/internet-drafts/draft-ietf-tls-sslv3-diediedie-03.txt, and is not supported anymore by many servers since poodle.
-Application level security.
+Secure Socket Layer (SSL) and Transport Layer Security (TLS) are two protocols ensuring secure communications from the application layer, securing the transport layer.
+The last version of SSL (version 3.0) dates from 1996~\cite{rfc6101} and has been found a critical vulnerability by Google in 2014~\cite{poodle}.
+As a consequence, SSL 3.0 support is being dropped, and the IETF is working on a depracation of the standard~\cite{sslv3-diediedie-03}.
+
+\noindent TLS is the successor of SSL is at version 1.2~\cite{rfc5246}, the next version being developped by the IETF~\cite{tls13-05}.
+TLS uses X.509 certificates for the asymmetric cryptography used during the key exchange.
+Those certificates come in pair: one public and a second private, and rely on a chain of trust for the validity of the certificate.
+Usually, the term certificate is use for the public certificate, and private key for the corresponding private certificate.
+
+\noindent The figure~\ref{fig:tls-handshake} shows the messages exchanged during the establishement of a connection between two peers.
+
+\begin{figure}[ht]
+\begin{verbatim}
+Client                                             Server
+
+ClientHello                -------->
+                                              ServerHello
+                                             Certificate*
+                                       ServerKeyExchange*
+                                      CertificateRequest*
+                           <--------      ServerHelloDone
+Certificate*
+ClientKeyExchange
+CertificateVerify*
+[ChangeCipherSpec]
+Finished                   -------->
+                                       [ChangeCipherSpec]
+                           <--------             Finished
+Application Data           <------->     Application Data
+\end{verbatim}
+\caption{TLS handshake}{The * denotes messages that are sent only in some situations. Reproduce from~\cite{rfc5246}.}
+\label{fig:tls-handshake}
+\end{figure}
+
+\begin{enumerate}
+	\item Client: ClientHello. The client sends a series of specifications to the server, such as the protocol version and the supported cipher suites.
+	\item Server: ServerHello and others. The server replies by setting the protocol and the cipher suite to use.
+	It can send its own certificate and request the client's.
+	In some cases, the server can already begin the key echange protocol (\textit{e.g.} Diffie-Hellman).
+	\item Client: ClientKeyExchange and others.
+The client veryfies the validity of the server certificate, if any, and sends its own.
+If necessary, the client may have to change its cipher specification based on the constraints of the server, or request a new specification from the server.
+	The client can also participate in the key exchange protocol.
+	\item Server: Finished.
+If the client asked for a new cipher specification, the server can set a new one.
+Otherwise, the exchange is simply finished and the connection has been successfully established.
+\end{enumerate}
+
+Once the connection is established, the peers can exchange application data using the symmetric key negociated during the handshake.
 

### 2015-05-28 implementation.tex
drivers and openssl

@@
 \section{Software}
+In order to implement all the required cryptographic protocol, several softwares are needed.
+Their role is summarized in table~\ref{tab:soft-role}.
+
+\begin{table}[ht]
+\center
+\begin{tabular}{|l|l|} \hline
+OpenVPN & SSL/TLS \\ \hline
+OpenSSH & SSH \\ \hline
+OpenSSL & Cryptographic ciphers implementation \\ \hline
+Strongswan & IPsec \\ \hline
+\end{tabular}
+\caption{Role of the softwares}{Each software implements different protocols. OpenSSL is particular since it implements the cryptographic algorithms themselves and is used by the other applications.}
+\label{tab:soft-role}
+\end{table}
 
@@
 
-\subsection{OpenSSL}
+\subsection{OpenSSL}\label{sec:implem-openssl}
 
-OpenSSL offers an high-level interface called EVP to be used by other applications.
+OpenSSL is an implementation of the SSL/TLS protocols.
+Unlike OpenVPN, it does implement all the cryptographic suites supported by the protocols in C and assembly.
+OpenSSL his mainly used as a cryptographic library and offers an high-level interface called EVP to be used by other applications.
+
+OpenSSL can extend its features by supporting cyrptographic engines.
+Those engines offer custom implementation for specific ciphers.
+In a sense, a crypto engine is to OpenSSL what the crypto kernel modules are to the Linux Crypto API.
+
+\noindent A widely used engine is one of Intel's, proposing acceleration for the AES operations on its processor using the AES-NI instruction set.
+
+Two engines will be used in this work:
+\begin{itemize}
+	\item cryptodev, to gain access to the BA411E from the user space.
+	\item silex engine, a custom-made engine to access the BA414E from user space using UIO.
+\end{itemize}
 
@@
 This driver takes the rightern path of figure~\ref{fig:os-path-generic-b}, which is the best because the most versatile.
-By pluging the driver into the Linux Crypto API, we offer a standard kernel interface that can be used by any other kernel driver, such as the ESP driver, or user-space application via the cryptodev driver and OpenSSL engine.
+By pluging the driver into the Linux Crypto API, we offer a standard kernel interface that can be used by any other kernel driver, such as the ESP driver, or user-space applications via the cryptodev driver and OpenSSL engine.
 
@@
 The former is using an IRQ-based driver, whilst the latter is actively polling the hardware.
-The reason why the IRQ-based driver can not be used by kernel-space application is because the current implementation uses sleep methods that, when called on other kernel drivers, put the kernel in panic mode and require the system to reboot.
-An alternative is under development and is further discussed in the last part of this work, in section~\ref{sec:future-work}.
+The reason why the IRQ-based driver can not be used by kernel-space application is because the current implementation uses sleep methods in order to free ressources while the hardware is busy.
+The problem is that the functions that called the device driver do not support to sleep during their execution.
+Hence, when the device driver tells the scheduler to put him into sleep mode, the parent function panics and crashes the kernel, forcing the system to reboot.
+An alternative is under development and is further discussed in the last part of this work, in section~\ref{sec:future-work}. \newline{}
 
@@
 At the moment of development, there is no asymmetric cryptography interface in the Linux Crypto API\footnote{A request for comment patch as been submitted to the Linux kernel mailing list~\cite{crypto-api-pk-encryption} in late April 2015, proposing a standard interface for public key encryption in the crypto API.}.
-The BA414E can thus not be accessed using the same path as the BA411E and a user-space driver will be needed, that is the leftern path of figure~\ref{fig:os-path-generic-b}.
-As presented in section~\ref{sec:theory-driver}, we will need to rely on an UIO driver to access the device from the user space.
+
+Pressed by exhibitions and the complexity of interfacing the functions with the crypto API, Barco Silex' developmement team took the choice to access the hardware from user space.
+
+\noindent The BA414E can thus not be accessed using the same path as the BA411E and a user-space driver will be needed, that is the leftern path of figure~\ref{fig:os-path-generic-b}.
+The complete work flow thus needs three elements:
+\begin{itemize}
+	\item A device driver handling the IRQ, initializing the device and creating a simple file system that will be used from the user space.
+	\item The UIO kernel module supervising the memory access of the user space.
+	\item The custom OpenSSL silex engine that translates the requests from OpenSSL into operations processable by the IP core.
+\end{itemize}

### 2015-05-28 technical-background.tex
cryptodev

@@
 \noindent Of course, this type of driver can also actively poll the status of the device, removing the need for a kernel module as it would not rely on interruptions.
+
+\subsection{Linux Crypto API}
+The Linux Crypto API is a cryptographic framework giving access to cryptographic capabilities in the kernel space~\cite{morris2003}.
+Initially, it was developed to support IPsec operations, but it evolved into a general-purpose API.
+
+By providing such an interface, the kernel applications\footnote{Used here as a misnomer to designate the code executed in kernel space.} can used cryptographic primitives without having to rely on their concrete implementation.
+In order to add such an implementation, a kernel has to be developed and register itself with the Linux Crypto API upon its loading.
+Those kernel modules can either be software implementation of some cryptogrpahic ciphers, or a device driver that will offload those operation to a dedicated hardware device.\newline{}
+
+The user space can also gain access to the Linux Crypto API, by using an intermediary virtual device \texttt{/dev/crypto}, \textit{aka} CryptoDev~\cite{cryptodev}.
+That support can be added by registering the CryptoDev kernel module with the kernel, and then access it from the user space, usually using a cryptographic library engine (see section~\ref{sec:implem-openssl}).
 
@@
 There exist several major implementations of VPN: SSL, IPsec, L2TP and PPTP.
-The later was developped by a vendor consortium leaded by Microsoft and proposed in the RFC 2637 and will not be discussed further.
+The later was developed by a vendor consortium leaded by Microsoft and proposed in the RFC 2637 and will not be discussed further.
 
@@
 \input{tls-ipsec}
-\caption{}{}
+\caption{SSL/TLS and IPsec integration into the TCP/IP network stack}{SSL/TLS works in the application layer in order to secure the transport layer, whilst IPsec secures the Internet layer.}
 \label{fig:tls-ipsec}
@@
 
-\noindent TLS is the successor of SSL is at version 1.2~\cite{rfc5246}, the next version being developped by the IETF~\cite{tls13-05}.
+\noindent TLS is the successor of SSL is at version 1.2~\cite{rfc5246}, the next version being developed by the IETF~\cite{tls13-05}.
 TLS uses X.509 certificates for the asymmetric cryptography used during the key exchange.
@@
 Since ESP can use a null cipher, it arguably offers the same features as AH.
-The existance of AH is actually historical; when the protocol was developped in the 1990s, there were laws in the United States and other countries preventing the export of product including encryption features.
+The existance of AH is actually historical; when the protocol was developed in the 1990s, there were laws in the United States and other countries preventing the export of product including encryption features.
 Having an authentication only counterparty allowed such product to be exported without encryption capabilities.\newline{}

### 2015-05-28 abstract.tex
Bunch of stuff. AAAAAAAAAAAAAAAH
Theory: SSL/TLS.
Comments from Bastien, changed title, added part of the intro.

@@
 \selectlanguage{english}
-% \renewcommand{\abstractname}{\Large{The \lego Motorcyclist robot}}
 \addcontentsline{toc}{chapter}{\numberline{}Abstract}
@@
 	\begin{center}
-		\textbf{Implementation of high level cryptographic protocol using a SoC platform}
+		\textbf{Implementation of high level cryptographic protocols using a SoC platform}
 	\end{center}
-This is an abstract.
+This is an abstract. Yep.
 \end{abstract}
@@
 \selectlanguage{francais}
-% \renewcommand{\abstractname}{\Large{Le robot motard \lego}}
 \addcontentsline{toc}{chapter}{\numberline{}Résumé}
@@
 	\small \textit{
-	par Quentin Delhaye, Master en ingénieur civil en informatique, à finalité spécialisée, Université Libre de Bruxelles, 2014--2015.}
+	par Quentin Delhaye, Master en ingénieur civil en informatique, à finalité finalité Software and Critical Systems Design, Université Libre de Bruxelles, 2014--2015.}
 	\normalsize
 	\end{verse}
+	\begin{center}
+		\textbf{Implementation of high level cryptographic protocols using a SoC platform}
+	\end{center}
+
+Abstract, bitch !
 \end{abstract}

### 2015-05-28 conclusion.tex
Comments from Bastien, changed title, added part of the intro.

@@
 \paragraph{Cryptodev}
-If patches adding the GCM support to cryptodev have already been released, those are not compatible with Barco Silex' driver.
+If patches adding the GCM support to cryptodev have already been released, those are not compatible with OpenSSL EVP interface.
 Adapting the interface would open the GCM hardware offload to the whole user space applications park.
@@
 % This work is just the beginning of something far larger. The aim is to build a fully secured platform, from software to hardware, using maximum security at each stage.
+
+\section{Conclusion}
+
+Security is everywhere.
+But security is not cheap, security is power-hungry.

### 2015-05-28 introduction.tex
Some shit.
Comments from Bastien, changed title, added part of the intro.

@@
 % What is hardware offloading?
+In 1965, Gordon Moore, co-founder of Intel, stated that for the same footprint, the number of transistors in an integrated circuit should double every two years~\cite{moore1965}.
+It became the Moore's law, an empirical prediction that drived inovation for fifty years.
+Nowadays, it is reaching its limits, as Intel in delaying the manufacturing of its next-generation processors.
+
+The technological limitation is not the only parameter slowing the development of new ICs.
+The second law of Moore states that the cost to design, manufacture and test the new architectures doubles every four years.
+This economical growth will eventually collide with the Moore's law, stopping its progression.
+
+The domination of multi-purpose processors is reaching its limits, and even the biggest actors on the market are considering alternatives to the Moore's Law.
+
+\noindent A report of the Natural Resources Defense Council~\cite{nrdc2014} forecasts the US datacenter electrity bill to reach \$13.7 billion by 2020.
+As CPUs are largely responsible for the power consumption, it makes their design critical.
+A solution is to use a hybrid architecture, combining a general-purpose micro-processor with an integrated circuit configurable on-the-fly.
+This is a path Intel is exploring with its new Xeon product line, a processor aimed at servers, that combines a Xeon CPU with an FPGA from Altera\cite{xeon2014}.\newline{}
+
+% on its way to expand that market with the merger with Altera~\cite{2015-intel-altera-merger}.
+
+\section{Aim and purpose of this thesis}
+
+% Hardware offloading consists in using a dedicated hardware device to process data instead of the CPU.
+
+In the scope of this work, the testing platform will be the SoCFPGA SoCrates from Altera (see figure~\ref{socrates-photo}.
+This SoC combines an ARM dual-core Cortex-A9 with an Altera Cyclone V.
+Its ethernet interface will allow us to set the platform in a networked environment and test network security protocols.
+
+This work will focus on the network security and the impact of hardware offloading of cryptographic operations.
 
@@
 Several solutions exists: abandonning general purpose processors altoghether and turn towards ARM cores, or add an FPGA to the processor for example.
-Intel proposed the latter solution in 2014~\cite{intel-xeon-fpga} and is on its way to expand that market with the merger with Altera~\cite{intel-altera-merger}.
+Intel proposed the latter solution in 2014~\cite{intel-xeon-fpga} and is on its way to expand that market with the merger with Altera~\cite{2015-intel-altera-merger}.
 
-\section{Challenge}
+There a re diffirent ways to add cryptographic capabilities to an application:
+\begin{itemize}
+	\item Having the cipher built in.
+	\item Using a cryptographic library, such as OpenSSL.
+	\item Using the Linux Crypto API.
+\end{itemize}
 
-\section{Network security}
+Present the two IP cores that will be used in this work: BA411E and BA414E.
+
+\begin{figure}[ht]
+\includegraphics[width=\linewidth]{socrates-photo}
+\caption{Altera SoCrates platform}{Combines an ARM Cortex-A9 with an Altera Cyclone V.}
+\label{socrates-photo}
+\end{figure}
+
+\section{Context}
+The work presented in this thesis has been conducted under the supervision of Sébastien Rabou, in Barco Silex' office in Louvain-la-Neuve.
+
+\section{Structure}
+The present document will begin with the technical ground necessary to this work.
+After a presentation on how to drive a hardware device from the operating system and details on device driver design, we will adress the cryptography.
+Only most relevant part of cryptography will be detailed: message digests, symmetric cryptography and more specifically the CBC and GCM modes of operation, and asymmetric cryptography with an accent on RSA and Diffie-Hellman.
+
+The next chapter will present the implementation methods for the protocols theoretically introduced.
+Four applications are considered: OpenVPN, OpenSSH, OpenSSL and Strongswan.
+
+Before presenting the results of those implementations, the tests protocols will be discussed and the platforms introduced.
+Three tests will be undertaken: TLS connections, latency and file transfer.
+
+The results will address those three tests for different implementations, and the impact of hardware offloading will be analyzed and compared.
+
+Finally, some future works and directions that the project could take are gathered in the last chapter, followed by a concluding word.

### 2015-05-28 results.tex
Comments from Bastien, changed title, added part of the intro.

@@
 As for the combination of encryption and authentication, IPsec is between 13\% and 32\% faster than OpenVPN.
-IPsec loses its advantage with the increase of the payload size, the time lost when moving around the data between the user and the kernel space being less important compared to the securoty operations.
+IPsec loses its advantage with the increase of the payload size, the time lost when moving around the data between the user and the kernel space being less important compared to the security operations.
 

### 2015-05-28 technical-background.tex
Comments from Bastien, changed title, added part of the intro.

@@
 
-In order for the operating system to comunicate with the device, they will need to share some memory.
-That memory can be directly mapped and accessed from the user-space using \texttt{/dev/mem}, or can use a direct memory access module (DMA).
-From the operating system, we build a bunch of scatterlist in the kernel-space memory, then map those pages to memory descriptor that have a physical address on the DMA.
-They can be mapped three different ways: \texttt{DMA\_BIDIRECTIONAL}, \texttt{DMA\_TO\_DEVICE} or \texttt{DMA\_FROM\_DEVICE}.
-When the CPU write something in those descriptors and synchronize them with the DMA, it does not have to care about them anymore, the DMA in now in charge to send them to the device where registers are ready to read the incomming data.
-The same goes from the device to the CPU: when the device wants to communicate data to the OS, it writes it on the DMA that will transfer them to the CPU, triggering a flag on the way to notify it.
+The operating system and the device need to share some memory to communicate.
+One possibility is to directly map and access from the user-space using \texttt{/dev/mem}, and an other is to use a direct memory access module (DMA).
 
@@
 	\item The data path is a master.
-That is the operating system will not have a direct access to the memory registers of the hardware devide.
-Instead, they will communicate via a DMA bus: one peer places data on the DMA and notifies the other that he can fetch it.
-	\item The control (ctrl) path is a slave. The operatin system can directly read and write on this part of the device memory.
+That is the operating system will not have a direct access to the data registers of the hardware device.
+Instead, they will communicate via a DMA: one peer places data on the DDR memory and notifies the other that he can fetch it.
+	\item The control (ctrl) path is a slave. The operating system can directly read and write on this part of the device memory.
 \end{itemize}
 
-A master interface has the advantage to offload a part of the memory management to the device, but there is a space overhead to connect the DMA as well as an extra delay to process its data.
-A slave interface is usually used for sporadic data, avoiding the overhead of a DMA.
-As an example, the data path of Barco Silex' symmetric cryptography IP core is a master, as there is a lot of data to process, whilst the same data path of their asymmetric cryptography IP core is a slave, for the operands are larger and the computation takes more time.
+A master interface has the advantage to offload data access operations to the device, but there is a ressource and area overhead to connect the DMA as well as an extra complexity to process its data.
+A slave interface is usually used for sporadic data (\textit{e.g.} status, control flags), avoiding the overhead of a DMA.
+As an example, the data path of Barco Silex' symmetric cryptography IP core is a master, as there is a lot of data to process, whilst the same data path of their asymmetric cryptography IP core is a slave, for the operands are larger and the computation takes more time.\newline{}
+
+If a master interface with a DMA is used, the operating system buils scatterlists in the kernel-space memory, then map those pages to memory descriptor that have a physical address on the DMA.
+They can be mapped three different ways: \texttt{DMA\_BIDIRECTIONAL}, \texttt{DMA\_TO\_DEVICE} or \texttt{DMA\_FROM\_DEVICE}.
+When the CPU write something in those descriptors and synchronize them with the DMA, it does not have to care about them anymore, the DMA in now in charge to send them to the device where registers are ready to read the incomming data.
+The same goes from the device to the CPU: when the device wants to communicate data to the OS, it writes it on the DMA that will transfer them to the CPU, triggering a flag on the way to notify it.
 
@@
 
-There are two ways to get the status of a hardware device: either by using interruption, or by actively polling the device, relentlessly asking if it finished its operations.
-The first case is the cleanest and the most common: when the device has something to send to the driver, or if anything unexpected happened, it sends an interruption request (IRQ) to the processor, which will in turn execute the interruption service routine (ISR) registered by the driver~\citep[chap.
+There are two ways to get the status of a hardware device: either by using interruption, or by actively polling the status register fo the device, relentlessly asking if it finished its operations or tasks.
+The first case is the cleanest and the most common: when the hardware device has something to send to the driver, or if anything unexpected happened, it sends an interruption request (IRQ) to the processor, which will execute the interruption service routine (ISR) registered by the driver~\citep[chap.
 10]{Corbet:2005:LDD:1209083}.
-The second case is the easiest and is always guaranteed to work, but won't let go of the processor willingly, loading it at 100\%, and avoiding any other task to be executed.
-Fortunately, modern monolithic kernels such as the Linux kernel from 2.6 provide preemptive scheduling~\cite{Santhanam2003}, that is the scheduler interrupts the running task and assigns the processor ressources it used to an other one.
+The second case is the easiest and is always guaranteed to work, but will monopolize the processir, loading it at 100\%, and avoiding any other task to be executed.
+Fortunately, modern monolithic kernels such as the Linux kernel from 2.6 provide preemptive scheduling~\cite{Santhanam2003}, \textit{i.e.} the scheduler interrupts the running task and assigns the processor ressources it used to an other one.
 Hence, systems with a lot of processes in need for CPU ressources would not be stalled, but it would not change anything if the only process heavily requesting processor time is the one using the driver.\newline{}
@@
 The vast majority of drivers are meant to be run in kernel space, but some use cases may require the user space to directly access the hardware.
-In those cases, one can use the Userspace I/O framework.
+In those cases, one can use the Userspace I/O framework on a slave interface.
 
@@
 An example of such a cipher is illustrated on the figure~\ref{fig:cbc-encrypt} with the Cipher Block Chaining (CBC) mode.
-This particular mode is easy to implement and has been robust for a long time, before the rise of the padding oracle attack~\cite{vaudenay2002}.
+This particular mode is easy to implement and has been considered robust for a long time, before the rise of the padding oracle attack~\cite{vaudenay2002}.
 
 \indent An other interresting block cipher mode is the Counter (CTR).
-Each block of data is treated independently from the others (see figure~\ref{fig:ctr-encrypt}, making the scheme parallelizable and thus more interresting on hardware.
+Each block of data is treated independently from the others (see figure~\ref{fig:ctr-encrypt}), making the scheme parallelizable and thus more interresting on hardware.
 
@@
 The advantage of this mode is that the authentication can be computed in parallel with the encryption, itself parallelizable.
-This mode can thus be highly optimized in hardware, as proved Barco Silex with their BA415 IP core reaching up to 100Gbps~\cite{barco-ba415}.
+This mode can thus be highly optimized in hardware, as proved by Barco Silex with their BA415 IP core reaching up to 100Gbps~\cite{barco-ba415}.
 

### 2015-05-28 introduction.tex
intro

@@
 
+On a different scale, Altera developed an ARM based SoC FPGA platform, combining an FPGA with an ARM Cortex-A9.
+This platform allows an increase of performance, lower power consumption and reduced board size compared with a solution with an FPGA separated from the CPU.
+Such architecture is aimed at embedded and low-energy systems, where not only the performance, but also the power consumption matters.
 % on its way to expand that market with the merger with Altera~\cite{2015-intel-altera-merger}.
@@
 \section{Aim and purpose of this thesis}
-
-% Hardware offloading consists in using a dedicated hardware device to process data instead of the CPU.
-
-In the scope of this work, the testing platform will be the SoCFPGA SoCrates from Altera (see figure~\ref{socrates-photo}.
-This SoC combines an ARM dual-core Cortex-A9 with an Altera Cyclone V.
+The aim of this thesis is to implement various popular cryptographic schemes on the board, and make use of the embedded hardware device to improve the performance and decrease the ressources utilization.
+This work will focus on the network security and the impact of hardware offloading of cryptographic operations.
+The testing platform will be the SoCFPGA SoCrates from Altera (see figure~\ref{socrates-photo}), embedding an ARM dual-core Cortex-A9 with an Altera Cyclone V.
 Its ethernet interface will allow us to set the platform in a networked environment and test network security protocols.
 
-This work will focus on the network security and the impact of hardware offloading of cryptographic operations.
+% Altera and Xilinx are trying to push those SoC plateforms to the market to popularize the FPGAs.
 
-Altera and Xilinx are trying to push those SoC plateforms to the market to popularize the FPGAs.
-
-Power is critical nowadays, esapcially in datacenters.
-Several solutions exists: abandonning general purpose processors altoghether and turn towards ARM cores, or add an FPGA to the processor for example.
-Intel proposed the latter solution in 2014~\cite{intel-xeon-fpga} and is on its way to expand that market with the merger with Altera~\cite{2015-intel-altera-merger}.
-
-There a re diffirent ways to add cryptographic capabilities to an application:
-\begin{itemize}
-	\item Having the cipher built in.
-	\item Using a cryptographic library, such as OpenSSL.
-	\item Using the Linux Crypto API.
-\end{itemize}
-
-Present the two IP cores that will be used in this work: BA411E and BA414E.
+% Power is critical nowadays, esapcially in datacenters.
+% Several solutions exists: abandonning general purpose processors altoghether and turn towards ARM cores, or add an FPGA to the processor for example.
+% Intel proposed the latter solution in 2014~\cite{intel-xeon-fpga} and is on its way to expand that market with the merger with Altera~\cite{2015-intel-altera-merger}.
 
@@
 \section{Context}
-The work presented in this thesis has been conducted under the supervision of Sébastien Rabou, in Barco Silex' office in Louvain-la-Neuve.
+The work presented in this thesis has been conducted under the supervision of Sébastien Rabou, in Barco Silex' office in Louvain-la-Neuve, between November 2014 and May 2015.\newline{}
+
+% There are diffirent ways to add cryptographic capabilities to an application:
+% \begin{itemize}
+% 	\item Having the cipher built in.
+% 	\item Using a cryptographic library, such as OpenSSL.
+% 	\item Using the Linux Crypto API.
+% \end{itemize}
+
+In order to make use of the embedded FPGA, two crypto IP cores of Barco Silex will be used: the BA411E~\cite{barco-ba411e} for AES, and the BA414E~\cite{barco-ba414e} for the public key operations.
 

### 2015-05-29 conclusion.tex
Correction of results + future work

@@
 % TODO refer to a stack analysis of an ipsec file transfer, showing the most expensive instructions. If I remember correctly, it was something like the DMA transfer and DMA something freeing.
-A better alternative would be to avoid descriptors altogether by modifiying the interface with the IP so that it can use the scatterlist directly.
-We would then spare a lot of DMA mapping instruction and thus some precious cycles on the software side.
+A better alternative would be to avoid descriptors altogether by modifiying the interface with the IP so that it can use the scatterlists directly.
+We would then spare a lot of DMA mapping instructions and thus some precious cycles on the software side.
 
-As we already remarked in \ref{sec:section-where-we-see-the-CPU-usage-of-ba411e-usage-with-active-polling}, the use cases involving IPsec were conducted using a previous revision of the driver still activelly polling the IP for its results.
-A better and cleaner way to proceed is to use interruption routines, as shown in \ref{sec:section-ba411e-with-irq}.
+As we already remarked in section \ref{sec:implem-ba411e-driver}, the use cases involving IPsec were conducted using a previous revision of the driver still activelly polling the IP for its results.
+A better and cleaner way to proceed is to use interruption routines.
 However, the kernel does not support their current implementation and panics upon usage.
@@
 \paragraph{Registering public key verification with the Linux Crypto API}
-As we saw in \ref{sec:section-with-RSA-BA414}, the driver is already capable of offloading a large portion of public key operations to the IP, but only with very specific libraries at the time being -- openSSL in our case.
-The next step is to register the very same operations with the Linux Crypto API so it can be used without having to rely on a custom openSSL engine.
+The BA414E driver is already capable of offloading a large portion of public key operations to the IP core, but only with very specific libraries at the time being -- OpenSSL in our case.
+The next step is to register the very same operations with the Linux Crypto API so they can be used without having to rely on a custom OpenSSL engine.
 This feature would require to work very closely to the linux kernel developpement.
@@
 
-Such a conditional offloading would be interresting for applications involving mainly very large packets and a few periodic smaller ones, like large data transfer between two hosts on a infrastructure supporting ethernet jumbo frames\footnote{Jumbo frames have an ethernet MTU of 9000 bytes, whilst standard frames are set to 1500.} with periodic ICMP heartbeat.
+Such a conditional offloading would be interresting for applications involving mainly very large packets and a few periodic smaller ones, like large data transfer between two hosts on a infrastructure supporting ethernet jumbo frames\footnote{Jumbo frames have an ethernet MTU of 9000 bytes, whilst standard frames are set to 1500.} with periodic ICMP heartbeats.
 
@@
 \paragraph{MAC offloading}
-While the symetric and asymetric encryption ciphers IPs are usable from the operating system, the IP computing MACs does not have a usable driver yet.
+While the symmetric and asymmetric encryption ciphers IP cores are usable from the operating system, the IP core computing MACs does not have a usable driver yet.
 Wherever there is encryption, authentication is also needed.

### 2015-05-29 graphs/ftp-bench-comparison.tex
Correction of results + future work

@@
     };
-    \label{software}
+    \label{software-cpu}
 
@@
     };
-    \label{ba411e}
+    \label{ba411e-cpu}
 
@@
     };
-    \label{out-of-tunnel}%"oot" for "out of tunnel"
-\legend{software, ba411e, out-of-tunnel}
+    \label{out-of-tunnel-cpu}%"oot" for "out of tunnel"
 \end{axis}
@@
         ymajorgrids = true,
-        ylabel = {Throughtput [MB/s]},
+        ylabel = {Throughput [MB/s]},
         xlabel = {},
@@
         scaled y ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
-        legend style={at={(0.5,-0.25)}, anchor=north,legend columns=2},
+        legend style={at={(0.5,-0.15)}, anchor=north,legend columns=4},
         enlarge x limits=0.1,
@@
     };
-    \label{soft-tp}
+    \label{software}
 
@@
     };
-    \label{ba411e-tp}
+    \label{ba411e}
 
@@
     };
-    \label{oot-tp}%"tp" for "throughput"
-\legend{}
+    \label{out-of-tunnel}%"tp" for "throughput"
+\legend{software, ba411e, out-of-tunnel}
 \end{axis}

### 2015-05-29 graphs/ipsec-ftp-bench.tex
Correction of results + future work

--- 2015-05-26 graphs/ipsec-ftp-bench.tex
@@
     };
-    \label{out-of-tunnel}%"oot" for "out of tunnel"
+    \label{out-of-tunnel-cpu}%"oot" for "out of tunnel"
 
@@
     };
-    \label{inside tunnel}%"it" for "in tunnel"
-\legend{software, ba411e, out-of-tunnel, inside tunnel}
+    \label{inside tunnel-cpu}%"it" for "in tunnel"
 \end{axis}
@@
     };
-    \label{oot-tp}%"tp" for "throughput"
+    \label{out-of-tunnel}%"tp" for "throughput"
 
@@
     };
-    \label{it-tp}
-\legend{}
+    \label{inside tunnel}
+\legend{software, ba411e, out-of-tunnel, inside tunnel}
 \end{axis}

### 2015-05-29 implementation.tex
Correction of results + future work

@@
 
-OpenVPN relies on a cryptographic library, OpenSSL in our case, for all the security operations: encryption, signature, certificate management, etc.
+OpenVPN relies on a cryptographic library, OpenSSL in our case, for all the security operations: encryption, signature, certificate management, etc.\newline{}
+
+A previous study highlighted that OpenVPN would divide the throughput by two from raw to tunneled without encryption nor authentication~\cite{Hoekstra2011}.
+Such a behaviour would not be surprising, considered all the transfers and context switches the data would have to operate and provoke in an OpenVPN tunnel.
 
@@
 % connection using SSL/TLS, then AES, but without fragmentation (unlike openvpn)
+OpenSSH is an implementation of the Secure Shell (SSH) network protocol defined in RFC 4251~\cite{rfc4251}.
+SSH lies in the application layer, as does SSL/TLS.
+It is very close to SSL/TLS in its principle and can also establish secure channel between two peers, ensuring confidentiality and authentication for the communication.
+However, this work focus is on OpenVPN and the underlying SSL/TLS protocol, and as such SSH is not studied any further than a comparison use case.\newline{}
+
 OpenSSH relies on an external cryptographic library for all its security operations.
@@
 \begin{description}
-	\item[\texttt{aes\_arm}] Assembly implementation of AES. This version is optimized to use the ARMv7 instruction set.
+	\item[\texttt{aes\_arm}] Assembly implementation of AES. This version is optimized to use the ARMv7 NEON instruction set.
 	\item[\texttt{sha256-generic}] C implementation of SHA-256.
-	\item[\texttt{gfmult128}] Multiplication in $GF(2^128)$, needed by the GCM mode.
+	\item[\texttt{gfmult128}] Multiplication in $GF(2^{128})$, needed by the GCM mode.
 	\item[\texttt{ghash}] GHASH function needed by the authentication of the GCM mode.
@@
 	\item[\texttt{cbc}] C based CBC mode.
-	\item[\texttt{gcm}] C base GCM mode.
-	\item[\texttt{crypto\_null}] Null cipher. This basically does nothing on the plaintext, but it is still needed to propose the interface in the kernel.
+	\item[\texttt{gcm}] C based GCM mode.
+	\item[\texttt{crypto\_null}] Null cipher. This basically does nothing on the plaintext, but it is still needed to make the interface available in the kernel.
 	\item[\texttt{tun}] TUN/TAP network device, needed by OpenVPN.
@@
 \begin{table}[ht]
+\center
 \begin{tabular}{|l|c|}\hline
@@
 
-\subsection{BA411E Driver}
+\subsection{BA411E Driver}\label{sec:implem-ba411e-driver}
 This driver takes the rightern path of figure~\ref{fig:os-path-generic-b}, which is the best because the most versatile.

### 2015-05-29 results.tex
Correction of results + future work

@@
 It already shows that when using the hardware, the connection latency is divided by 2 for low security RSA, and up to by 7.8 for higher security parameters.
-The figure~\ref{fig:openvpn-tls-bench} shows the number of TLS connections per minute for three RSA exponent sizes: 1024-, 2048- and 4096-bit.
-The higher the exponent size, the higher the performance boost.
+The figure~\ref{fig:openvpn-tls-bench} shows the number of TLS connections per minute for three RSA key sizes: 1024-, 2048- and 4096-bit.
+The higher the key size, the higher the performance boost.
 
@@
 \noindent RSA-2048 is a much more common configuration, espacially since the NIST deprecated RSA-1024 in 2013.
-The full software implementation is visibly affected by the increase of the exponent size: the CPU usage doubles and the server processes three time less connections.
+The full software implementation is visibly affected by the increase of the key size: the CPU usage doubles and the server processes three time less connections compared to the RSA-1024 results.
 At the same time, the hardware loses less than 20\% connections for a third of the CPU usage.
@@
 
-\noindent The results obtained for RSA-4096 can be interpreted similarly to those of RSA-2048, except that the CPU usage is exactly the same for th hardware configuration.
+\noindent The results obtained for RSA-4096 can be interpreted similarly to those of RSA-2048, except that the CPU usage is exactly the same for the hardware configuration.
 One way too look at those results is to directly compare the raw performance, and the hardware can then process almost six times more connections per minute than the software.
@@
 Such interesting results, particularly regarding the CPU usage, are possible because at least 87\% of the operations are RSA and Diffie-Hellman operations, which are entirely offloaded in hardware.
-Nevertheless, OpenVPN still needs to proceed to some extra computations (such as SHA-1 integrity), and the hardware operations are not instantaneous, so the performance gain can only be that high.
+Nevertheless, OpenVPN still needs to proceed to some extra computations (such as a default SHA-1 integrity), and the hardware operations are not instantaneous, so the performance gain can only be that high.
+Having a CPU usage of less than 5\% for both RSA-2048 and -4096 in hardware shows that the device is actually the bottleneck at this point.
 
@@
 \subsection{OpenVPN}
-The figures~\ref{fig:ping-benchmark-openvpn}~and~\ref{fig:ping-benchmark-openvpn-soft} shows the results for different payload sizes, which raw results are presented in table~\ref{tab:ping-benchmark-openvpn}
+The figures~\ref{fig:ping-benchmark-openvpn-soft}~and~\ref{fig:ping-benchmark-openvpn} shows the results for different payload sizes, whose raw results are presented in table~\ref{tab:ping-benchmark-openvpn}
 
@@
 \input{graphs/openvpn-ping-benchmark-soft}
-\caption{OpenVN: ping average response time -- software}{Compares the response for different packet size when using a full-software implementation.
+\caption{OpenVPN: ping average response time -- software}{Compares the response for different packet size when using a full-software implementation.
 OpenVPN adds up to 53\% extra delay when not using encryption nor authentication (none:none).}
@@
 \multicolumn{1}{|l|}{raw} & 1.444 & 1.929 & 2.811 & 4.322 \\ \hline
-\multicolumn{1}{|l|}{none-none} & 2.066 & 2.561 & 4.293 & 6.117 \\ \hline
-\multicolumn{1}{|l|}{soft-cbc256-none} & 2.044 & 2.760 & 5.092 & 7.166\\ \hline
-\multicolumn{1}{|l|}{ba411e-cbc256-none} & 2.301 & 2.783 & 5.377 & 8.009\\ \hline
-\multicolumn{1}{|l|}{soft-cbc256-sha256} & 2.415 & 3.061 & 5.997 & 9.135 \\ \hline
-\multicolumn{1}{|l|}{ba411e-cbc256-sha256} & 2.416 & 3.052 & 5.963 & 9.207 \\ \hline
+\multicolumn{1}{|l|}{none:none} & 2.066 & 2.561 & 4.293 & 6.117 \\ \hline
+\multicolumn{1}{|l|}{soft-cbc256:none} & 2.044 & 2.760 & 5.092 & 7.166\\ \hline
+\multicolumn{1}{|l|}{ba411e-cbc256:none} & 2.301 & 2.783 & 5.377 & 8.009\\ \hline
+\multicolumn{1}{|l|}{soft-cbc256:sha256} & 2.415 & 3.061 & 5.997 & 9.135 \\ \hline
+\multicolumn{1}{|l|}{ba411e-cbc256:sha256} & 2.416 & 3.052 & 5.963 & 9.207 \\ \hline
 \end{tabular}
-\caption{OpenVN: ping average response time}{the latency is given in milliseconds.}
+\caption{OpenVN: ping average response time}{the latency is given in milliseconds.
+The key is <implementation>-<encryption>:<authentication>.
+The difference between ``raw" an ``none:none" is that the former is processed outside the tunnel, whilst the latter is processed inside the tunnel, but without encryption/authentication.}
 \label{tab:ping-benchmark-openvpn}
@@
 
-Firstly, is the impact of OpenVPN's operation on the delay: when no encryption nor authentication is used, the exchange is 33\% to 53\% longer.
-This is simply because the packet has to go through a virtual interface, hence even without the security time overhead, the latency is much longer.
+First is the impact of OpenVPN's operation on the delay: when no encryption nor authentication is used, the exchange is 33\% to 53\% longer.
+This is simply because the packet has to go through a virtual interface, hence even without the security operations time overhead, the latency is much longer.
 
 When combining encryption and authentication, the software and hardware implementation are neck and neck.
-Eventhough the MAC is not offload, the hardware implementation could have performed better.
+Eventhough the MAC is not offloaded, the hardware implementation could have performed better.
 The reason is not only all the packets have to go through a virtual interface in the kernel, the payload also has to be sent to the hardware through the kernel.
@@
 \subsection{IPsec}\label{sec:lantecy-ipsec}
-The figures~\ref{fig:ping-benchmark-ipsec} and \ref{fig:ping-benchmark-ipsec-soft-hard}  studies the same payload size as OpenVPN, but with an extra software implementation of AES mode GCM.
+The figures~\ref{fig:ping-benchmark-ipsec-soft} and \ref{fig:ping-benchmark-ipsec-soft-hard}  studies the same payload size as OpenVPN, but with an extra software implementation of AES mode GCM.
 The raw results are in table~\ref{tab:ping-benchmark-ipsec}.
@@
 \caption{IPsec: ping average response time -- software/Hardware}{\newline{}(a) Processing GCM in hardware yields at least as good results as CBC.
-Comparing the CBC mode without authentication in hardware with the GCM mode in software thus makes sense.\newline{}(b) With or without authentication, the better results by the same margin.}
+Comparing the CBC mode without authentication in hardware with the GCM mode in software thus makes sensena yields 42\% less delay for larger payload size.\newline{}(b) With or without authentication, the hardware offers better results up to 13\% less delay.}
 \label{fig:ping-benchmark-ipsec-soft-hard}
@@
 \multicolumn{1}{|l|}{raw} & 1.444 & 1.929 & 2.811 & 4.322 \\ \hline
-\multicolumn{1}{|l|}{none-none} & 1.545 & 2.045 & 2.997 & 4.703 \\ \hline
-\multicolumn{1}{|l|}{soft-cbc256-none} & 1.581 & 2.134 & 3.910 & 6.426 \\ \hline
-\multicolumn{1}{|l|}{ba411e-cbc256-none} & 1.639 & 2.080 & 3.445 & 5.345 \\ \hline
-\multicolumn{1}{|l|}{soft-cbc256-sha256} & 1.645 & 2.322 & 4.762 & 7.975 \\ \hline
-\multicolumn{1}{|l|}{ba411e-cbc256-sha256} & 1.635 & 2.246 & 4.170 & 6.929 \\ \hline
+\multicolumn{1}{|l|}{none:none} & 1.545 & 2.045 & 2.997 & 4.703 \\ \hline
+\multicolumn{1}{|l|}{soft-cbc256:none} & 1.581 & 2.134 & 3.910 & 6.426 \\ \hline
+\multicolumn{1}{|l|}{ba411e-cbc256:none} & 1.639 & 2.080 & 3.445 & 5.345 \\ \hline
+\multicolumn{1}{|l|}{soft-cbc256:sha256} & 1.645 & 2.322 & 4.762 & 7.975 \\ \hline
+\multicolumn{1}{|l|}{ba411e-cbc256:sha256} & 1.635 & 2.246 & 4.170 & 6.929 \\ \hline
 \multicolumn{1}{|l|}{soft-gcm256} & 1.651 & 2.388 & 5.383 & 9.241 \\ \hline
@@
 \end{tabular}
-\caption{IPsec: ping average response time}{the latency is given in milliseconds. Some results for the GCM mode in hardware are non-existant.}
+\caption{IPsec: ping average response time}{the latency is given in milliseconds.
+Even if there is no results for a hardware implementation of GCM, the result should have been close to those of \textit{ba411e-cbc256:none}.}
 \label{tab:ping-benchmark-ipsec}
@@
 
-When combining encryption and authentication, the hardware implementation steadily takes the advantage over the software with the encrease of the payload length.
+When combining encryption and authentication, the hardware implementation steadily takes the advantage over the software with the increase of the payload length.
 If they both have the same latency at default payload size, the hardware is up to 13\% faster for 16000 bytes payloads.
@@
 \subsection{Comparison}
-The figure~\ref{fig:ping-benchmark-comparison} summarizes the results for OpenVPN and IPsec for the most realistic use case: the combination of encryption and authentication.
+The figure~\ref{fig:ping-benchmark-comparison} summarizes the results for OpenVPN and IPsec for the most realistic use case: the combination of encryption and authentication (aes256cbc:sha256).
 
@@
 
-\noindent In hardware, as OpenVPN could not take advantage of the accelaration of the encryption, IPsec is even faster, ranging from 25\% to 33\% faster than OpenVPN.
+\noindent In hardware, as OpenVPN could not take effeciently advantage of the accelaration of the encryption, IPsec is even faster, ranging from 25\% to 33\% faster than OpenVPN.
 
@@
 \end{tabular}
-\caption{File transfer over an SSH tunnel}{The throughput is in MB/s.}
+\caption{File transfer over an SSH tunnel}{The throughput (Tp.) is in MB/s.}
 \label{tab:openssh-bench}
@@
 \caption{FTP file transfer over an OpenVPN tunnel}{The results inside the tunnel with no encryption nor authentication highlights the overhead OpenVPN adds to the transfer.
-From there, the throughput only change by a small extent.
+From there, the throughput only changes by a small extent.
 The poor hardware results also shows the heaviness of OpenVPN on hardware offloading.
@@
 \end{tabular}
-\caption{FTP file transfer over an OpenVPN tunnel}{The throughput is in MB/s.}
+\caption{FTP file transfer over an OpenVPN tunnel}{The throughput (Tp.) is in MB/s.}
 \label{tab:openvpn-ftp-bench}
@@
 
-As it was already the case for the ping benchmark, the overhead imposed by the manipulation of OpenVPN is extremely heavy: the CPU usage jumps from 7.16\% to 42.60\%, and the throughput drops by 55\%, from 11.39MB/s to 5.18MB/s.
+As it was already the case for the ping benchmark, the overhead imposed by the manipulations of OpenVPN is extremely heavy: without encryption nor authentication, the CPU usage jumps from 7.16\% outide the tunnel to 42.60\% inside, and the throughput drops by 55\%, from 11.39MB/s to 5.18MB/s.
 
-\noindent Surprisingly enough, encrypting the data only changes the throughput by 8\%, but the CPU usage is almost doubled.
-A fair interpretation would be that the encryption is no the bottleneck, the transfer bewteen the user and kernel mode, as well as the fragmentation, are.
-However for the hardware, the performance collapse to 3.35MB/s.
+\noindent Encrypting the data only changes the throughput by 8\%, but the CPU usage is almost doubled.
+Such a small decrease of performance is coherent with the results of the ping benchmark for the chunck of data size OpenVPN is supposed to process: around 1kB.
+% A fair interpretation would be that the encryption is not the bottleneck, the transfer bewteen the user and kernel mode, as well as the fragmentation, are.
+However for the hardware, the performance collapse to 3.35MB/s, 35\% less than without encryption.
 
@@
 \end{tabular}
-\caption{FTP file transfer over an IPsec tunnel}{The throughput is in MB/s.}
+\caption{FTP file transfer over an IPsec tunnel}{The throughput (Tp.) is in MB/s.}
 \label{tab:ipsec-ftp-bench}
@@
 
-\noindent An explanation could be that the operating system prevent the driver to monopolizing the CPU with its polling and preempt it regularly, effectively lowering the ressource usage, but also limiting the performance.
+\noindent An explanation could be that the operating system prevents the driver from monopolizing the CPU with its polling and preempts it regularly, effectively lowering the ressource usage, but also limiting the performance.
 
-Adding the authentication yields expected results for : an increase of CPU usage and lower throughput.
+Adding the authentication yields expected results: an increase of CPU usage and lower throughput.
 Both implementation lose 2MB/s and a CPU usage increase of 20\% and 16\%.
@@
 
-The softaware GCM results allows to open a discussion on this mode used in conjunction with IPsec.
+The softaware GCM results allow to open a discussion on this mode used in conjunction with IPsec.
 The GCM performance presented clearly shows a drop of throughput and an increase of CPU usage, illustrating the fact that those operations are hard on the software.
-With an hardware offload, we could expect not only a drastic drop of the CPU usage, but an increase of throughput as well, since it's CPU-limited in those results.
+With a hardware offload, we could expect not only a drastic drop of the CPU usage, but an increase of throughput as well, since the performance of the software is CPU-limited.
 As we already discussed with the latency results in section~\ref{sec:lantecy-ipsec}, the results of a hardware implementation of GCM should be at least as good as those of AES-CBC wihtout authentication.
@@
 Note that the software results for the GCM are achieved using a C-based implementation of galois-field multiplications.
-As we saw in chapter~\ref{chap:theory}, modern processor designers tend to add specialized instruction sets aimed at AES-GCM enhancement.
+Modern processor designers tend to add specialized instruction sets aimed at AES-GCM enhancement~\cite{intel-gcm2011}, making assembly implementation more efficient.
 Should further tests be conducted concerning IPsec paired with GCM, it would be wise to compare with an assembly implementation exploiting ARM NEON instruction set.
@@
 \input{graphs/ftp-bench-comparison}
-\caption{Comparison of file transfer methods}{Out of the three methods, OpenSSH is to be prefered is the raw performance is the objective, but IPsec offers an higher effeciency.
+\caption{Comparison of file transfer methods}{Out of the three methods, OpenSSH is to be prefered if the raw performance is the objective, but IPsec offers a higher effeciency.
 At the same time, OpenVPN is worse on all fronts.
@@
 A look at the figure~\ref{fig:openssl-speed} shows that fragmenting at small size is a bad idea, especially when the hardware is reached from the user space.
-OpenVPN sends one third of its packets at sizes that are highly ineffective, whilst IPsec manages it better and limit the small packets to less than 10\%.
+OpenVPN sends one third of its packets at sizes that are highly ineffective, whilst IPsec manages a better frafmentation and limits the small packets to less than 10\% of the total.
 
@@
 \input{graphs/openssl-speed}
-\caption{OpenSSL benchmark for AES-256-CBC}{compare a run of \texttt{openssl speed} for three implementation of AES: in software by OpenSSL (soft), in software by the standard Linux kernel module (soft-drv) and in hardware (hard).
+\caption{OpenSSL benchmark for AES-256-CBC}{compare a run of \texttt{openssl speed} for three implementations of AES: in software by OpenSSL (soft), in software by the standard Linux kernel module (soft-drv) and in hardware (hard).
 The last two have to go through cryptodev in order to be reached from the user space.}

### 2015-05-29 test-proto.tex
intro
Correction of results + future work

@@
 
-\subsection{Altera Socrates SoCFPGA}
+\subsection{EBV Socrates board}
+The board if based on the SocFPGA from Altera: a combination of ARM cores and Altera Cyclone V FPGA.
 
@@
 
-\section{TLS Connections}
+\subsection{TLS Connections}
 This benchmark is done only with OpenVPN 2.3.6.
@@
 
-\section{Response time -- latency}
+\subsection{Response time -- latency}
 This use case exchanges ICMP request of various sizes via the \texttt{ping} command.
@@
 
-\section{File transfer}
+\subsection{File transfer}
 The file transfered is an uncompressed block of 128MB of random data generated using the following command:
@@
 All the results on the throughput and CPU usage take into account only the file transfer.
-The server authentication, the connection establishement and any other securoty negociation is ignored in this particular test case.
+The server authentication, the connection establishement and any other security negociation is ignored in this particular test case.

### 2015-05-29 conclusion.tex
Conclusion

@@
 
-% This work is just the beginning of something far larger. The aim is to build a fully secured platform, from software to hardware, using maximum security at each stage.
-
 \section{Conclusion}
 
-Security is everywhere.
-But security is not cheap, security is power-hungry.
+This thesis studied the implementation of cryptographic protocols on a platform with hardware offloading, using realistic use cases.
+The analysis of the results lead to the conclusion that a user-space VPN such as OpenVPN is not a good choice when the objective is to profit from a dedicated physical device.
+Instead, if one has all the operations he wants registered in the kernel, he should turn toward IPsec.
+OpenVPN took over the market ten years ago when IPsec was not a standard feature of the Linux kernel yet and was quite complex to implement.
+Technologic inertia left OpenVPN at the top of the tree, but modern infrastructures should favor IPsec.\newline{}
+
+Security is needed everywhere.
+But security is ressource-hungry, and consequently increase the power consumption of the platform.
+An ARM core is not meant for comple cryptographic operations.
+Combining it with an FPGA like Altera did opens a new field of application for such a SoC platform.
+Embedding this SoC into a low-energy platform could add cryptogrpahic capabilities in places it was not possible before.
+
+The strength of this systems also lies into its configurability.
+It is not a closed and frozen system, the client has a plaform that can evolve and is reconfigurable according to his needs.\newline{}
+
+This work is just the beginning of something bigger.
+This constitutes a preliminary proof-of-concept for the integration of cryptographic protocols.
+The aim is to build a fully secured platform, from software to hardware, using maximum security at each stage.
+
+\begin{center}
+Security is the goal, hardware offloading is the way.
+\end{center}

### 2015-05-29 aknowledgements.tex
stuff
conclusion
FPGA stuff and new caption command.
implementation introduction
Theory and shit
Aknowledgements

@@
 \chapter*{Aknowledgements}
+
+% Talk about the internship? => "Development of verification environment for very high throughput authenticated encryption algorithm"
+% Thanks a bunch of people here:
+
+% - Frédéric Robert for his support, advices
+
+% - Sébastien Rabou for his insight and the conversations we had on the vast and exciting fields that are semiconductors and security.
+
+% - Batien Heneffe for his extensive help
+
+% Anthony Debruyn for one awesome LaTeX trick.
+
+Before diving into the thesis \textit{per se}, I would like to thank the people who have contributed in a way or an other to this work.\newline{}
+
+\noindent First of all, I am gratefull to Professor Frédéric Robert for his advice and support.\newline{}
+
+\noindent I would also like to thank Sébastien Rabou for his insight and the conversations we had on the vast and exciting fields that are semiconductors and security.\newline{}
+
+\noindent I am also thankful for the extensive help Bastien Heneffe provided for the seven months I spent at Barco Silex office.\newline{}
+
+\noindent Finally, I would like to thank Olivier L'Heureux, Florian Hannesse and Barco Silex team for their help and welcom.\newline{}
+
+\noindent And thank you to Anthony Debruyn for one awesome \LaTeX ~trick.

### 2015-05-29 abstract.tex
Correction of results + future work
Abstracts

@@
 	\small \textit{
-	by Quentin Delhaye,  Master in Computer Science and Engineering, Professional Focus, Université Libre de Bruxelles, 2014--2015.}
+	by Quentin Delhaye,  Master in Computer Science and Engineering, Professional Focus in Software and Critical Systems Design, Université Libre de Bruxelles, 2014--2015.}
 	\normalsize
@@
 	\end{center}
-This is an abstract. Yep.
+The aim of this thesis is to study the impact on the performance and ressource usage of the implementation of cryptographic protocols on a SoC platform.
+The platform consist of the combination of an ARM core with an Altera FPGA, on which Barco Silex' crypto IP cores are programmed.
+
+The focus is set on specific cryptographic schemes: AES and its CBC mode of operation, as well as RSA and Diffie-Hellman, all being supported in hardware.
+The GCM mode of AES is also part of the results of the software implementation. Doing so opens the discussion on the performance gain if the device driver could use the hardware.
+The two main network protocols implemented to use those operations are TLS/SSL and IPsec.
+Several applications are considered for the implementation, among which OpenVPN, OpenSSL, OpenSSH and Strongswan.
+Implementing those on the platform allows realistic benchmarks on TLS connections, latency and file transfer.
+
+The results are unequivocal: the overhead added by OpenVPN not only delivers poor performance (56\% drop without encryption nor authentication), but also do not benefit from the hardware, yielding even worse results when offloading.
+With IPsec, on the other hand, the hardware uses four times less CPU than the software for similar performance on file transfer.
+As for the TLS connections, the hardware is able to process up to 7.8 times more connections than the software at high security parameters, and for 19 times less CPU utilization.
+
+% Network security, SSL/TLS and IPsec compared, different ways to offload cryptographic operations from the user and kernel space of the operating system.
+~\newline{}
+\textbf{Keywords:} \textit{Cryptography, network security, hardware offloading, SoC platform, IPsec}
 \end{abstract}
@@
 	\end{center}
+Le but de ce mémoire est d'étudier l'implémentation de protocol cryptographiques sur une plateforme SoC et leur impact sur les performances et l'utilisation des ressources.
+La plateforme consiste en la combinaison d'un cœur ARM avec un FPGA d'Altera sur lequel sont programmées des IP cores de Barco Silex.
 
-Abstract, bitch !
+L'accent est mis sur certains algorithmes cryptographiques~: AES et son mode CBC, ainsi que RSA et Diffie-Hellman, tous étant supportés par le hardware.
+Le mode GCM d'AES fait aussi partie des résultats des implémentations software, ouvrant ce faisant la discussion sur le gain de performance qu'apporterait un pilote tirant parti du hardware.
+Les deux principaux protocoles réseaux utilisant ces opérations qui sont implémentés sont TLS/SSL et IPsec.
+Plusieurs application sont utilisées pour implémenter ces protocoles~: OpenVPN, OpenSSH, OpenSSL et Strongswan.
+Leur implémentation permet d'effectuer des tests réalistes de connexions TLS, de latence et de transfert de fichier.
+
+Les résultats sont sans appel~: le coût imposé par OpenVPN non seulement diminue les performance (56~\% de perte sans chiffrement ni authentification), mais en plus il n'exploite pas le hardware, donnant des résultats encore plus mauvais quand les opérations sont déchargées sur le périphérique.
+En revanche, IPsec combiné au hardware utilise quatre fois moins de CPU que le software pour des performances similaires de transfert de fichier.
+Concernant les connections TLS, le hardware est capable de traiter jusqu'à 7.8 fois plus de connexions par minute que le software avec des paramètre de sécurité élevés, et ce en utilisant 19 fois moins de CPU.
+
+~\newline{}
+\textbf{Mots-clés:} \textit{Cryptographie, sécurité réseau, hardware offloading, plaeforme SoC, IPsec}
 \end{abstract}

### 2015-05-29 article/graphs/ftp-bench-comparison.tex
article

@@
         ymin=0, ymax=100,
-        symbolic x coords={oot, openSSH, OpenVPN, IPSec},
+        symbolic x coords={Normal, openSSH, OpenVPN, IPSec},
         xtick = data,
@@
     coordinates {
-        (oot, 0)
+        (Normal, 0)
         (openSSH, 89.29)
@@
     coordinates {
-        (oot, 0)
+        (Normal, 0)
         (openSSH, 68.29)
@@
     };
-    \label{hardware}
+    \label{hardware-cpu}
 
@@
     coordinates {
-        (oot, 7.16)
+        (Normal, 7.16)
         (openSSH, 0)
@@
     };
-    \label{oot}%"oot" for "out of tunnel"
-\legend{software, hardware, oot}
+    \label{Normal-cpu}%"Normal" for "out of tunnel"
 \end{axis}
@@
         ymin=0, ymax=12,
-        symbolic x coords={oot, openSSH, OpenVPN, IPSec},
+        symbolic x coords={Normal, openSSH, OpenVPN, IPSec},
         xtick = data,
@@
         % legend style={at={(0.5,-0.25)}, anchor=north,legend columns=2},
+        legend style={at={(0.5,-0.45)}, anchor=north,legend columns=3},
         enlarge x limits=0.1,
+        nodes near coords,
+        enlarge y limits={value=0.4,upper},
+        every node near coord/.append style={rotate=90, anchor=west},%This needs to be after 'ybar'
     ]
@@
     coordinates {
-        (oot, 0)
+        (Normal, 0)
         (openSSH, 8.19)
@@
     coordinates {
-        (oot, 0)
+        (Normal, 0)
         (openSSH, 10.39)
@@
     };
-    \label{hardware-tp}
+    \label{hardware}
 
@@
     coordinates {
-        (oot, 11.39)
+        (Normal, 11.39)
         (openSSH, 0)
@@
     };
-    \label{oot-tp}%"tp" for "throughput"
-\legend{}
+    \label{Normal}%"tp" for "throughput"
+\legend{software, hardware, Normal}
 \end{axis}

### 2015-05-29 article/graphs/openvpn-tls-bench.tex
article

@@
         ymin=0, ymax=100,
-        symbolic x coords={RSA-1024, RSA-2048, RSA-4096},
+        symbolic x coords={Low, Medium, High},
         xtick = data,
@@
     coordinates {
-        (RSA-1024, 40.32)
-        (RSA-2048, 92.14)
-        (RSA-4096, 81.97)
+        (Low, 40.32)
+        (Medium, 92.14)
+        (High, 81.97)
     };
-    \label{software}
+    \label{software-cpu}
 
@@
     coordinates {
-        (RSA-1024, 13.29)
-        (RSA-2048, 4.82)
-        (RSA-4096, 4.78)
+        (Low, 13.29)
+        (Medium, 4.82)
+        (High, 4.78)
     };
-    \label{hardware}
-\legend{software, hardware}
+    \label{hardware-cpu}
 \end{axis}
@@
         ymajorgrids = true,
-        ylabel = {Throughtput [MB/s]},
+        ylabel = {Connections per minute},
         xlabel = {},
         ymin=0,
-        symbolic x coords={RSA-1024, RSA-2048, RSA-4096},
+        symbolic x coords={Low, Medium, High},
         xtick = data,
@@
         % legend style={at={(0.5,-0.25)}, anchor=north,legend columns=2},
+        legend style={at={(0.5,-0.20)}, anchor=north,legend columns=2},
         enlarge x limits=0.4,
+        nodes near coords,
+        enlarge y limits={value=0.45,upper},
+        every node near coord/.append style={rotate=90, anchor=west},%This needs to be after 'ybar'
     ]
@@
     coordinates {
-        (RSA-1024, 445.4)
-        (RSA-2048, 155.6)
-        (RSA-4096, 19.6)
+        (Low, 445.4)
+        (Medium, 155.6)
+        (High, 19.6)
     };
-    \label{soft-tp}
+    \label{software}
 
@@
     coordinates {
-        (RSA-1024, 509.3)
-        (RSA-2048, 420.9)
-        (RSA-4096, 115.5)
+        (Low, 509.3)
+        (Medium, 420.9)
+        (High, 115.5)
     };
-    \label{hardware-tp}
-\legend{}
+    \label{hardware}
+\legend{software, hardware}
 \end{axis}

### 2015-05-29 abstract.tex
corrections

@@
 The aim of this thesis is to study the impact on the performance and ressource usage of the implementation of cryptographic protocols on a SoC platform.
-The platform consist of the combination of an ARM core with an Altera FPGA, on which Barco Silex' crypto IP cores are programmed.
+The platform consists of the combination of an ARM core with an Altera FPGA, on which Barco Silex' crypto IP cores are programmed.
 
@@
 	\small \textit{
-	par Quentin Delhaye, Master en ingénieur civil en informatique, à finalité finalité Software and Critical Systems Design, Université Libre de Bruxelles, 2014--2015.}
+	par Quentin Delhaye, Master en ingénieur civil en informatique, à finalité Software and Critical Systems Design, Université Libre de Bruxelles, 2014--2015.}
 	\normalsize
@@
 	\end{center}
-Le but de ce mémoire est d'étudier l'implémentation de protocol cryptographiques sur une plateforme SoC et leur impact sur les performances et l'utilisation des ressources.
+Le but de ce mémoire est d'étudier l'implémentation de protocoles cryptographiques sur une plateforme SoC et leur impact sur les performances et l'utilisation des ressources.
 La plateforme consiste en la combinaison d'un cœur ARM avec un FPGA d'Altera sur lequel sont programmées des IP cores de Barco Silex.
@@
 En revanche, IPsec combiné au hardware utilise quatre fois moins de CPU que le software pour des performances similaires de transfert de fichier.
-Concernant les connections TLS, le hardware est capable de traiter jusqu'à 7.8 fois plus de connexions par minute que le software avec des paramètre de sécurité élevés, et ce en utilisant 19 fois moins de CPU.
+Concernant les connections TLS, le hardware est capable de traiter jusqu'à 5.89 fois plus de connexions par minute que le software avec des paramètre de sécurité élevés, et ce en utilisant 19 fois moins de CPU.
 

### 2015-05-29 conclusion.tex
corrections

@@
 
-Although the present work presents some promissing results, the implementation can certainly be improved in several ways and some further experiments should be conducted.
+Although the present work presents some promising results, the implementation can certainly be improved in several ways and some further experiments should be conducted.
 
 \paragraph{Driver improvement}
-The driver of the BA411E can be made less ressources hungry by improving the initialisation of the descriptors and their linking, but the gain would not be significant enough to justify the time investment at this point.
+The driver of the BA411E can be made less resources hungry by improving the initialisation of the descriptors and their linking, but the gain would not be significant enough to justify the time investment at this point.
 % TODO refer to a stack analysis of an ipsec file transfer, showing the most expensive instructions. If I remember correctly, it was something like the DMA transfer and DMA something freeing.
-A better alternative would be to avoid descriptors altogether by modifiying the interface with the IP so that it can use the scatterlists directly.
+A better alternative would be to avoid descriptors altogether by modifying the interface with the IP so that it can use the scatterlists directly.
 We would then spare a lot of DMA mapping instructions and thus some precious cycles on the software side.
 
-As we already remarked in section \ref{sec:implem-ba411e-driver}, the use cases involving IPsec were conducted using a previous revision of the driver still activelly polling the IP for its results.
+As we already remarked in section \ref{sec:implem-ba411e-driver}, the use cases involving IPsec were conducted using a previous revision of the driver still actively polling the IP for its results.
 A better and cleaner way to proceed is to use interruption routines.
@@
 The next step is to register the very same operations with the Linux Crypto API so they can be used without having to rely on a custom OpenSSL engine.
-This feature would require to work very closely to the linux kernel developpement.
+This feature would require to work very closely to the linux kernel development.
 Indeed, if the signature verification using public key cryptography has been available in the kernel since 2013, a public key encryption API has only been proposed in late April 2015~\cite{crypto-api-pk-encryption} and is still under request for comments.
@@
 %TODO at least add a pseudo-code to illustrate this shit.
-Using this breakeven point, one could patch the cryptodev cryptodev engine of OpenSSL to branch on the packet size, using the hardware if the packet is large enough, or fallback on the default software implementation otherwise.
+Using this break even point, one could patch the cryptodev cryptodev engine of OpenSSL to branch on the packet size, using the hardware if the packet is large enough, or fallback on the default software implementation otherwise.
 %openssl-1.0.2a/crypto/engine/eng_cryptodev.c
@@
 	% \item if we always branch to the fallback functions, the time overhead due to crytodev would lower the performance; %No, we don't leave the user space, the overhead is minimum.
-	\item as the breakeven point is around 1024kB, the performance for a network application would be very close to those of a full software implementation, knowing that the ethernet frame size, the MTU, is set by default at 1500 bytes.
+	\item as the break even point is around 1024kB, the performance for a network application would be very close to those of a full software implementation, knowing that the ethernet frame size, the MTU, is set by default at 1500 bytes.
 \end{itemize}
 
-Such a conditional offloading would be interresting for applications involving mainly very large packets and a few periodic smaller ones, like large data transfer between two hosts on a infrastructure supporting ethernet jumbo frames\footnote{Jumbo frames have an ethernet MTU of 9000 bytes, whilst standard frames are set to 1500.} with periodic ICMP heartbeats.
+Such a conditional offloading would be interesting for applications involving mainly very large packets and a few periodic smaller ones, like large data transfer between two hosts on a infrastructure supporting ethernet jumbo frames\footnote{Jumbo frames have an ethernet MTU of 9000 bytes, whilst standard frames are set to 1500.} with periodic ICMP heartbeats.
 
@@
 
-However, one library definitely worth to keep an eye on is mbed TLS, formally known as PolarSSL, recently bought by ARM~\cite{2015-arm-buy-polarssl}.
+However, one library definitely worth to keep an eye on is mbed TLS, formerly known as PolarSSL, recently bought by ARM~\cite{2015-arm-buy-polarssl}.
 We can expect the future releases of this library to be more optimized for ARM platforms, and maybe the software footprint and overhead to be reduced.
@@
 OpenVPN took over the market ten years ago when IPsec was not a standard feature of the Linux kernel yet and was quite complex to implement.
-Technologic inertia left OpenVPN at the top of the tree, but modern infrastructures should favor IPsec.\newline{}
+Technological inertia left OpenVPN at the top of the tree, but modern infrastructures should favour IPsec.\newline{}
 
 Security is needed everywhere.
-But security is ressource-hungry, and consequently increase the power consumption of the platform.
-An ARM core is not meant for comple cryptographic operations.
-Combining it with an FPGA like Altera did opens a new field of application for such a SoC platform.
-Embedding this SoC into a low-energy platform could add cryptogrpahic capabilities in places it was not possible before.
+But security is resource-hungry, and consequently increase the power consumption of the platform.
+An ARM core is not meant for complex cryptographic operations.
+Combining it with an FPGA like Altera did open a new field of application for such a SoC platform.
+Embedding this SoC into a low-energy platform could add cryptographic capabilities in places it was not possible before.
 
-The strength of this systems also lies into its configurability.
-It is not a closed and frozen system, the client has a plaform that can evolve and is reconfigurable according to his needs.\newline{}
+The strength of this systems also lies into its reconfigurability.
+It is not a closed and frozen system, the client has a platform that can evolve and is reconfigurable according to his needs.\newline{}
 

### 2015-05-29 implementation.tex
corrections

@@
 \section{Software}
-In order to implement all the required cryptographic protocol, several softwares are needed.
+In order to implement all the required cryptographic protocol, several software are needed.
 Their role is summarized in table~\ref{tab:soft-role}.
@@
 \end{tabular}
-\caption{Role of the softwares}{Each software implements different protocols. OpenSSL is particular since it implements the cryptographic algorithms themselves and is used by the other applications.}
+\caption{Role of the software}{Each software implements different protocols. OpenSSL is particular since it implements the cryptographic algorithms themselves and is used by the other applications.}
 \label{tab:soft-role}
@@
 Since OpenVPN aims at network security at the application level, it needs virtual network interface to read and write the data.
-The \texttt{tun} adapter will be used when OpenVPN wants to establish a point-to-point connection between two client, while the \texttt{tap} device is used when there is one server acting as a bridge, hence managing lots of clients at the same time.
+The \texttt{tun} adapter will be used when OpenVPN wants to establish a point-to-point connection between two clients, while the \texttt{tap} device is used when there is one server acting as a bridge, hence managing lots of clients at the same time.
 
-OpenVPN as two choices when it wants to sends its data over the internet: either encapsulating its IP packet into TCP or UDP packets.
+OpenVPN as two choices when it wants to send its data over the internet: either encapsulating its IP packet into TCP or UDP packets.
 The problem of encapsulating IP into TCP is that inside IP packets, there already are TCP frames.
@@
 UDP is the unreliable counterparty of TCP, offering a better alternative for encapsulation.
-If we refer to the figure~\ref{fig:openvpn-packet-flow}, \texttt{iperf} would send its regular \texttt{IP(TCP)} packet over the virtual \texttt{tun} device, which will then compress, fragment and encrypt the frame before encaspulating it into a new UDP packet that will finally be sent through the physical \texttt{eth0} interface.
+If we refer to the figure~\ref{fig:openvpn-packet-flow}, \texttt{iperf} would send its regular \texttt{IP(TCP)} packet over the virtual \texttt{tun} device, which will then compress, fragment and encrypt the frame before encapsulating it into a new UDP packet that will finally be sent through the physical \texttt{eth0} interface.
 
@@
 
-\noindent At this point, it is important to note that if the figure~\ref{fig:openvpn-packet-flow} shows a workflow in which the encryption taking place before the fragmentation, a quick look at the source code proves the opposite.
-A sample of the said code has been writen down in the listing~\ref{list:openvpn-workflow}.
+\noindent At this point, it is important to note that if the figure~\ref{fig:openvpn-packet-flow} shows a work flow in which the encryption taking place before the fragmentation, a quick look at the source code proves the opposite.
+A sample of the said code has been written down in the listing~\ref{list:openvpn-workflow}.
 
@@
 \begin{lstlisting}[caption={openvpn compress then encrypt -- sample from \texttt{forward.c}.
-It clearly shows that the order of operations in the packet workflow is compression, then fragmentation and finally encryption.}, label=list:openvpn-workflow, float]% The 'float' option makes the listing unbreakable.
+It clearly shows that the order of operations in the packet work flow is compression, then fragmentation and finally encryption.}, label=list:openvpn-workflow, float]% The 'float' option makes the listing unbreakable.
 /* Compress, fragment, encrypt and HMAC-sign an outgoing packet. */
@@
 
-An other way to look at this workflow is to completely isolate the OpenVPN block from both hosts, and exporting it to an independent device such as a router.
+Another way to look at this workflow is to completely isolate the OpenVPN block from both hosts, and exporting it to an independent device such as a router.
 The hosts would be stripped off the virtual interface -- \texttt{tun0} would become physical, and would communicate normally over the internet, at least from their point of view.
-All their communications would be routed by the router of their local area network, incomming on a physical \texttt{tun0} interface and outputed to the internet on the \texttt{eth0} interface.
-Such implementation are used when an adminstrator wishes to secure a whole local area network without bloating each client with a VPN application.
+All their communications would be routed by the router of their local area network, incoming on a physical \texttt{tun0} interface and output to the internet on the \texttt{eth0} interface.
+Such implementation are used when an administrator wishes to secure a whole local area network without bloating each client with a VPN application.
 In this case, there exist router firmware packages running OpenVPN, such as DD-WRT and OpenWRT.
@@
 
-A previous study highlighted that OpenVPN would divide the throughput by two from raw to tunneled without encryption nor authentication~\cite{Hoekstra2011}.
+A previous study highlighted that OpenVPN would divide the throughput by two from raw to tunnelled without encryption nor authentication~\cite{Hoekstra2011}.
 Such a behaviour would not be surprising, considered all the transfers and context switches the data would have to operate and provoke in an OpenVPN tunnel.
 
-\noindent Note that if one wants to use the \texttt{none} cipher with a version of OpenVPN ealier than 2.3.7, he would have to update it using a community patch~\cite{openvpn-patch-none}.
+\noindent Note that if one wants to use the \texttt{none} cipher with a version of OpenVPN earlier than 2.3.7, he would have to update it using a community patch~\cite{openvpn-patch-none}.
 
@@
 OpenSSH relies on an external cryptographic library for all its security operations.
-Up until the version 6.7 published in october 2014, it had to be compiled against OpenSSL.
-However, after the infamous security vulnerability heartbleed in april 2014, the developpers took a step to move towards LibreSSL, a fork of OpenSSL managed by OpenBSD  developers.
+Up until the version 6.7 published in October 2014, it had to be compiled against OpenSSL.
+However, after the infamous security vulnerability heartbleed in April 2014, the developers took a step to move towards LibreSSL, a fork of OpenSSL managed by OpenBSD  developers.
 Still, there is no official support for any other cryptographic library.
@@
 
-\noindent The first customization is a response to an old vulnarability\footnote{A vulnerability note as been issued by Carnegie Mellon University Computer Emergency Response Team in early 2009 (last revision)~\cite{CERT2009}, in response to a research of University of London~\cite{Albrecht:2009} presenting a plaintext-recovering attack against SSH when CBC mode is used, but the OpenSSH update took only place in october 2014.}.
+\noindent The first customization is a response to an old vulnarability\footnote{A vulnerability note as been issued by Carnegie Mellon University Computer Emergency Response Team in early 2009 (last revision)~\cite{CERT2009}, in response to a research of University of London~\cite{Albrecht:2009} presenting a plaintext-recovering attack against SSH when CBC mode is used, but the OpenSSH update took only place in October 2014.}.
 Even if it is not recommended, the user can still enable this mode by explicitly configuring it in the \texttt{sshd\_config} options.
@@
 \noindent The second has been taken to prevent the user to strip himself from data authenticity and integrity.
-However, in the case of testing and benchmarking, skipping the MAC can be interresting, especially if it is not offloaded in hardware such as the encryption in our case.
+However, in the case of testing and benchmarking, skipping the MAC can be interesting, especially if it is not offloaded in hardware such as the encryption in our case.
 In order to overcome this limitation, we wrote a patch to apply on OpenSSH 6.7, available in appendix~\ref{chap:openssh-patch}.
@@
 
-OpenSSL can extend its features by supporting cyrptographic engines.
+OpenSSL can extend its features by supporting cryptographic engines.
 Those engines offer custom implementation for specific ciphers.
@@
 
-% Talk about the assembly implementation, maybe show the difference in an appendice with the C implementation. Not /that/ interresting.
+% Talk about the assembly implementation, maybe show the difference in an appendice with the C implementation. Not /that/ interesting.
 
@@
 % TODO talk about IKE and SA management
-Strongswan is a full implementation of IPsec relying on the kernel drivers for the networking part, on the Linux Crypto API for the cryptographic part, and on user space crypto libraries for the connection negociation.
-An other popular implementation is ipsec-tools, but its development lags behind modern Linux and is not up-to-date with the 3.14 Linux kernel headers, making its cross-compilation difficult.
+Strongswan is a full implementation of IPsec relying on the kernel drivers for the networking part, on the Linux Crypto API for the cryptographic part, and on user space crypto libraries for the connection negotiation.
+Another popular implementation is ipsec-tools, but its development lags behind modern Linux and is not up-to-date with the 3.14 Linux kernel headers, making its cross-compilation difficult.
 %Use strongswan 5.3.0
@@
 
-The figure~\ref{fig:ipsec-workflow} illustrates the workflow of Alice communicating with Bob via an IPsec ESP tunnel.
-The XFRM (read ``transform") framework is implementing IPsec and handles the incomming and outgoing packets for established VPN connections~\cite{rosen2014}.
+The figure~\ref{fig:ipsec-workflow} illustrates the work flow of Alice communicating with Bob via an IPsec ESP tunnel.
+The XFRM (read ``transform") framework is implementing IPsec and handles the incoming and outgoing packets for established VPN connections~\cite{rosen2014}.
 Its name comes from the fact that the kernel transforms packet frames to incorporate IPsec security.
@@
 We can also clearly see one of the main advantages of IPsec: it works in the kernel space.
-Since it does not require a virtual network interface like OpenVPN, the only transfer between the user/kernel space happens when the former wishes to send a packet on the network, passing it to the later -- or \textit{vice versa} for incomming packets.
+Since it does not require a virtual network interface like OpenVPN, the only transfer between the user/kernel space happens when the former wishes to send a packet on the network, passing it to the later -- or \textit{vice versa} for incoming packets.
 
@@
 }
-\caption{IPsec user/kernel space workflow, using \texttt{ping} as a test case.}{}
+\caption{IPsec user/kernel space work flow, using \texttt{ping} as a test case.}{}
 \label{fig:ipsec-workflow}
@@
 	\item[\texttt{cryptodev}] Creates \texttt{/dev/crypto}, giving access to the crypto API in the user space.
-	\item[\texttt{uio}] User-space I/O driver, allowing to access the hardware memory from the user space, needed by our implementation of the BA414E driver.
+	\item[\texttt{uio}] User-space I/O driver, allowing accessing the hardware memory from the user space, needed by our implementation of the BA414E driver.
 \end{description}
@@
 \subsection{BA411E Driver}\label{sec:implem-ba411e-driver}
-This driver takes the rightern path of figure~\ref{fig:os-path-generic-b}, which is the best because the most versatile.
-By pluging the driver into the Linux Crypto API, we offer a standard kernel interface that can be used by any other kernel driver, such as the ESP driver, or user-space applications via the cryptodev driver and OpenSSL engine.
+This driver takes the righten path of figure~\ref{fig:os-path-generic-b}, which is the best because the most versatile.
+By pluging the driver in the Linux Crypto API, we offer a standard kernel interface that can be used by any other kernel driver, such as the ESP driver, or user-space applications via the cryptodev driver and OpenSSL engine.
 
@@
 
-Pressed by exhibitions and the complexity of interfacing the functions with the crypto API, Barco Silex' developmement team took the choice to access the hardware from user space.
+Pressed by exhibitions and the complexity of interfacing the functions with the crypto API, Barco Silex' development team took the choice to access the hardware from user space.
 
-\noindent The BA414E can thus not be accessed using the same path as the BA411E and a user-space driver will be needed, that is the leftern path of figure~\ref{fig:os-path-generic-b}.
+\noindent The BA414E can thus not be accessed using the same path as the BA411E and a user-space driver will be needed, that is the path to the left of figure~\ref{fig:os-path-generic-b}.
 The complete work flow thus needs three elements:

### 2015-05-29 introduction.tex
corrections

@@
 In 1965, Gordon Moore, co-founder of Intel, stated that for the same footprint, the number of transistors in an integrated circuit should double every two years~\cite{moore1965}.
-It became the Moore's law, an empirical prediction that drived inovation for fifty years.
+It became the Moore's law, an empirical prediction that drived innovation for fifty years.
 Nowadays, it is reaching its limits, as Intel in delaying the manufacturing of its next-generation processors.
@@
 
-\noindent A report of the Natural Resources Defense Council~\cite{nrdc2014} forecasts the US datacenter electrity bill to reach \$13.7 billion by 2020.
+\noindent A report of the Natural Resources Defense Council~\cite{nrdc2014} forecasts the US datacenter electricity bill to reach \$13.7 billion by 2020.
 As CPUs are largely responsible for the power consumption, it makes their design critical.
@@
 % Power is critical nowadays, esapcially in datacenters.
-% Several solutions exists: abandonning general purpose processors altoghether and turn towards ARM cores, or add an FPGA to the processor for example.
+% Several solutions exists: abandonning general purpose processors altogether and turn towards ARM cores, or add an FPGA to the processor for example.
 % Intel proposed the latter solution in 2014~\cite{intel-xeon-fpga} and is on its way to expand that market with the merger with Altera~\cite{2015-intel-altera-merger}.
@@
 The present document will begin with the technical ground necessary to this work.
-After a presentation on how to drive a hardware device from the operating system and details on device driver design, we will adress the cryptography.
+After a presentation on how to drive a hardware device from the operating system and details on device driver design, we will address the cryptography.
 Only most relevant part of cryptography will be detailed: message digests, symmetric cryptography and more specifically the CBC and GCM modes of operation, and asymmetric cryptography with an accent on RSA and Diffie-Hellman.

### 2015-05-29 results.tex
corrections

@@
 \end{tabular}
-\caption{TLS connections per minute}{measures obtained with ten clients concurently connecting to an OpenVPN server.}
+\caption{TLS connections per minute}{measures obtained with ten clients concurrently connecting to an OpenVPN server.}
 \label{tab:tls-con}
@@
 
-\noindent RSA-2048 is a much more common configuration, espacially since the NIST deprecated RSA-1024 in 2013.
+\noindent RSA-2048 is a much more common configuration, especially since the NIST deprecated RSA-1024 in 2013.
 The full software implementation is visibly affected by the increase of the key size: the CPU usage doubles and the server processes three time less connections compared to the RSA-1024 results.
@@
 If we look at the efficiency, the software can process 0.24 connection per percentage of CPU usage, whilst the hardware can process 24 of them.
-The efficiency is thus multipled by a factor 1000.
+The efficiency is thus multiplied by a factor 1000.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The following tests are conducted after the connection has been established, so as the clients do not need to undergo any new key negociation.
+The following tests are conducted after the connection has been established, so as the clients do not need to undergo any new key negotiation.
 
@@
 When combining encryption and authentication, the software and hardware implementation are neck and neck.
-Eventhough the MAC is not offloaded, the hardware implementation could have performed better.
+Even though the MAC is not offloaded, the hardware implementation could have performed better.
 The reason is not only all the packets have to go through a virtual interface in the kernel, the payload also has to be sent to the hardware through the kernel.
@@
 
-\noindent It is interresting to note that for packets of 16000 bytes, the hardware adds an extra 12\% to the latency when using only encryption, but only 1\% when adding the authentication.
+\noindent It is interesting to note that for packets of 16000 bytes, the hardware adds an extra 12\% to the latency when using only encryption, but only 1\% when adding the authentication.
 In the second case, the synchronization between the confidentiality and the authentication slows the process down up to the point that the user/kernel space transfers are less an overhead than with confidentiality alone.
-If the MAC were to be done in hardware as well, the extra delay would probably be even larger as the data would have to be transfered to two different IP cores via two different drivers\footnote{A unifed driver is under development, gathering the encryption and authentication in the same kernel module.}.
+If the MAC were to be done in hardware as well, the extra delay would probably be even larger as the data would have to be transferred to two different IP cores via two different drivers\footnote{An unified driver is under development, gathering the encryption and authentication in the same kernel module.}.
 
@@
 \caption{IPsec: ping average response time -- software/Hardware}{\newline{}(a) Processing GCM in hardware yields at least as good results as CBC.
-Comparing the CBC mode without authentication in hardware with the GCM mode in software thus makes sensena yields 42\% less delay for larger payload size.\newline{}(b) With or without authentication, the hardware offers better results up to 13\% less delay.}
+Comparing the CBC mode without authentication in hardware with the GCM mode in software thus makes sense, and yields 42\% less delay for larger payload size.\newline{}(b) With or without authentication, the hardware offers better results up to 13\% less delay.}
 \label{fig:ping-benchmark-ipsec-soft-hard}
@@
 \caption{IPsec: ping average response time}{the latency is given in milliseconds.
-Even if there is no results for a hardware implementation of GCM, the result should have been close to those of \textit{ba411e-cbc256:none}.}
+Even if there are no results for a hardware implementation of GCM, the result should have been close to those of \textit{ba411e-cbc256:none}.}
 \label{tab:ping-benchmark-ipsec}
@@
 However, the process of GCM packets in hardware should be at least as fast as CBC packets.
-We can thus compare the software implementation of the GCM mode with a hardware implementation of CBC only (wihtout authentication).
+We can thus compare the software implementation of the GCM mode with a hardware implementation of CBC only (without authentication).
 The figure~\ref{fig:ping-benchmark-ipsec-gcm} shows an improved delay by up to 42\% when using the hardware.
@@
 
-\noindent In hardware, as OpenVPN could not take effeciently advantage of the accelaration of the encryption, IPsec is even faster, ranging from 25\% to 33\% faster than OpenVPN.
+\noindent In hardware, as OpenVPN could not take efficiently advantage of the acceleration of the encryption, IPsec is even faster, ranging from 25\% to 33\% faster than OpenVPN.
 
@@
 When only encryption is used, both implementation performs almost as well as outside the tunnel, but the software already saturates the CPU, whilst the hardware only uses half the same ressources.
-About 41\% of the operations using the CPU when accelarating with the hardware involve interuptions and kernel memory management.
+About 41\% of the operations using the CPU when accelerating with the hardware involve interruptions and kernel memory management.
 
@@
 
-As it was already the case for the ping benchmark, the overhead imposed by the manipulations of OpenVPN is extremely heavy: without encryption nor authentication, the CPU usage jumps from 7.16\% outide the tunnel to 42.60\% inside, and the throughput drops by 55\%, from 11.39MB/s to 5.18MB/s.
+As it was already the case for the ping benchmark, the overhead imposed by the manipulations of OpenVPN is extremely heavy: without encryption nor authentication, the CPU usage jumps from 7.16\% outside the tunnel to 42.60\% inside, and the throughput drops by 55\%, from 11.39MB/s to 5.18MB/s.
 
 \noindent Encrypting the data only changes the throughput by 8\%, but the CPU usage is almost doubled.
-Such a small decrease of performance is coherent with the results of the ping benchmark for the chunck of data size OpenVPN is supposed to process: around 1kB.
+Such a small decrease of performance is coherent with the results of the ping benchmark for the chunk of data size OpenVPN is supposed to process: around 1kB.
 % A fair interpretation would be that the encryption is not the bottleneck, the transfer bewteen the user and kernel mode, as well as the fragmentation, are.
@@
 \input{graphs/ipsec-ftp-bench}
-\caption{FTP file transfer over an IPsec tunnel}{The advantage of the hardware is a lower CPU usage, at the cost of a few percents lower throughput.
+\caption{FTP file transfer over an IPsec tunnel}{The advantage of the hardware is a lower CPU usage, at the cost of a few percent lower throughput.
 The GCM mode is also tested to discuss its implementation in hardware.
@@
 In both cases, the hardware implementation exhibited slightly lower performance, but a CPU usage three to four times lower.
-On embedded platforms, this is as important as the raw performance, because lower CPU usage means underclocked CPU, resulting in a lower power consumption.\newline{}
+On embedded platforms, this is as important as the raw performance, because lower CPU usage means under clocked CPU, resulting in a lower power consumption.\newline{}
 
-The softaware GCM results allow to open a discussion on this mode used in conjunction with IPsec.
+The software GCM results allow opening a discussion on this mode used in conjunction with IPsec.
 The GCM performance presented clearly shows a drop of throughput and an increase of CPU usage, illustrating the fact that those operations are hard on the software.
 With a hardware offload, we could expect not only a drastic drop of the CPU usage, but an increase of throughput as well, since the performance of the software is CPU-limited.
-As we already discussed with the latency results in section~\ref{sec:lantecy-ipsec}, the results of a hardware implementation of GCM should be at least as good as those of AES-CBC wihtout authentication.
+As we already discussed with the latency results in section~\ref{sec:lantecy-ipsec}, the results of a hardware implementation of GCM should be at least as good as those of AES-CBC without authentication.
 If it were the case, we would increase the throughput by at least 67\% and the decrease the CPU usage by at least 84\%.
@@
 Modern processor designers tend to add specialized instruction sets aimed at AES-GCM enhancement~\cite{intel-gcm2011}, making assembly implementation more efficient.
-Should further tests be conducted concerning IPsec paired with GCM, it would be wise to compare with an assembly implementation exploiting ARM NEON instruction set.
+Should further test be conducted concerning IPsec paired with GCM, it would be wise to compare with an assembly implementation exploiting ARM NEON instruction set.
 Some are being developed~\cite{Conrado2013,Danilo2013}, but none have been committed to the Linux kernel repository yet.
@@
 \input{graphs/ftp-bench-comparison}
-\caption{Comparison of file transfer methods}{Out of the three methods, OpenSSH is to be prefered if the raw performance is the objective, but IPsec offers a higher effeciency.
+\caption{Comparison of file transfer methods}{Out of the three methods, OpenSSH is to be prefered if the raw performance is the objective, but IPsec offers a higher efficiency.
 At the same time, OpenVPN is worse on all fronts.
@@
 In this work, we have two objectives: the raw performance and the CPU usage.
-The first objetive is fulfilled by OpenSSH; even if it is a user-space application that needs to transfer the data through the kernel in order to use the hardware, it still outperforms the software by 27\%.
+The first objective is fulfilled by OpenSSH; even if it is a user-space application that needs to transfer the data through the kernel in order to use the hardware, it still outperforms the software by 27\%.
 
@@
 A look at the figure~\ref{fig:openssl-speed} shows that fragmenting at small size is a bad idea, especially when the hardware is reached from the user space.
-OpenVPN sends one third of its packets at sizes that are highly ineffective, whilst IPsec manages a better frafmentation and limits the small packets to less than 10\% of the total.
+OpenVPN sends one third of its packets at sizes that are highly ineffective, whilst IPsec manages a better fragmentation and limits the small packets to less than 10\% of the total.
 

### 2015-05-29 technical-background.tex
corrections

@@
 \includegraphics[width=\linewidth]{barco-ba411e-ipcore}
-\caption{Barco Silex' AES IP Core}{There are two channels to comunicate with the IP core: the data and control paths~\cite{barco-ba411e}.}
+\caption{Barco Silex' AES IP Core}{There are two channels to communicate with the IP core: the data and control paths~\cite{barco-ba411e}.}
 \label{fig:barco-ba411e-ipcore}
@@
 
-A master interface has the advantage to offload data access operations to the device, but there is a ressource and area overhead to connect the DMA as well as an extra complexity to process its data.
+A master interface has the advantage to offload data access operations to the device, but there is a resource and area overhead to connect the DMA as well as an extra complexity to process its data.
 A slave interface is usually used for sporadic data (\textit{e.g.} status, control flags), avoiding the overhead of a DMA.
@@
 
-If a master interface with a DMA is used, the operating system buils scatterlists in the kernel-space memory, then map those pages to memory descriptor that have a physical address on the DMA.
+If a master interface with a DMA is used, the operating system builds scatterlists in the kernel-space memory, then map those pages to memory descriptor that have a physical address on the DMA.
 They can be mapped three different ways: \texttt{DMA\_BIDIRECTIONAL}, \texttt{DMA\_TO\_DEVICE} or \texttt{DMA\_FROM\_DEVICE}.
-When the CPU write something in those descriptors and synchronize them with the DMA, it does not have to care about them anymore, the DMA in now in charge to send them to the device where registers are ready to read the incomming data.
+When the CPU write something in those descriptors and synchronize them with the DMA, it does not have to care about them any more, the DMA in now in charge to send them to the device where registers are ready to read the incoming data.
 The same goes from the device to the CPU: when the device wants to communicate data to the OS, it writes it on the DMA that will transfer them to the CPU, triggering a flag on the way to notify it.
@@
 Under Unix, the kernel runs under the highest level and it has no restriction, whereas applications run on the lowest level.
-We call those two levels kernel space and user space, segragating the privilege and the memory between the two modes (\citet{tanenbaum2014} and \citet[chap. 2]{Corbet:2005:LDD:1209083}).
+We call those two levels kernel space and user space, segregating the privilege and the memory between the two modes (\citet{tanenbaum2014} and \citet[chap. 2]{Corbet:2005:LDD:1209083}).
 The figure~\ref{fig:user-kernel} shows this separation.
@@
 \input{user-kernel}
-\caption{User and kernel space separation}{The applications and libraries from the the user space communicates with the kernel space via the system calls.}
+\caption{User and kernel space separation}{The applications and libraries from the user space communicates with the kernel space via the system calls.}
 \label{fig:user-kernel}
@@
 10]{Corbet:2005:LDD:1209083}.
-The second case is the easiest and is always guaranteed to work, but will monopolize the processir, loading it at 100\%, and avoiding any other task to be executed.
-Fortunately, modern monolithic kernels such as the Linux kernel from 2.6 provide preemptive scheduling~\cite{Santhanam2003}, \textit{i.e.} the scheduler interrupts the running task and assigns the processor ressources it used to an other one.
-Hence, systems with a lot of processes in need for CPU ressources would not be stalled, but it would not change anything if the only process heavily requesting processor time is the one using the driver.\newline{}
+The second case is the easiest and is always guaranteed to work, but will monopolize the processor, loading it at 100\%, and avoiding any other task to be executed.
+Fortunately, modern monolithic kernels such as the Linux kernel from 2.6 provide preemptive scheduling~\cite{Santhanam2003}, \textit{i.e.} the scheduler interrupts the running task and assigns the processor resources it used to another one.
+Hence, systems with a lot of processes in need for CPU resources would not be stalled, but it would not change anything if the only process heavily requesting processor time is the one using the driver.\newline{}
 
@@
 In order to add such an implementation, a kernel has to be developed and register itself with the Linux Crypto API upon its loading.
-Those kernel modules can either be software implementation of some cryptogrpahic ciphers, or a device driver that will offload those operation to a dedicated hardware device.\newline{}
+Those kernel modules can either be software implementation of some cryptographic ciphers, or a device driver that will offload those operations to a dedicated hardware device.\newline{}
 
@@
 Now rises the question of what and when authenticating.
-\citet{Bellare2000} prooved that the most secure solution is to encrypt then compute the MAC from the ciphertext.
+\citet{Bellare2000} proved that the most secure solution is to encrypt then compute the MAC from the ciphertext.
 We will see in section~\ref{sec:theory-network} that if IPsec follows this recommandation, SSL/TLS does not and MAC first the plaintext then encrypt the message.
@@
 
-Eventhough the signature algorithms are often more efficient than the usual MACs and the key pairs key stay in use for very long period of time, asymmetric cryptography can not supplant its symmetric counterpart.
+Even though the signature algorithms are often more efficient than the usual MACs and the key pairs key stay in use for very long period of time, asymmetric cryptography can not supplant its symmetric counterpart.
 The throughput rates of modern asymmetric schemes are still well under the symmetric schemes, plus the key sizes are much larger.
@@
 
-In practice, public-key cryptography is used to estalish a secure channel between two or more peers and exchange a shared secret key that will be used the encrypt the subsequent communications.
+In practice, public-key cryptography is used to establish a secure channel between two or more peers and exchange a shared secret key that will be used the encrypt the subsequent communications.
 
@@
 \begin{enumerate}
-	\item randomy choose two large primes $p$ and $q$;
+	\item randomly choose two large primes $p$ and $q$;
 	\item compute the modulus $n = p q$, and consequently we have \[\phi(n) = (p-1)(q-1)\]with $\phi(n)$ as the Euler function;
@@
 
-The scheme is based on modular artihmetic, which is secure because of the factoring problem.
+The scheme is based on modular arithmetic, which is secure because of the factoring problem.
 For an opponent to decrypt a ciphertext $c$ wihtout knowing the private key $(d, n)$, he would need to compute the $e^{th}$ roots of $c$, modulo $n$.
@@
 Imagine that somehow, an opponent lays his hand on the shared secret.
-If that secret has already been used, he can decipher all data transfered during past connections.
-However, if the secret is new for every new connection, the compromission of the shared secret does not jeopardize past communications.
+If that secret has already been used, he can decipher all data transferred during past connections.
+However, if the secret is new for every new connection, the compromise of the shared secret does not jeopardize past communications.
 This is perfect forward secrecy: using a new key to protect the previous messages.
@@
 We will see in chapter~\ref{chap:results} that while a 1024-bit key size is easily manageable by full software implementation, hardware offloading become a necessity for 4096-bit key sizes.
-Moreover, 1024-bit parameter size, both RSA and Diffie-Hellman, are disallowed by the NIST recommandations since 2013~\cite{nist-sp800-131A}.
+Moreover, 1024-bit parameter size, both RSA and Diffie-Hellman, are disallowed by the NIST recommendations since 2013~\cite{nist-sp800-131A}.
 
@@
 % Present the TCP/IP layering
-Computer networks are described by two models: the TCP/IP stack and the OSI model, each splitting the network workflow into more or less abstract layers.
+Computer networks are described by two models: the TCP/IP stack and the OSI model, each splitting the network work flow into more or less abstract layers.
 The RFC 1122~\cite{rfc1122} defines the TCP/IP, and the international standard ISO/IEC 7498-1~\cite{ISOIEC7498} defines the OSI stack as in the table~\ref{tab:tcp-ip-stack}.
@@
 \end{tabularx}
-\caption{TCP/IP and OSI model comparison}{They are globaly the same, except for the application layer of the TCP/IP stack which merge togheter the three upper layers of the OSI model.
-Between parenthsis are examples of protocols resting on each layer.}
+\caption{TCP/IP and OSI model comparison}{They are globally the same, except for the application layer of the TCP/IP stack which merge together the three upper layers of the OSI model.
+Between parenthesis are examples of protocols resting on each layer.}
 \label{tab:tcp-ip-stack}
@@
 
-The present work will focus and SSL/TLS and IPsec, the figure~\ref{fig:ipsec-tls} showing their respective integration in the network stack.
+The present work will focus and SSL/TLS and IPsec, the figure~\ref{fig:tls-ipsec} showing their respective integration in the network stack.
 
@@
 The last version of SSL (version 3.0) dates from 1996~\cite{rfc6101} and has been found a critical vulnerability by Google in 2014~\cite{poodle}.
-As a consequence, SSL 3.0 support is being dropped, and the IETF is working on a depracation of the standard~\cite{sslv3-diediedie-03}.
+As a consequence, SSL 3.0 support is being dropped, and the IETF is working on a deprecation of the standard~\cite{sslv3-diediedie-03}.
 
@@
 
-\noindent The figure~\ref{fig:tls-handshake} shows the messages exchanged during the establishement of a connection between two peers.
+\noindent The figure~\ref{fig:tls-handshake} shows the messages exchanged during the establishment of a connection between two peers.
 
@@
 	It can send its own certificate and request the client's.
-	In some cases, the server can already begin the key echange protocol (\textit{e.g.} Diffie-Hellman).
+	In some cases, the server can already begin the key exchange protocol (\textit{e.g.} Diffie-Hellman).
 	\item Client: ClientKeyExchange and others.
-The client veryfies the validity of the server certificate, if any, and sends its own.
+The client verifies the validity of the server certificate, if any, and sends its own.
 If necessary, the client may have to change its cipher specification based on the constraints of the server, or request a new specification from the server.
@@
 
-Once the connection is established, the peers can exchange application data using the symmetric key negociated during the handshake.
+Once the connection is established, the peers can exchange application data using the symmetric key negotiated during the handshake.
 
@@
 IPsec is a protocol whose implementation is a modification of the IP stack in the kernel space.
-It is a network/internet level security; it examines incomming IP packets and checks if there exists a security association with the destination, and decrypt it on-the-fly if necessary.
+It is a network/internet level security; it examines incoming IP packets and checks if there exists a security association with the destination, and decrypt it on-the-fly if necessary.
 
@@
 
-\noindent The IKEv2 protocol uses four different types of exchanges to fulfill its role:
+\noindent The IKEv2 protocol uses four different types of exchanges to fulfil its role:
 \begin{itemize}
@@
 \noindent The ESP payload includes an optional IV and an optional Traffic Flow Confidentiality (TFC) padding.
-The TFC padding allows to modify the length of the packet with dummy data, somewhat hiding the true nature of the payload.
+The TFC padding allows modifying the length of the packet with dummy data, somewhat hiding the true nature of the payload.
 It differs from the standard padding because its length is not limited to 255 bytes like the latter.
@@
 Since ESP can use a null cipher, it arguably offers the same features as AH.
-The existance of AH is actually historical; when the protocol was developed in the 1990s, there were laws in the United States and other countries preventing the export of product including encryption features.
+The existence of AH is actually historical; when the protocol was developed in the 1990s, there were laws in the United States and other countries preventing the export of product including encryption features.
 Having an authentication only counterparty allowed such product to be exported without encryption capabilities.\newline{}
@@
 \begin{itemize}
-	\item Transport for end-to-end service and protecting the payload. IPsec examines the incomming payload, checks it has not been tampered with and pass it to the next layer.
+	\item Transport for end-to-end service and protecting the payload. IPsec examines the incoming payload, checks it has not been tampered with and pass it to the next layer.
 	\item Tunnel protects the entire IP packet by encapsulating it into a new IP packet.

### 2015-05-29 test-proto.tex
corrections

@@
 The experimental environment is built around a standard x86 host and an ARM Cortex-A9 alongside an Altera Cyclone V FPGA as the target.
-Both are linked toghether through a network capped by 100Mbps switches.
+Both are linked together through a network capped by 100Mbps switches.
 Both stations have gigabits ethernet interface and could hence be directly connected to each other, but in that case the communication would be limited by the I/O transfers of the storage units -- a hard drive disk in one case, an micro-SD card in the second -- on which we can not depend to set a constant throughput limitation, as it is highly influenced by the data block size and general health of the support.
@@
 \subsection{x86 host}
-The desktop host runs on Windows 7 Professional 64-bit, but a virtual machine using a Linux distribution is used for the developement and testings.
+The desktop host runs on Windows 7 Professional 64-bit, but a virtual machine using a Linux distribution is used for the development and testings.
 
@@
 
-\noindent The user should also be aware that if Streamline if a very powerfull and versatile tool, it is not reliable at very high resolution.
-The data should not be analyzed in windows smaller than a millisecond.
+\noindent The user should also be aware that if Streamline if a very powerful and versatile tool, it is not reliable at very high resolution.
+The data should not be analysed in windows smaller than a millisecond.
 
@@
 
-The table~\ref{tab:software-version} summarizes the softwares used and tjeir respective version.
+The table~\ref{tab:software-version} summarizes the software used and their respective version.
 
@@
 
-\noindent As AES mode CBC and SHA-2 are widely used, they will be used with a serious security paramter using keys of 256-bit.
+\noindent As AES mode CBC and SHA-2 widely are used, they will be used with a serious security parameter using keys of 256-bit.
 Some test case will also experiment some implementation of AES-GCM to compare its performance and justify the need for its offloading in hardware.
@@
 
-All the tests will be conducted with the ESP protocol in tunnel mode, so that we have the worst case scenario; as AH imposes a smaller overhead, the perfomance could only be better.
+All the tests will be conducted with the ESP protocol in tunnel mode, so that we have the worst case scenario; as AH imposes a smaller overhead, the performance could only be better.
 
@@
 This benchmark is done only with OpenVPN 2.3.6.
-Since there is no standard support for those operation in the kernel yet, it would not have made sense to use IPsec, since it would have had to fallback to OpenSSL, then following the same path as OpenVPN do.
+Since there is no standard support for those operations in the kernel yet, it would not have made sense to use IPsec, since it would have had to fallback to OpenSSL, then following the same path as OpenVPN do.
 As soon as the public key operations can be plugged into the Linux Crypto API, this use case should however be tested.
@@
 For this use case, the board is configured in server mode, so that it can accept connections from any client.
-The server does not connect with a specific client and waits for incomming connection requests (see listing~\ref{list:openvpn-server}).
+The server does not connect with a specific client and waits for incoming connection requests (see listing~\ref{list:openvpn-server}).
 The virtual machine will then execute then clients in parallel using the script~\ref{list:openvpn-client-script}.
 The only option differentiating the clients is their IP address and port number.
-Otherwise, all the clients share the same basic configuration file (see listing~\ref{list:openvpn-client}), which tell them to renogociate a new connection every second.
+Otherwise, all the clients share the same basic configuration file (see listing~\ref{list:openvpn-client}), which tell them to renogotiate a new connection every second.
 Hence, if a connection could be made with no delay and if the processes scheduling were ideal, the server would have to address 600 connections per minute.
@@
 
-The security parameter tested is a standard TLS-DHE-RSA, hence forcing the peers the renogociate a new shared secret and ephemeral Diffie-Hellman parameters at each connection attempt.
+The security parameter tested is a standard TLS-DHE-RSA, hence forcing the peers the renogotiate a new shared secret and ephemeral Diffie-Hellman parameters at each connection attempt.
 
@@
 \subsection{File transfer}
-The file transfered is an uncompressed block of 128MB of random data generated using the following command:
+The file transferred is an uncompressed block of 128MB of random data generated using the following command:
 \begin{lstlisting}[language=bash]
@@
 
-For OpenSSH, the commmand \texttt{scp} will be used to transfer the data securely over an SSH tunnel.
+For OpenSSH, the command \texttt{scp} will be used to transfer the data securely over an SSH tunnel.
 As for OpenVPN and IPsec, a tunnel will be established beforehand, and then the simple \texttt{ftp} command will allow the client to fetch the file on the server.
@@
 All the results on the throughput and CPU usage take into account only the file transfer.
-The server authentication, the connection establishement and any other security negociation is ignored in this particular test case.
+The server authentication, the connection establishment and any other security negotiation is ignored in this particular test case.

### 2015-06-21 presentation/ipsec-ftp-cpu.tex
ipsec file transfer

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        % title = {FTP transfer inside an IPSec tunnel},
        width  = \textwidth,
        height = 0.75\paperheight,
        major x tick style = transparent,
        xbar=2pt,
        bar width=8pt,
        % enlarge x limits={abs=1},
        xmajorgrids = true,
        xlabel = {CPU usage},
        ylabel = {},
        % ylabel near ticks,
        % yticklabel pos=right,
        xmin=0, xmax=100,
        symbolic y coords={cipher null, aes:none, aes:sha},
        ytick = data,
        ymajorticks=false,
        scaled x ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.15)}, anchor=north,legend columns=4},
        % enlarge x limits=0.1,
    ]

\addplot[style={black,fill=LimeGreen,postaction={pattern=north east lines},mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (0,cipher null)
        (63.74,aes:none)
        (74.64,aes:sha)
    };
    \label{software-cpu}

\addplot[style={black,fill=gray,postaction={pattern=north east lines},mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (7.16,cipher null)
        (0,aes:none)
        (0,aes:sha)
    };
    \label{out-of-tunnel-cpu}%"oot" for "out of tunnel"

\addplot[style={black,fill=RedOrange,postaction={pattern=north east lines},mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (0,cipher null)
        (14.87,aes:none)
        (17.25,aes:sha)
    };
    \label{ba411e-cpu}

\addplot[style={black,fill=NavyBlue,postaction={pattern=north east lines},mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (14.68,cipher null)
        (0,aes:none)
        (0,aes:sha)
    };
    \label{inside tunnel-cpu}%"it" for "in tunnel"
% \legend{software, ba411e, out-of-tunnel, inside tunnel}
\end{axis}

### 2015-06-21 presentation/ipsec-ftp-throughput.tex
ipsec file transfer

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        % title = {FTP transfer inside an IPSec tunnel},
        width  = \textwidth,
        height = 0.75\paperheight,
        major x tick style = transparent,
        xbar=2pt,
        bar width=8pt,
        % enlarge x limits={abs=1},
        xmajorgrids = true,
        xlabel = {Throughput [MB/s]},
        ylabel = {},
        ylabel near ticks,
        yticklabel pos=right,
        xmin=0, xmax=15,
        symbolic y coords={cipher null, aes:none, aes:sha},
        ytick = data,
        scaled x ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.15)}, anchor=north,legend columns=4},
        % enlarge x limits=0.1,
    ]

\addplot[style={black,fill=ForestGreen,mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (0,cipher null)
        (8.83,aes:none)
        (6.47,aes:sha)
    };
    \label{software}

\addplot[style={black,fill=black,mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (11.39,cipher null)
        (0,aes:none)
        (0,aes:sha)
    };
    \label{out-of-tunnel}%"tp" for "throughput"

\addplot[style={black,fill=BrickRed,mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (0,cipher null)
        (8.52,aes:none)
        (5.80,aes:sha)
    };
    \label{ba411e}

\addplot[style={black,fill=MidnightBlue,mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (10.21,cipher null)
        (0,aes:none)
        (0,aes:sha)
    };
    \label{inside tunnel}
% \legend{software, ba411e, out-of-tunnel, inside tunnel}
\end{axis}

### 2015-06-21 presentation/openvpn-ftp-cpu.tex
Openvpn file transfer

--- 2015-06-21 presentation/ipsec-ftp-cpu.tex
@@
         % yticklabel pos=right,
-        xmin=0, xmax=100,
+        xmin=0, xmax=115,
+        xtick={50, 100},
         symbolic y coords={cipher null, aes:none, aes:sha},
@@
         (0,cipher null)
-        (63.74,aes:none)
-        (74.64,aes:sha)
+        (76.60,aes:none)
+        (76.03,aes:sha)
     };
@@
         (0,cipher null)
-        (14.87,aes:none)
-        (17.25,aes:sha)
+        (83.74,aes:none)
+        (80.89,aes:sha)
     };
@@
     coordinates {
-        (14.68,cipher null)
+        (42.60,cipher null)
         (0,aes:none)

### 2015-06-21 presentation/openvpn-ftp-throughput.tex
Openvpn file transfer

--- 2015-06-21 presentation/ipsec-ftp-throughput.tex
@@
         scaled x ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
-        legend style={at={(0.5,-0.15)}, anchor=north,legend columns=4},
+        legend style={at={(1.3,-0.1)}, anchor=north,legend columns=2},
         % enlarge x limits=0.1,
@@
         (0,cipher null)
-        (8.83,aes:none)
-        (6.47,aes:sha)
+        (4.78,aes:none)
+        (3.87,aes:sha)
     };
@@
         (0,cipher null)
-        (8.52,aes:none)
-        (5.80,aes:sha)
+        (3.35,aes:none)
+        (2.84,aes:sha)
     };
@@
     coordinates {
-        (10.21,cipher null)
+        (5.18,cipher null)
         (0,aes:none)
@@
     \label{inside tunnel}
-% \legend{software, ba411e, out-of-tunnel, inside tunnel}
+\legend{\small software, raw, ba411e, OpenVPN}
 \end{axis}

### 2015-06-22 presentation/openvpn-tls-connections.tex
First draft presentation

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        % title = {FTP transfer inside an IPSec tunnel},
        width  = \textwidth,
        height = 0.65\paperheight,
        major x tick style = transparent,
        xbar=2pt,
        bar width=8pt,
        % enlarge x limits={abs=1},
        xmajorgrids = true,
        xlabel = {Connections per minute},
        ylabel = {},
        ylabel near ticks,
        yticklabel pos=right,
        xmin=0, xmax=600,
        xtick={0,100,200,300,400,500},
        symbolic y coords={RSA-1024, RSA-2048, RSA-4096},
        ytick = data,
        scaled x ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(1.3,-0.1)}, anchor=north,legend columns=1},
        % enlarge x limits=0.1,
    ]

\addplot[style={black,fill=ForestGreen,mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (445.4,RSA-1024)
        (155.6,RSA-2048)
        (19.6,RSA-4096)
    };
    \label{soft-tp}

\addplot[style={black,fill=BrickRed,mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (509.3,RSA-1024)
        (420.9,RSA-2048)
        (115.5,RSA-4096)
    };
    \label{ba411e-tp}
\legend{\small software, ba411e}
\end{axis}

### 2015-06-22 presentation/openvpn-tls-cpu.tex
First draft presentation

%%%%%%%%%%%%%%%%%%%%%%%%
% throughput
%%%%%%%%%%%%%%%%%%%%%%%%
\begin{axis}[
        % title = {FTP transfer inside an IPSec tunnel},
        width  = \textwidth,
        height = 0.65\paperheight,
        major x tick style = transparent,
        xbar=2pt,
        bar width=8pt,
        % enlarge x limits={abs=1},
        xmajorgrids = true,
        xlabel = {CPU usage},
        ylabel = {},
        % ylabel near ticks,
        % yticklabel pos=right,
        xmin=0, xmax=115,
        xtick={50, 100},
        symbolic y coords={RSA-1024, RSA-2048, RSA-4096},
        ytick = data,
        ymajorticks=false,
        scaled x ticks = false,%Disable the *10^4 exponent applied to all y axis markings.
        legend style={at={(0.5,-0.15)}, anchor=north,legend columns=4},
        % enlarge x limits=0.1,
    ]

\addplot[style={black,fill=LimeGreen,postaction={pattern=north east lines},mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (40.32,RSA-1024)
        (92.14,RSA-2048)
        (81.97,RSA-4096)
    };
    \label{software}

\addplot[style={black,fill=RedOrange,postaction={pattern=north east lines},mark=none}, nodes near coords={\pgfmathfloatifflags{\pgfplotspointmeta}{0}{}{\pgfmathprintnumber{\pgfplotspointmeta}}}]
    coordinates {
        (13.29,RSA-1024)
        (4.82,RSA-2048)
        (4.78,RSA-4096)
    };
    \label{ba414e}
% \legend{software, ba411e, out-of-tunnel, inside tunnel}
\end{axis}
